### 2021-01-07 3_r_scripts/1_markdown_document.Rmd
Setting up project.

---
title: "Can we distinguish between the speed and scope of acute glucocorticoid responses?"
author: "Conor Taff, Laura Schoenle, and Maren Vitousek"
date: "Last Updated: `r format(Sys.time(), '%B %d %Y at %H:%M')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(here)
library(sjPlot)
```

```{r}
sessionInfo()
```

### 2021-01-07 3_r_scripts/1_main_script.R

@@ -1,92 +1,3 @@
 ###
-# I always start my scripts with a header giving a brief description of the purpose of the script.
-# In this case, the goal is to make a template for new analyses in the lab and to describe why
-# parts of the project folder are organized the way they are.
-#
-# Written by: Sabrina McNew & Conor Taff
-# Last updated: 10/13/2020
-# Run under R Studio 1.1.463 on Mac OSX 10.11.6
-# See markdown file for saving session info. If you aren't working in markdown it is also 
-# good practice to record package versions here.
-
-# Load packages ----
-  # Ending the line above with four dashes or four hashtags puts the chunk into your outline below. 
-  # This makes it a lot easier to find things later in a long script. Another way to create sections 
-  # is with "command + shift + R" in Rstudio. 
-
-  # I like to load all packages up front using 'pacman'. This way you can see right at the beginning
-  # what is required and you are less likely to break your code when modifying it by moving a line
-  # up to before the required package was loaded. I have a few packages I almost always load, but most
-  # of these would change each time depending on the project. Just remember to add packages to the list
-  # up here when you want to use them. You'll get an error if these packages aren't installed on your local
-  # machine yet and will need to run install.packages(xxx)
-    pacman::p_load(tidyverse, here, plyr, sjPlot)
-      # 'here' is useful for referring to relative path names in your project directory.
-
-# Load data ----
-  # here is where I would load any raw data that you have added to the project
-  # I'm reading in two files from
-    #https://knowledge.domo.com/Training/Self-Service_Training/Onboarding_Resources/Fun_Sample_Datasets
-    d_hero <- read.delim(here::here("1_raw_data/raw_hero_data.txt"))      #super hero attributes
-    d_power <- read.delim(here::here("1_raw_data/raw_hero_power.txt"))    #super hero powers
-
-# Clean up data
-    # Make any kind of exclusions you want here and document them well. Apply transformations to columns or 
-    # standardize them in some way.
-    # You might change column names here too (though maybe that should be done when you make the files). 
-      colnames(d_power)[1] <- "name"    #I changed this to match the other data file.
-        
-# Set values ----
-    
-    # Any values that you are setting that will be used through the script should go here. For example
-    # if you have a control and treatment group, you can set plotting colors for them here. Then everywhere
-    # you plot below you can refer to these objects. That way changing the color for all treatment plots can
[…]

### 2021-01-08 3_r_scripts/1_main_script.R
updated simulation

@@ -1,3 +1,290 @@
-###
-# 
+### Script for speed vs. scope manuscript
+# Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
+
+
+## Load packages ----
+    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
+
+## Load RWBL field data ----
+        rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
+        rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
+    
+    ## Replacing labeling on a couple rows to make them conform to level names
+        rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
+        rw_dat$LH.stage[38]<-"early-breeding"
+    
+    ## Make a list of colors in order of rows to plot LH stages in different colors
+        collist<-rep(NA,nrow(rw_dat))
+        for(i in 1:nrow(rw_dat)){
+          if(rw_dat$LH.stage[i]=="early-breeding"){collist[i]<-"lightblue"}
+          if(rw_dat$LH.stage[i]=="late-breeding"){collist[i]<-"gray60"}
+          if(rw_dat$LH.stage[i]=="molt"){collist[i]<-"red"}
+        }
+
+## Analysis of RWBL field data ----
+        ## Calculate slope of increase on log scale in first 8 minutes for each sample
+            for(i in 1:nrow(rw_dat)){
+              calc <- as.data.frame(matrix(nrow = 3, ncol = 3))
+              colnames(calc) <- c("time", "cort", "lcort")
+              calc[, 1] <- t(rw_dat[i, c("tube1_tsec", "tube2_tsec", "tube3_tsec")])
+              calc[, 2] <- t(rw_dat[i, c("tube1_cort", "tube2_cort", "tube3_cort")])
+              calc[, 3] <- log(calc[, 2] + 1)
+              rownames(calc) <- c(1, 2, 3)
+              calc <- subset(calc, calc$time < 600 & is.na(calc$cort) == FALSE)
+              calc$time2 <- calc$time / 60
+              ifelse(nrow(calc) > 1, slope <- coef(lm(calc$cort ~ calc$time2))[2], slope <- NA)
+              rw_dat$slope[i] <- slope
[…]

### 2021-01-08 3_r_scripts/1_markdown_document.Rmd
updated simulation
Lots of new plots.

@@
 knitr::opts_chunk$set(echo = TRUE)
-library(knitr)
-library(dplyr)
-library(ggplot2)
-library(here)
-library(sjPlot)
 ```
 
+# Red-Winged Blackbird Analysis
+
+![](rwbl.png)\
+
+Data from 52 RWBL that we caught in 2016 & 2017 in Ithaca between 18 April and 22 July.
+We tried to take a 7 point stress series with samples at 1, 3, 7, 15, 30, 45, 60 minutes.
+The exact time points vary a little bit though and we use the exact measures in all the plots and analyses.
+Some samples are also missing from the end or middle of some time series.
+
+## Descriptive Plots
+
+First, I'm just plotting the full stress series on the raw and log scale.
+There are a lot of possible choices here for how best to display this, so I'm dropping in a few different options at this point.
+The y axis scale is very different for the three breeding stages, which makes it hard to compare the shapes.
+There are a few options with changing the scales or log transforming or plotting in different facets that make it easier to compare.
+
+![](basic_rwbl_profile.png)\
+
+**Figure x.** Time course of corticosterone measurements in each life history stage by individual (gray lines and points) and by group (red line and confidence interval).
+Note that y axes differ between the plots.
+
+![](basic_rwbl_profile_log.png)\
+
+**Figure x.** Same as above but on log scale.
+
+![](summary_lines.png)\
+
+**Figure x.** Lines on the same panels. Left is normal scale right is log scale.
+
+## Response Shape by Life History
+
+A lot of classic studies ask about differences in absolute cort values at different life history stages (e.g., find some Wingfield studies).
+We can see here that there are big differences between early breeding vs.
+late breeding vs.
+molt.
+For the most part, early breeding is higher than the other two across the board.
+We can also ask though whether the shape of the stress response differs or not between stages.
+For this I'm plotting the responses as a percentage of maximum for each individual over our seven point time series.
+This lets you see very clearly how the shape of the time course differs between stages.
+
+I could do some actual statistics on this, but basically here you can see that:  
+- Early breeders start with a baseline value that is already at a higher percentage of their maximum value.  
+- Early & Late breeders achieve a higher percentage of their maximum value faster (i.e., they have a 'faster' cort response).  
+- Molting birds achieve maximum value later (i.e., slower response).  
+- As a result of those last two points, molting birds actually sustain their maximal *rate* of increase longer than the other stages.
+
+![](rwbl_s_profiles.png)\
+
+**Figure x.** Time course of corticosterone as a percentage of maximum corticosterone achieved.
+Panel A shows loess regressions for each individual sampled.
+Panel B shows the combined loess regressions with 95% confidence interval for each life history stage.
+
+![](rwbl_s_profiles_log.png)\
+
+**Figure x.** Same as above but on log scale.
+
+## How much variation?
+
+One thing that we might want to ask is just how much variation is there between individuals in the speed vs.
+the scope of cort responses?
+Is one more variable than the other?
+There might be different ways to get at this, but for now I'm just calculating the coefficient of variation for the slope, the time taken to reach 95% of maximal, and the maximal value (and baseline value) for each individual.
+Because the absolute values differ so much by stage I'm doing this separetly for each stage.
+
 ```{r}
-sessionInfo()
+cv_tab <- readRDS(here::here("3_r_scripts/cv_table.rds"))
+knitr::kable(cv_tab, digits = 1, align = "lccccccc")
 ```
+
+I don't love this way of comparing, but you can get a sense here of how much variation there is relative to the mean of each measure.
+It looks like there is at least a comparable amount of variation in initial slope as there is in max cort and maybe base cort has a bit more.
+In contrast, there isn't a huge amount of variation in the time it takes to reach 95% of maximal.
+
+## Slope vs. Max
+
+Here I'm asking whether the steepness of the initial slope of increase (in first 8 minutes) is correlated with the max cort value eventually achieved.
+Plotting both on the normal and on log scale and separately for the different life history stages.
+I'm not sure how much sense it really makes to look at this on the log scale, but I guess it is kind of what people are doing in a lot of analyses if they log transform without really thinking about it.
+Anyway, I'm plotting it just to look.
+
+![](slope_vs_max.png)\
+
+**Figure x.** Relationship between the slope in the first three samples (~8 minutes) and the maximal corticosterone value achieved.
+Panel A shows all samples combined and B-D show the three stages separately.
+R2 and P values are from simple linear models.
+The one outlier with very high cort is included, I think excluding would make the early breeding relationship significant.
+
+![](slope_vs_max_log.png)\
+
+**Figure x.** Same as above on log scale.
+
+## Slope vs. Time to Max
+
+Here I'm plotting the initial slope vs the elapsed time before each bird reached 95% of their maximal cort from the loess regressions.
+I do 95 rather than 100 because a lot of birds are still going up at the end of the 60 minutes.
+I guess it might be worth doing this at different numbers like 50 or 75% too...
+
+![](slope_vs_95.png)\
+
+**Figure x.** Normal scale
+
+![](slope_vs_95_log.png)\
+
+**Figure x.** Log scale
+
+## RWBL Conclusions
+
+There are probably some different things we could do with the data above, but I think we can make some general conclusions relevant to the speed vs. scope question.
+
++ In addition to varying in absolute cort measures across life stages, the actual shape of the cort response also seems to differ as life stage has slightly different effects on the speed vs.
+the scope of the cort response.
++ Within a life history stage, there is variation within individuals in both the speed and scope of the cort response, and it isn't clear that one of these aspects is substantially more or less variable than the other.
++ Pooling across contexts, the initial slope and maximal value of corticosterone are correlated, but the relationship is quite weak within each separate life stage, suggesting that the two can vary somewhat independently.
++ There is little evidence that initial slope is directly related to time to achieving 95% of the maximal value, but the time series here does not capture the downward return to base cort and there was relatively little variation in the time required to reach 95%, so this may not be surprising.
+It may be better to look at the time to reach 70%, since that seems to be the part of the curve where individuals and stages differ most from each other.
++ Measuring only at 30 minutes misses a lot of these features of the cort response.
+
+# Simulation
+
+This is already described in more detail in the manuscript file I started, but basically the simulation works in two steps.
+I'm not sure it is worth making an actual package, but I could use some catchy function names.
+A package could be something like `cortResponse`.
+Right now my functions are just called `cort_sim1` and `cort_sim2`.
+
+First, one function generates a population with 'true' values for base, speed, max, time at plateau, and time at return to base for each along with a 'performance' metric (i.e., fitness) that is derived from those true underlying values.
+You can set the mean and sd for each parameter, the number of animals to generate, and the degree of correlation between each parameter.
+Animals are then drawn from a multivariate normal distribution with the set values.
+Note that the function will allow you to set any values, but a lot of combinations won't work if the produce negative values or if different traits are set to have strong correlations in a way that makes it impossible to generate a multivariate normal distribution; a message is given in these cases to try different values.
+I think I could extend this first function to also have slopes across a gradient for some parameters (e.g., the 'true' slope of response in base cort across a temperature gradient).
+
+Second, the output from the first function is used to simulate corticosterone time courses for each animal (or multiple times for each animal).
+In this function you can set the degree of error from the 'true' value for each parameter.
+This can be thought of as measurement error or as individual variation associated with repeat sampling of the same animal.
+After this sampling, a loess smoothed curve is fit for each series of parameters for each animal and then 'cort' values are recorded at each minute of the full time course.
+The function saves a list with 1) the full time course for each sample, 2) a downsampled data set that has cort measured at just the selected time points (by default 2 & 30 minutes), and 3) a rank ordered time course to visualize the degree of crossing.
+
+Using this simulation, we can explore a lot of different scenarios and ask how changing different parameters or sampling regimes might influence our ability to detect 'true' effects.
+There are actually so many possibilities here that I'm not sure what makes the most sense to focus on:
+
+- Sample different 'species' or 'populations' that have different shape, max, timing, or correlation between different parameters.
+Can you detect those differences?
+How much is ability to detect influenced by sampling regime?
+- Sample 1/2/3 time points. Can you detect speed vs. max separately? Change variation in time to reach max etc and see how that changes.
+
+## Demonstration
+
+## Speed vs. Scope

### 2021-01-08 3_r_scripts/1_RWBL_script.R

@@ -1,0 +1,535 @@
+### Script for speed vs. scope manuscript
+# Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
+
+
+## Load packages ----
+    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
+
+## Load RWBL field data ----
+        rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
+        rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
+    
+    ## Replacing labeling on a couple rows to make them conform to level names
+        rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
+        rw_dat$LH.stage[38]<-"early-breeding"
+    
+    ## Make a list of colors in order of rows to plot LH stages in different colors
+        collist<-rep(NA,nrow(rw_dat))
+        for(i in 1:nrow(rw_dat)){
+          if(rw_dat$LH.stage[i]=="early-breeding"){collist[i]<-"lightblue"}
+          if(rw_dat$LH.stage[i]=="late-breeding"){collist[i]<-"gray60"}
+          if(rw_dat$LH.stage[i]=="molt"){collist[i]<-"red"}
+        }
+
+## Analysis of RWBL field data ----
+        ## Loop through data frame to calculate slope and max cort achieved for each bird's samples
+            for(i in 1:nrow(rw_dat)){
+              # Set up dataframe for each bird with just the first three samples (less than 10 minutes)
+                calc <- as.data.frame(matrix(nrow = 3, ncol = 3))
+                colnames(calc) <- c("time", "cort", "lcort")
+                calc[, 1] <- t(rw_dat[i, c("tube1_tsec", "tube2_tsec", "tube3_tsec")])
+                calc[, 2] <- t(rw_dat[i, c("tube1_cort", "tube2_cort", "tube3_cort")])
+                calc[, 3] <- log(calc[, 2] + 1)
+                rownames(calc) <- c(1, 2, 3)
+                calc <- subset(calc, calc$time < 600 & is.na(calc$cort) == FALSE)
+              # Calculate the slope over the first 8 minutes from a simple linear regression. Both normal and log scale.   
+                calc$time2 <- calc$time / 60
+                ifelse(nrow(calc) > 1, slope <- coef(lm(calc$cort ~ calc$time2))[2], slope <- NA)
+                rw_dat$slope[i] <- slope
[…]

### 2021-01-08 3_r_scripts/1_main_script.R

@@ -1,290 +1,0 @@
-### Script for speed vs. scope manuscript
-# Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
-
-
-## Load packages ----
-    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
-
-## Load RWBL field data ----
-        rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
-        rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
-    
-    ## Replacing labeling on a couple rows to make them conform to level names
-        rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
-        rw_dat$LH.stage[38]<-"early-breeding"
-    
-    ## Make a list of colors in order of rows to plot LH stages in different colors
-        collist<-rep(NA,nrow(rw_dat))
-        for(i in 1:nrow(rw_dat)){
-          if(rw_dat$LH.stage[i]=="early-breeding"){collist[i]<-"lightblue"}
-          if(rw_dat$LH.stage[i]=="late-breeding"){collist[i]<-"gray60"}
-          if(rw_dat$LH.stage[i]=="molt"){collist[i]<-"red"}
-        }
-
-## Analysis of RWBL field data ----
-        ## Calculate slope of increase on log scale in first 8 minutes for each sample
-            for(i in 1:nrow(rw_dat)){
-              calc <- as.data.frame(matrix(nrow = 3, ncol = 3))
-              colnames(calc) <- c("time", "cort", "lcort")
-              calc[, 1] <- t(rw_dat[i, c("tube1_tsec", "tube2_tsec", "tube3_tsec")])
-              calc[, 2] <- t(rw_dat[i, c("tube1_cort", "tube2_cort", "tube3_cort")])
-              calc[, 3] <- log(calc[, 2] + 1)
-              rownames(calc) <- c(1, 2, 3)
-              calc <- subset(calc, calc$time < 600 & is.na(calc$cort) == FALSE)
-              calc$time2 <- calc$time / 60
-              ifelse(nrow(calc) > 1, slope <- coef(lm(calc$cort ~ calc$time2))[2], slope <- NA)
-              rw_dat$slope[i] <- slope
[…]

### 2021-01-10 3_r_scripts/1_markdown_document.Rmd
Updating markdown.

@@
 
+## Time to Max vs. Max
+
+Here I'm plotting the max cort value reached vs. the time that individuals reached that value. Do individuals with the fastest cort responses also reach higher levels of circulating cort?
+
+![](time95_vs_max.png)\
+
+**Figure x.**
+
+![](time95_vs_maxl.png)\
+
+**Figure x.**
+
 ## RWBL Conclusions
@@
 It may be better to look at the time to reach 70%, since that seems to be the part of the curve where individuals and stages differ most from each other.
++ The time taken to reach maximal cort is related to the absolute value of max cort, but the relationship is positive.
+That means that birds that take longer to reach their max level also reach higher levels, presumably because they continue increasing for longer.
+That is really a problem for disentangling measure of speed and scope with time point measures.
 + Measuring only at 30 minutes misses a lot of these features of the cort response.
@@
 
+This is a demonstration of the output and plotting function for the simulation using all the default settings. I've been playing with what makes sense as defaults, but currently this is set such that:
+
++ Base cort is
+
+![](demo.png)\
+
+In the default plot the three panels show  
+1) The simulated dataset that is reduced to the defined sampling time points (2, 15, 30).  
+2) The full time series dataset. This is fit with a loess curve from the 'true' values sampled with error. The 'wiggliness' can be adjusted.  
+3) The rank profile of each animal over the time series.
+
 ## Speed vs. Scope

### 2021-01-10 3_r_scripts/1_RWBL_script.R

@@ -3,7 +3,7 @@
 
 
 ## Load packages ----
-    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
+    pacman::p_load(ggplot2, here, tidyverse, viridis, gridExtra)
 
 ## Load RWBL field data ----
         rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
@@ -320,216 +320,25 @@
           guides(fill = FALSE, color = FALSE)
         ggsave(here::here("3_r_scripts/slope_vs_95_log.png"),
                p2, device = "png", width = 8.4, height = 3.1, units = "in")
+        
+## Time95 vs. max RWBL ----
+        p1 <- ggplot(data = rw_dat, mapping = aes(x = time95 / 60, y = max.loess, color = LH.stage, fill = LH.stage)) +
+          geom_point() + geom_smooth(method = "lm") +
+          scale_fill_viridis(discrete = TRUE) + scale_color_viridis(discrete = TRUE) +
+          theme_classic() + xlab("Time to 95% of max") + ylab("Max cort (ng/ul)") +
+          facet_wrap(~ LH.stage, scale = "free") +
+          guides(fill = FALSE, color = FALSE)
+        ggsave(here::here("3_r_scripts/time95_vs_max.png"),
+               p1, device = "png", width = 8.4, height = 3.1, units = "in")
+        
+        p2 <- ggplot(data = rw_dat, mapping = aes(x = time95 / 60, y = max.l, color = LH.stage, fill = LH.stage)) +
+          geom_point() + geom_smooth(method = "lm") +
+          scale_fill_viridis(discrete = TRUE) + scale_color_viridis(discrete = TRUE) +
+          theme_classic() + xlab("Time reaching 95% of maximal") + ylab("log max cort (ng/ul)") +
+          facet_wrap(~ LH.stage, scale = "free") +
+          guides(fill = FALSE, color = FALSE)
+        ggsave(here::here("3_r_scripts/time95_vs_maxl.png"),
+               p2, device = "png", width = 8.4, height = 3.1, units = "in")
                 
-## Create simulation function ----
-  # The first function simulates cort response parameters for a population of animals of size n. We
-    # can think of these as the 'true' phenotypes of these individuals. For purposes of this simulation
-    # the cort response is described with five simulated parameters that can covary or not.
-    # The parameters estimated are: 
-          #base cort
-          #time to reach max cort
-          #max cort
[…]

### 2021-01-10 3_r_scripts/2_cortResponse_script.R

@@ -5,322 +5,6 @@
 ## Load packages ----
     pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
 
-## Load RWBL field data ----
-        rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
-        rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
-    
-    ## Replacing labeling on a couple rows to make them conform to level names
-        rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
-        rw_dat$LH.stage[38]<-"early-breeding"
-    
-    ## Make a list of colors in order of rows to plot LH stages in different colors
-        collist<-rep(NA,nrow(rw_dat))
-        for(i in 1:nrow(rw_dat)){
-          if(rw_dat$LH.stage[i]=="early-breeding"){collist[i]<-"lightblue"}
-          if(rw_dat$LH.stage[i]=="late-breeding"){collist[i]<-"gray60"}
-          if(rw_dat$LH.stage[i]=="molt"){collist[i]<-"red"}
-        }
-
-## Analysis of RWBL field data ----
-        ## Loop through data frame to calculate slope and max cort achieved for each bird's samples
-            for(i in 1:nrow(rw_dat)){
-              # Set up dataframe for each bird with just the first three samples (less than 10 minutes)
-                calc <- as.data.frame(matrix(nrow = 3, ncol = 3))
-                colnames(calc) <- c("time", "cort", "lcort")
-                calc[, 1] <- t(rw_dat[i, c("tube1_tsec", "tube2_tsec", "tube3_tsec")])
-                calc[, 2] <- t(rw_dat[i, c("tube1_cort", "tube2_cort", "tube3_cort")])
-                calc[, 3] <- log(calc[, 2] + 1)
-                rownames(calc) <- c(1, 2, 3)
-                calc <- subset(calc, calc$time < 600 & is.na(calc$cort) == FALSE)
-              # Calculate the slope over the first 8 minutes from a simple linear regression. Both normal and log scale.   
-                calc$time2 <- calc$time / 60
-                ifelse(nrow(calc) > 1, slope <- coef(lm(calc$cort ~ calc$time2))[2], slope <- NA)
-                rw_dat$slope[i] <- slope
[…]

### 2021-01-10 3_r_scripts/1_markdown_document.Rmd
Changed method of curve fitting in simulation.

@@
 
-## Speed vs. Scope
+Below I'm mostly cropping plots at 60 minutes, but the time series goes out all the way to where the birds return to baseline, which is also sampled as part of the simulation. Like this:
+
+![](demo_fullx.png)\
+
+## Relative variation in speed or scope
+
+Measuring cort at 30 minutes at interpreting differences as variation in max cort implicitly assumes that variation in maxcort exists while variation in speed does not (or at least that it is so much smaller than max variation that it can be ignored).
+These three panels illustrate scenarios where cort is measured at 2, 15, and 30 minutes under different scenarios with respect to the amount of variation in speed or scope.
+In all three cases here, I assume that the population average maxcort is the same and that the average time to reach max cort is exactly 30 minutes (that means that we have timed our 30 minute sample perfectly for this population).
+Responses are measured with error, so even in cases where there is no 'true' variation, some between individual differences are visible.
+The correlation between all separate components of the curves are set to 0 here for simplicity.
+There are three scenarios:
+
+A) Max cort varies but there is no variation in speed.  
+B) Both max cort and speed vary in reasonably comparable amounts.  
+C) Max cort does not vary but speed does.
+
+![](relative_variation.png)\
+
+Even though the sampling is the same in all three of these cases, the information encoded in the variation in cort levels at 15 and 30 minutes is very different.
+
+**Panel A** Both the 15 and 30 minute samples accurately reflect variation in max cort.
+Note though that the 15 minute measure actually does a better job of capturing the rank order of max cort (though not the absolute measure) because it is taken before any individuals have started to decline.
+At 30 minutes, the sample is still a reasonably good measure of max cort, but some individuals have started to decline already while others are going up.
+The estimate of average max cort on an absolute scale is better, but the individual ranks are at least a bit out of order.
+Note that this is driven by small differences in the amount of time individuals sustain their highest level and how long they take to return to baseline even in the absence of any underlying variation in the 'true' value of speed to reach max cort.
+
+**Panel B** There is variation in both speed and max cort.
+Because speed and max cort are not correlated here, the measures at both 15 and 30 minutes are completely confounded between the two aspects and represent some combination of speed and scope that is likely impossible to disentangle without a more complete time series for each individual.
+
+**Panel C** There is no variation in max cort, but ther is variation in the speed to reach maximum.
+At 15 minutes, there appears to be substantial variation in absolute cort level, but this is completely attributable to variation in the speed rather than maximum cort.
+At 30 minutes, even though we have timed our sample exactly at the population mean, our measure is not telling us much of anything because the variation is all in the speed of response and we are sampling a mix of individuals on the way up, at their plateau, and on the way down.
+In this case it would be better to have samples earlier than the mean population peak value if we were interested in understanding variation in speed.
+
+## Correlation between speed and scope
+
+Now consider the case of how much of a correlation there is between the speed and scope of the response.
+
+![](sp_sc_correlation.png)\
+
+**Panel A** Here speed and scope have a very strong positive correlation where faster responders also reach higher maximum values.
+This is good in some ways because it means that even though we can't tell speed and scope apart, by measuring one we are accurately measuring both (and maybe in this case you can just think of 'strong' responders as being both fast and high).
+This is a nice world, but we don't really have data to tell whether this is the world we are in.
+
+**Panel B** Here this is like the one above where there is no correlation.
+Speed and scope are confounded and a high value at 15/30 minutes could either be because of a fast responder or a high maximum but you can't really tell which.
+
+**Panel C** Here there is a strong correlation but in the other direction such that fast responders actually end up leveling off at lower maximal cort values.
+I think most people might assume this usually isn't the case, but in the RWBL data above, this is actually closer to what we see (though this is a stronger correlation just to illustrate).
+In this case, measures at 15/30 minutes that are interpretted as max cort values could be really misleading.

### 2021-01-10 3_r_scripts/2_cortResponse_script.R

@@ -19,7 +19,7 @@
                                base_mu = 5, base_sd = 1.5, base_min = 0.5,
                                speed_mu = 30, speed_sd = 5, speed_min = 10,
                                max_mu = log(50), max_sd = log(1.2), 
-                               maxtime_mu = 10, maxtime_sd = 3, maxtime_min = 4,
+                               maxtime_mu = 15, maxtime_sd = 3, maxtime_min = 5,
                                return_mu = 90, return_sd = 15, 
                                cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0,
                                cor_speed_max = 0, cor_speed_maxtime = 0, cor_speed_return = 0,
@@ -97,7 +97,7 @@
                                 sample_times = 1,
                                 timecourse_max = 170,
                                 performance_contributions = c(0, 0, 0, 0, 0, 1),
-                                sm_span = .6) #relative contribution of base, speed, max, maxtime, return, and error to performance
+                                sm_span = .2) #relative contribution of base, speed, max, maxtime, return, and error to performance
               {
             data2 <- data[rep(seq_len(nrow(data)), sample_times), ]
             data2$sample <- rep(seq(1, sample_times, 1), each = nrow(data))
@@ -143,18 +143,62 @@
                         mean = 0,
                         sd = return_error * data2[which(data2$group == "return"), "x"])
              
-            
-            time_seq <- data.frame(time = seq(0, timecourse_max, 1))
-            preds <- as.data.frame(matrix(nrow = length(unique(data2$animal_sample)), ncol = nrow(time_seq)))
-            colnames(preds) <- time_seq$time
-            rownames(preds) <- unique(data2$animal_sample)
-            for(i in 1:length(unique(data2$animal_sample))){
-              temp <- subset(data2, data2$animal_sample == unique(data2$animal_sample)[i])
-              suppressWarnings(
-                fits <- predict(loess(temp$y ~ temp$x, span = sm_span), newdata = time_seq$time)
-              )
-              preds[i, ] <- fits
-            }
[…]

### 2021-01-11 3_r_scripts/1_markdown_document.Rmd
Made a bunch of changes to the simulation.

@@
 Some samples are also missing from the end or middle of some time series.
+
+Note that I have removed one bird from early season that has an astronomically high cort value even starting at baseline and through the full series.
+It doesn't really make any qualitative difference, but including that bird just makes it hard to plot because it is so much higher than any others.
 
@@
 First, one function generates a population with 'true' values for base, speed, max, time at plateau, and time at return to base for each along with a 'performance' metric (i.e., fitness) that is derived from those true underlying values.
+I have now also added two additional parameters for the initial slope (a period of time with a potentially steeper slope in the first part of the response) and a 'fastpct' that defines what percentage of the total stress response (max - base) occurs during that steep initial slope period.
 You can set the mean and sd for each parameter, the number of animals to generate, and the degree of correlation between each parameter.
 Animals are then drawn from a multivariate normal distribution with the set values.
-Note that the function will allow you to set any values, but a lot of combinations won't work if the produce negative values or if different traits are set to have strong correlations in a way that makes it impossible to generate a multivariate normal distribution; a message is given in these cases to try different values.
+Note that the function will allow you to set any values, but some combinations won't work if the produce negative values or if different traits are set to have strong correlations in a way that makes it impossible to generate a multivariate normal distribution; a message is given in these cases to try different values.
 I think I could extend this first function to also have slopes across a gradient for some parameters (e.g., the 'true' slope of response in base cort across a temperature gradient).
@@
 Second, the output from the first function is used to simulate corticosterone time courses for each animal (or multiple times for each animal).
-In this function you can set the degree of error from the 'true' value for each parameter.
-This can be thought of as measurement error or as individual variation associated with repeat sampling of the same animal.
+In this function you can set the degree of noise from the 'true' value for each parameter.
+This should be thought of as within-individual variation in cort response expression.
+You can also set an 'assay error'.
+This should be thought of as measurement error associated with the assay or pipetting, etc and it is only applied to the simulated dataset that is downsampled from the full timecourse to be measured at the set 'bleed times'.
 After this sampling, a loess smoothed curve is fit for each series of parameters for each animal and then 'cort' values are recorded at each minute of the full time course.
@@
 
-This is a demonstration of the output and plotting function for the simulation using all the default settings. I've been playing with what makes sense as defaults, but currently this is set such that:
-
-+ Base cort is
+This is a demonstration of the output and plotting function for the simulation using all the default settings.
+I've been playing with what makes sense as defaults, but probably any individual study would want to change them for species specific values.
 
@@
 In this case, measures at 15/30 minutes that are interpretted as max cort values could be really misleading.
+
+# Simulate RWBL Data
+
+So one of the major challenges to disentangling speed and scope is just getting a better handle on how to parameterize these shapes for different species.
+How much variation is there in when individuals reach their max?
+How does that compare to variation in the max value itself?
+How do the relative contributions of those different parameters influence our ability to detect causal relationships between underlying traits and fitness when we only have measurements taken with error?
+
+There are a lot of ways we could play with this simulator to address those questions, but they involve a lot of guesswork.
+I think one promising approach that makes sense here is to start with the empirical data from the RWBL dataset that has a (relatively) fine scale time course measured and create a simulation that produces similar curves.
+Then we can ask how easily we can identify relationships between cort phenotype and fitness from the simulated data.
+
+Here is an attempt at that where I have tweaked the simulation parameters until I produce curves that look qualitatively like the RWBL data.
+I've collected the data with the same 7 point time series as we used in the empirical study, but we also have access to the full time course and the true values underlying it.
+For simplicity, I'm only using the early season records from RWBL.
+
+Like in the empirical data above, I've encoded a positive correlation between the initial slope and the max cort value achieved. All other traits are left uncorrelated.
+
+![](rwbl_sim.png)\
+
+**Figure x.** Simulated data plotted against red-winged blackbird data from the early breeding season.
+Simulated data are sampled at 7 time points corresponding to those used in the empirical study and plotted by animal (panel A) and with a smoothed curve and confidence interval for the simulated versus empirical data (panel B).
+
+So this does a good job of simulating data that looks like the RWBL data. Next lets use that as a starting point to explore whether we can recover relationships between speed, scope and performance.
+
+## Scenario 1: Speed predicts performance
+
+In this first scenario, I'm simulating the data with the same parameters as above.
+We'll take a population of 50 animals, with cort measured at 30 and 55 minutes (since 55 is the average max time here).
+Speed of the cort response explains 50% of the variation in performance, scope explains 0%, and the other 50% is random.

### 2021-01-11 3_r_scripts/1_RWBL_script.R

@@ -8,6 +8,7 @@
 ## Load RWBL field data ----
         rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
         rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
+        rw_dat <- subset(rw_dat, rw_dat$band != "1372-33452")
     
     ## Replacing labeling on a couple rows to make them conform to level names
         rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
@@ -46,6 +47,23 @@
                 values.l <- gsub("NA", 0, values.l)
                 rw_dat$max.t[i] <- which.max(values.l)
             } 
+        
+        # Calculate max from loess regression
+          for(i in 1:nrow(rw_dat)){	
+            yy <- as.vector(t(rw_dat[i, 38:45]))	
+            xx <- as.vector(t(rw_dat[i, 30:37]))
+            dd <- as.data.frame(cbind(yy, xx))
+            colnames(dd) <- c("yy", "xx")
+            ss <- loess(yy ~ xx, span = 1.5) 
+            max <- max(predict(ss, seq(min(na.omit(dd$xx)), max(na.omit(dd$xx)))))
+            pct <- predict(ss, seq(min(na.omit(dd$xx)), max(na.omit(dd$xx)))) / max
+            se <- seq(min(na.omit(dd$xx)), max(na.omit(dd$xx)))
+            rw_dat$thirtypct[i] <- pct[which(1800 == se)]
+            rw_dat$time95[i] <- se[min(which(pct > 0.95))]
+            rw_dat$time.max[i] <- se[which.max(se)]
+            pred <- predict(ss,seq(min(na.omit(dd$xx)), max(na.omit(dd$xx))))
+            rw_dat$max.loess[i] <- max(pred)
+          }
         
         ### Make a long version of the data for easier plotting
               rwd_long1 <- as.data.frame(pivot_longer(rw_dat, cols = tube1_tsec:tube8_cort,

### 2021-01-11 3_r_scripts/2_cortResponse_script.R

@@ -17,62 +17,89 @@
           #time of return to base cort level 
           cort_sim1 <- function(n = 20, 
                                base_mu = 5, base_sd = 1.5, base_min = 0.5,
-                               speed_mu = 30, speed_sd = 5, speed_min = 10,
-                               max_mu = log(50), max_sd = log(1.2), 
-                               maxtime_mu = 15, maxtime_sd = 3, maxtime_min = 5,
+                               slope_mu = 15, slope_sd = 1,
+                               fastpct_mu = 0.65, fastpct_sd = 0.05, 
+                               speed_mu = 25, speed_sd = 5, speed_min = 10,
+                               max_mu = log(50), max_sd = 0.21, 
+                               maxtime_mu = 10, maxtime_sd = 3, maxtime_min = 5,
                                return_mu = 90, return_sd = 15, 
-                               cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0,
-                               cor_speed_max = 0, cor_speed_maxtime = 0, cor_speed_return = 0,
-                               cor_max_maxtime = 0, cor_max_return = 0,
-                               cor_maxtime_return = 0)
+                               cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0, cor_base_slope = 0, cor_base_fastpct = 0,
+                               cor_speed_max = 0, cor_speed_maxtime = 0, cor_speed_return = 0, cor_speed_slope = 0, cor_speed_fastpct = 0,
+                               cor_max_maxtime = 0, cor_max_return = 0, cor_max_slope = 0, cor_max_fastpct = 0,
+                               cor_maxtime_return = 0, cor_maxtime_slope = 0, cor_maxtime_fastpct = 0,
+                               cor_return_slope = 0, cor_return_fastpct = 0,
+                               cor_slope_fastpct = 0)
               {
-                  # make variance covariance matrix
-                      cort_vcov <- matrix(nrow = 5, ncol = 5)
+                  # Build a full variance-covariance matrix based on the values for each parameter and correlation between parameters
+                      cort_vcov <- matrix(nrow = 7, ncol = 7)
                       cort_vcov[1, ] <- c(base_sd^2, cor_base_speed * base_sd * speed_sd, cor_base_max * base_sd * max_sd, 
-                                          cor_base_maxtime * base_sd * maxtime_sd, cor_base_return * base_sd * return_sd)
[…]

### 2021-01-27 3_r_scripts/1_markdown_document.Rmd
Made a bunch of changes to the algorithm.

@@
 
+This is a concept figure of what the simulation is doing:
+
+![](concept.png)\
+
 ## Demonstration
@@
 
+## Variation in speed but not max
+
+Here is another illustration showing what happens if there is no variation in maximal cort coupled with different amounts of variation in speed (A -> C = adding more variation).
+In some cases (panel A), apparent variation in max at 30 minutes is really just measuring speed (but at least measuring it well).
+In others (Panel C), variation in 'max' at 30 minutes just isn't measuring anything, but still looks like there is individual variation.
+
+![](var_in_speed_only.png)\
+
+# Extensions
+
+Leaving notes here for myself. I think that I can pretty easily add:
+
+- Calculate and output area under the curve as a default based on whatever sampling points are chosen.
+This wouldn't be an additional parameter, but would change as a consequence of the parameters set above.
+
+- Make an extra function that will calculate profile repeatability (Reed & Romero 2019) from a simulated dataset with >1 sample drawn per individual.
+I *think* I can do this, but looking back quickly at how they actually calculated this, it is kind of a mess that involves counting the number of crossovers between time points.
+I might still be able to code it that way though, will have to look in more detail.
+
+- Given a set of parameters, could I make the function tell you the optimal times to collect x number of samples if you want to separately estimate speed and scope?
+Presumably you would want a couple early to get at rapid increase.
+But if there is very little variation in baseline cort then it might be better not to 'waste' a baseline measure?
+To really get time to max you kind of need a few samples pre/at/post max right?
+
 # Simulate RWBL Data
@@
 Speed of the cort response explains 50% of the variation in performance, scope explains 0%, and the other 50% is random.
+
+- Scenarios to point out:
+
+If there is any variation in speed, observed repeatability in basecort can be entirely attributable to speed even if there is 0 variation in basecort.
+This situation may be better if there really is a lag between capture and increase, but how do you assess if there is individual variation in the time of that lag?

### 2021-01-27 3_r_scripts/2_cortResponse_script.R

@@ -3,7 +3,7 @@
 
 
 ## Load packages ----
-    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra)
+    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra, sp, rptR)
 
 ## Create simulation function ----
   # The first function simulates cort response parameters for a population of animals of size n. We
@@ -16,11 +16,11 @@
           #time at max cort level
           #time of return to base cort level 
           cort_sim1 <- function(n = 20, 
-                               base_mu = 5, base_sd = 1.5, base_min = 0.5,
-                               slope_mu = 15, slope_sd = 1,
-                               fastpct_mu = 0.65, fastpct_sd = 0.05, 
+                               base_mu = 5, base_sd = 2, base_min = 0.5,
+                               slope_mu = 15, slope_sd = 2,
+                               fastpct_mu = 0.55, fastpct_sd = 0.08, 
                                speed_mu = 25, speed_sd = 5, speed_min = 10,
-                               max_mu = log(50), max_sd = 0.21, 
+                               max_mu = log(45), max_sd = 0.21, 
                                maxtime_mu = 10, maxtime_sd = 3, maxtime_min = 5,
                                return_mu = 90, return_sd = 15, 
                                cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0, cor_base_slope = 0, cor_base_fastpct = 0,
@@ -65,7 +65,7 @@
                           animal = paste("id", 1:n, sep = "_"),
                           base = msam[, 1],
                           tmax = msam[, 2],
-                          max = exp(msam[, 3]),
+                          max = exp(msam[, 3]) + msam[, 1],
                           atmax = msam[, 4],
                           return = msam[, 5],
                           slope = msam[, 6],
@@ -94,7 +94,7 @@
                                 sim_dat$atmax + sim_dat$tmax,
                                 sim_dat$return),
                           y = c(sim_dat$base,
-                                (sim_dat$max - sim_dat$base) * sim_dat$fastpct,
+                                (sim_dat$max - sim_dat$base) * sim_dat$fastpct + sim_dat$base,
                                 sim_dat$max,
                                 sim_dat$max,
                                 sim_dat$base),
@@ -117,20 +117,20 @@
         # a simulated dataset with samples at specified time points, the full time course with cort at every time point, and a 
         # rank converted estimate. 
[…]

### 2021-04-06 3_r_scripts/Manuscript.Rmd
Made lots of changes to start writing a manuscript.

---
title: "Functions for simulating data and designing studies of physiological flexibility in the acute glucocorticoid response to stressors"
author: "Conor Taff"
date: "Last Updated: `r format(Sys.time(), '%B %d %Y at %H:%M')`"
output: pdf_document
bibliography: references.bib
csl: animal-behaviour.csl
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ABSTRACT

Wild animals often experience unpredictable challenges that demand rapid and flexible responses. The glucocorticoid mediated stress response is one of the major systems that allows vertebrates to rapidly adjust their physiology and behavior. Given it's role in responding to challenges, evolutionary physiologists have focused on the consequences of between-individual and, more recently, within-individual variation in the acute glucocorticoid response. Although sophisticated approaches are available to partition this variation statistically, empirical studies of physiological flexibility are severely limited by the logistical challenges of measuring the same animal multiple times during a single acute response or across multiple instances of acute responses. Empiricists have largely adopted the strategy of standardizing sampling as much as possible to allow for comparison between individuals, but this standardization also makes it very difficult to detect certain types of variation in the functional shape of acute response curves. Data simulation is a powerful approach when empirical data are limited, but has not been adopted to date in studies of physiological flexibility. In this paper, I describe a set of simulation functions that can generate realistic acute glucorticoid response data with user specified characteristics. Simulated animals can be sampled continuously through an acute response and across as many separate responses as desired, while varying key parameters (e.g., the degree of correlation between the speed and scope of a response). Using this simulation, I explore several possible scenarios to highlight areas where simulation might either provide new insight into physiological flexibility directly or aid in designing empirical studies that are better able to test the hypotheses of interest. Complete code and a worked vignette using the functions are provided so that the simulation can be adapted to any study system.

## INTRODUCTION

Animals live in a dynamic environment in which they regularly encounter unpredictable challenges. Successfully navigating these challenges often requires the ability to rapidly adjust behavior and physiology to match current conditions. For vertebrates, the glucocorticoid mediated stress response plays a major role in coordinating these changes when stressors are encountered [@wingfield1998; @sapolsky2000] and similar rapid response systems mediate changes in other taxa [@taborsky2020]. Because of the central role that this response plays in coping with challenges, a great deal of research effort over the past 15 years has focused on understanding whether between-individual differences in the magnitude of this response predict coping ability and, ultimately, fitness [@breuner2008; @schoenle2020].

More recently, a series of conceptual papers have asked whether the degree of within-individual variation in glucocorticoid modulation (i.e., endocrine flexibility) across different contexts or in response to different stressors might also be an important predictor of performance [@taff2016; @hau2016; @wada2014; @lema2013]. Perhaps the major limit to empirical progress, especially for within-individual variation, is the logistical difficulty of accurately characterizing the functional shape of the acute physiological stress response for an individual during a single acute response and across multiple acute responses occurring under different conditions. Often, these measures are strictly limited by the number of samples that can safely be taken from an animal during a single capture and the number of repeated captures that are possible [but see @koolhaas2011]. Given these limitations, data simulation is a powerful tool that could complement empirical work in this area, but that has not yet been applied to studies of endocrine flexibility.

Several recent papers have suggested that physiologists interested in endocrine flexibility should adopt a within-individual reaction norm approach [e.g., @hau2016; @taff2016]. This approach has been widely adopted in studies of behavioral flexibility where statistical approaches and empirical progress have developed synergistically [e.g., @araya2015; @dingemanse2010; @westneat2015]. This field has also benefited from simulation studies to evaluate optimal study design [@van2012] and packages that can create artificial datasets with desired patterns of between, within, and residual variance to evaluate the consequences of different patterns of variation on the ability to detect effects [see SQuID package, @allegue2017]. While these approaches are powerful, they have proven difficult to apply directly to endocrine flexibility data for two reasons. First, simulation studies suggest that many patterns may only be detectable with a level of repeated sampling that is possible for many behaviors (especially when collected autonomously), but that is currently not possible for most studies of endocrine flexibility. Second, and more fundamentally, these papers often focus on somewhat discrete measures of behavior (e.g., aggression score or activity level), whereas for acute glucocorticoid responses, the functional shape of the response itself may be the important trait and it may not be possible to summarize variation in the shape of the response with a single measure.

The function valued trait (FVT) framework is an alternative approach that explicitly considers the functional shape of a biological response [@gomulkiewicz2018; @stinchcombe2012; @kingsolver2015]. While FVT approaches have been suggested for studies of endocrine flexibility [@taff2016], I am not aware of any papers that have applied this framework to empirical data on acute glucocorticoid responses, probably because sufficient data are not available. Conceptually, however, this approach is a better match to the acute glucocorticoid response, because the shape of a response curve is explicitly considered as the phenotypic trait of interest. In some cases, it may make sense to estimate particular parameters of the curve (e.g., maximum rate of increase and maximum value reached) and then treat those parameters as phenotypic values for downstream analysis, although statistical methods also exist to analyze the shape of the entire curve directly without the need to extract discrete parameters [@kingsolver2015]. This approach has been used to study a variety of phenotypes where values can be measured continuously or pooled across many individuals from the same group to accurately estimate the shape of a curve [see Table 1 in @stinchcombe2012]. Applying the technique to endocrine flexibility at the within-individual level faces the same empirical challenges described for within-individual reaction norms above. Note that FVT and within-individual reaction norms approaches are not necessarily incompatible, but they have largely developed separately.

The recognition that characterizing the functional shape of an acute stress response is challenging goes back to the earliest studies conducted in wild animals. Early studies often employed various control groups and sampled individual animals at a variety of time points over a long period in order to describe the full response curve for a particular group [e.g., a species or a breeding stage, @wingfield1992]. These validations were considered essential to characterize key parameters of the acute response for each group being studied (i.e., baseline, rate of increase, maximum level, time of peak, and area under the curve; John Wingfield, personal communication). The challenge of estimating these parameters becomes much more difficult when trying to describe the response for an individual animal rather than for a group, because glucocorticoids can often only be measured at two or three time points and only a small number of times per animal [e.g., @vitousek2018]. Because these studies require an estimate for each individual, the solutions used by older studies that added additional animals to allow for sampling at more time points are not available.

For individual based studies, the most common approach to this problem is to standardize measurements as much as possible by measuring animals at the same time of the day during the same context, and by taking blood samples at standard times (often <3 and 30 minutes after capture) to characterize baseline and stress-induced glucocorticoids. This standardization allows for comparison between individuals, but in some cases it may also completely obscure the ability to detect variation in certain characteristics of the acute response curve. For example, if the speed and scope (maximum value) of the acute response vary independently, samples taken at only two time points cannot accurately capture variation in either parameter. Indeed, several discussions in recent years about methods such as the '3 minute rule' and the relative merits of 'area under the curve' versus time point measures of glucocorticoids are fundamentally related to a recognition of the importance of understanding variation in the functional shape of stress responses and whether different components of that shape covary within individuals [e.g., @cockrem2002; @small2017].

One of the characteristics of both the within-individual reaction norm and FVT literature is that empirical work has proceeded in very close coordination with simulation and statistical method development. In contrast, studies of endocrine flexibility often point to these methods, but don't address the ways that the particular logistical challenges of hormone measurement might suggest altered study designs. While many of the tools developed in these fields are transferable, studies of physiological flexibility would benefit from a focus on analysis development and testing that explicitly incorporates the particular details of these questions. One way to accomplish these goals is to use simulations, but to my knowledge no studies of physiological flexibility have developed simulations of the acute stress response that address the issues discussed above.

Data simulation is a powerful approach for several reasons. Because true parameter values (e.g., maximum glucocorticoid level) are known, it is possible to evaluate how well different study designs and analytical choices perform in recovering true patterns and how sensitive those designs are to different assumptions. Thus, simulation can tell us whether the study designs we use can *in principle* detect the patterns we predict given realistic effect sizes. Simulated data can also identify conditions under which current study designs will perform well or poorly. For example, if simulations suggest that the baseline paired with stress-induced paradigm only works well when the speed and scope of responses are positively correlated, then empirical work could seek to determine the degree of correlation for a particular study system as justification for the approach. This ability to highlight key assumptions and create data sets with known properties has the ability to both provide insight into physiological flexibility directly and to guide empirical work by improving study design and identifying key areas for subsequent sampling. In the rest of this paper, I develop a simple simulation of acute physiological stress responses and then briefly illustrate several possible applications of the simulation.

## DESCRIPTION OF THE SIMULATION

I developed a set of simulation functions in R version 4.0.2 [@rcore] to generate acute physiological response curves. This simulation makes no assumptions about the  mechanistic process that results in the shape of a glucocorticoid response. Rather, parameters are sampled to generate curves that are similar in shape and degree of variation to empirically observed responses (Figure 1). This simulation is designed to create data sets with realistic structure that can be used to better design and plan studies of physiological flexibility, to evaluate power of current study designs, and to evaluate the sensitivity of sampling regimes to any number of modifications to the shape of glucocorticoid response curves (e.g., changing covariation patterns between different features of the response). I explore a small number of scenarios in the next section, but I expect that many other scenarios can be addressed with these tools. For illustration purposes, I refer to simulated glucocorticoid responses, but the simulation applies equally well to any physiological mediator of a rapid response.

![](concept.png)\

***Figure 1.** Conceptual illustration of the structure of the simulation. For each simulated animal, seven parameters are sampled from a multivariate normal distribution. Together, these seven parameters define the turning points in an acute response curve. The mean and standard deviation for each parameter can be set along with the degree of covariation between each pair of parameters. Note that the simulation can easily be simplified as desired by setting some parameter mean or standard deviations to zero.*

The simulation is constructed as two main functions with several minor functions for downstream analysis. I have included a detailed description of the functions and arguments in the appendix along with a full vignette detailing the process of specifying values for each function (https://github.com/cct663/speed_vs_scope). Briefly, function `cort_sim1` samples the parameters shown in Figure 1 from an arbitrary number of animals. These parameters are sampled from a multivariate normal distribution with user specified mean, variance, and covariance for each parameter. I consider these values to be the 'true', unobserved, phenotype of the animal (setting aside the question of whether or not a 'true' physiological phenotype exists).

A second function, `cort_sim2`, starts with a population of animals generated from `cort_sim1` and samples observed acute glucocorticoid responses an arbitrary number of times for each animal. Two sources of variation in the observed relative to true parameter values can be specified. First, within-individual variation in expression is represented by specifying what amount of variation in the observation of each parameter is determined by the true value and what amount is determined by an additional randomly sampled response, based on the population parameters (this additional sampling maintains the user specified covariance structure of the population). Second, additional noise can be added to all measurements to represent measurement error (e.g., assay error, inaccurate pipetting, etc). After sampling the parameters, values are interpolated for each one minute time point and a localized regression is fit to create a smoothed curve that represents the observed glucocorticoid response.

The function also generates a simulated performance (e.g., fitness) measure, based on the underlying true values. In addition to returning the full curve, a down sampled dataset is returned that retains measurements made only at the user specified fixed time points (e.g., only at 2 and 30 minutes, as is typical in many field based studies). Data generated from this function can then be used in downstream analyses with any standard statistical approaches or software. For example, a user could perform an analysis to ask whether a known relationship between fitness and a particular true parameter is recovered in an analysis that includes only measures taken at particular time points. An additional convenience function summarizes the output of a simulation run in a multi-panel plot (Figure 2).

![](demo.png)\

***Figure 2.** Example of simulation output with default settings. Panel A shows the down sampled data set for this run with samples collected at 1, 15, and 30 minutes in this case. Panel B shows the full observed response curve for each animal. Panel C shows the rank order of glucocorticoid level at each time point for each animal. In each panel, the vertical dashed lines represent the three time points that might have been measured in a typical empirical study.*

Finally, given recent interest in estimating the repeatability of glucocorticoid regulation [@taff2018; @cockrem2013; @hau2016], I also included a function that takes input from `cort_sim2` and calculates the observed repeatability of several measures using package `rptR` [@stoffel2017]. Full details are included in the appendix, but this function returns repeatability for each individual time point specified in the down sampled data set, profile repeatability [@reed2019], and for area under the curve calculated as both increase (AUC~I~) and ground (AUC~G~) approaches [@pruessner2003]. For each AUC measure, I return repeatability for the full time course, for an estimate using only the observed values in the down sampled data set, and for the full data set constrained to the time period encompassing the observed data points.

## EXAMPLE APPLICATIONS OF SIMULATION

The goal of this simulation is to provide a flexible tool that can produce realistic datasets of physiological flexibility for a variety of different systems and scenarios. As such, there are many possible applications and here I briefly highlight a few possibilities. These are by no means exhaustive, and I hope the simulation will be a useful tool to guide empirical work for specific hypotheses and study systems. Within each scenario, I have illustrated how the simulation functions might be used to address the particular question of interest, but I have not fully explored all the possible permutations of parameters systematically, because these will depend to a large extent on the empirical details of the system being studied.

### *Simulating empirically parameterized data*

### *Exploring covariance between response components*

### *Comparing repeatability of different measures*

### *Designing optimal sampling strategies*

### *Detecting links between fitness and responses*

## DISCUSSION

More sophisticated simulations that incorporate mechanistic processes or other molecules or components of stress response system.

Empirical data is so limited and hard to collect, that it makes sense to proceed with simulations to guide and focus empirical work and to determine whether empirical study designs actually have the ability in principle to support hypotheses of interest given reasonable effect sizes.

Based on the brief scenarios developed above:

## ACKNOWLEDGEMENTS

I would like to thank John Wingfield for discussions about the ideas presented in this project and for providing an insightful overview of the history of study design in field based studies of acute stress responses. I also thank Maren Vitousek and all members of the Vitousek Lab for feedback and discussion on early versions of this project.

## REFERENCES

### 2021-04-06 3_r_scripts/2_cortResponse_script.R

@@ -23,6 +23,7 @@
                                max_mu = log(45), max_sd = 0.21, 
                                maxtime_mu = 10, maxtime_sd = 3, maxtime_min = 5,
                                return_mu = 90, return_sd = 15, 
+                               extra_n = 500,
                                cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0, cor_base_slope = 0, cor_base_fastpct = 0,
                                cor_speed_max = 0, cor_speed_maxtime = 0, cor_speed_return = 0, cor_speed_slope = 0, cor_speed_fastpct = 0,
                                cor_max_maxtime = 0, cor_max_return = 0, cor_max_slope = 0, cor_max_fastpct = 0,
@@ -58,11 +59,11 @@
                       mu_list <- c(base_mu, speed_mu, max_mu, maxtime_mu, return_mu, slope_mu, fastpct_mu)
                   
                   # Create a dataset by sampling from a multivariate normal distribution
-                        msam <- mvrnorm(n, mu = mu_list, Sigma = cort_vcov)
+                        msam <- mvrnorm(n + extra_n, mu = mu_list, Sigma = cort_vcov)
                       
                       # Make a dataframe from the matrix above with columns named by paramter  
                         sim_dat <- data.frame(
-                          animal = paste("id", 1:n, sep = "_"),
+                          animal = paste("id", 1:(n + extra_n), sep = "_"),
                           base = msam[, 1],
                           tmax = msam[, 2],
                           max = exp(msam[, 3]) + msam[, 1],
@@ -87,24 +88,48 @@
                   
                   # Convert simulated data to long form x/y dataframe
                         sim_dat2 <- data.frame(
-                          animal = rep(sim_dat$animal, 5),
+                          animal = rep(sim_dat$animal[1:n], 5),
                           x = c(rep(0, n),
-                                sim_dat$slope,
-                                sim_dat$tmax,
-                                sim_dat$atmax + sim_dat$tmax,
-                                sim_dat$return),
-                          y = c(sim_dat$base,
-                                (sim_dat$max - sim_dat$base) * sim_dat$fastpct + sim_dat$base,
-                                sim_dat$max,
-                                sim_dat$max,
-                                sim_dat$base),
+                                sim_dat$slope[1:n],
+                                sim_dat$tmax[1:n],
+                                sim_dat$atmax[1:n] + sim_dat$tmax[1:n],
[…]

### 2021-07-27 3_r_scripts/Manuscript.Rmd
Made updates and switched to SSH.

@@
 title: "Functions for simulating data and designing studies of physiological flexibility in the acute glucocorticoid response to stressors"
-author: "Conor Taff"
-date: "Last Updated: `r format(Sys.time(), '%B %d %Y at %H:%M')`"
-output: pdf_document
+author: 
+ - Conor C. Taff
+header-includes:
+  - \usepackage[font={footnotesize}, labelfont={bf}]{caption}
+fontsize: 12pt
+output: 
+  bookdown::word_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependcies: ["flafter"]
+  bookdown::pdf_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependencies: ["flafter"]
 bibliography: references.bib
@@
 knitr::opts_chunk$set(echo = TRUE)
+pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra, rptR, simcoRt, sp)
 ```
@@
 Data simulation is a powerful approach when empirical data are limited, but has not been adopted to date in studies of physiological flexibility.
-In this paper, I describe a set of simulation functions that can generate realistic acute glucorticoid response data with user specified characteristics.
+In this paper, I describe the `simcoRt` package, which includes functions that can generate realistic acute glucocorticoid response data with user specified characteristics.
 Simulated animals can be sampled continuously through an acute response and across as many separate responses as desired, while varying key parameters (e.g., the degree of correlation between the speed and scope of a response).
 Using this simulation, I explore several possible scenarios to highlight areas where simulation might either provide new insight into physiological flexibility directly or aid in designing empirical studies that are better able to test the hypotheses of interest.
-Complete code and a worked vignette using the functions are provided so that the simulation can be adapted to any study system.
+The package includes a vignette with details on use and code to reproduce all of the examples explored in this paper.
 
@@
 Perhaps the major limit to empirical progress, especially for within-individual variation, is the logistical difficulty of accurately characterizing the functional shape of the acute physiological stress response for an individual during a single acute response and across multiple acute responses occurring under different conditions.
-Often, these measures are strictly limited by the number of samples that can safely be taken from an animal during a single capture and the number of repeated captures that are possible [but see @koolhaas2011].
+Often these measures are strictly limited by the number of samples that can safely be taken from an animal during a single capture and the number of repeated captures that are possible [but see @koolhaas2011].
 Given these limitations, data simulation is a powerful tool that could complement empirical work in this area, but that has not yet been applied to studies of endocrine flexibility.
@@
 Several recent papers have suggested that physiologists interested in endocrine flexibility should adopt a within-individual reaction norm approach [e.g., @hau2016; @taff2016].
-This approach has been widely adopted in studies of behavioral flexibility where statistical approaches and empirical progress have developed synergistically [e.g., @araya2015; @dingemanse2010; @westneat2015].
+This approach has been widely adopted in studies of behavioral flexibility where statistical methods and empirical progress have developed synergistically [e.g., @araya2015; @dingemanse2010; @westneat2015].
 This field has also benefited from simulation studies to evaluate optimal study design [@van2012] and packages that can create artificial datasets with desired patterns of between, within, and residual variance to evaluate the consequences of different patterns of variation on the ability to detect effects [see SQuID package, @allegue2017].
@@
 This standardization allows for comparison between individuals, but in some cases it may also completely obscure the ability to detect variation in certain characteristics of the acute response curve.
-For example, if the speed and scope (maximum value) of the acute response vary independently, samples taken at only two time points cannot accurately capture variation in either parameter.
+For example, if the speed (rate of initial increase) and scope (maximum value) of the acute response vary independently, samples taken at only two time points cannot accurately capture variation in either parameter.
 Indeed, several discussions in recent years about methods such as the '3 minute rule' and the relative merits of 'area under the curve' versus time point measures of glucocorticoids are fundamentally related to a recognition of the importance of understanding variation in the functional shape of stress responses and whether different components of that shape covary within individuals [e.g., @cockrem2002; @small2017].
@@
 For example, if simulations suggest that the baseline paired with stress-induced paradigm only works well when the speed and scope of responses are positively correlated, then empirical work could seek to determine the degree of correlation for a particular study system as justification for the approach.
-This ability to highlight key assumptions and create data sets with known properties has the ability to both provide insight into physiological flexibility directly and to guide empirical work by improving study design and identifying key areas for subsequent sampling.
+This ability to highlight key assumptions and create data sets with known properties has the potential to both provide insight into physiological flexibility directly and to guide empirical work by improving study design and identifying key areas for subsequent sampling.
 In the rest of this paper, I develop a simple simulation of acute physiological stress responses and then briefly illustrate several possible applications of the simulation.
@@
 
-I developed a set of simulation functions in R version 4.0.2 [@rcore] to generate acute physiological response curves.
+I developed a package called `simcoRt` in R version 4.0.2 [@rcore] with functions to generate acute physiological response curves.
 This simulation makes no assumptions about the  mechanistic process that results in the shape of a glucocorticoid response.
-Rather, parameters are sampled to generate curves that are similar in shape and degree of variation to empirically observed responses (Figure 1).
+Rather, parameters are sampled to generate curves that are similar in shape and degree of variation to empirically observed responses (Figure \@ref(fig:concept-fig)).
 This simulation is designed to create data sets with realistic structure that can be used to better design and plan studies of physiological flexibility, to evaluate power of current study designs, and to evaluate the sensitivity of sampling regimes to any number of modifications to the shape of glucocorticoid response curves (e.g., changing covariation patterns between different features of the response).
@@
 For illustration purposes, I refer to simulated glucocorticoid responses, but the simulation applies equally well to any physiological mediator of a rapid response.
+The package can be installed in R using the following command.
 
-![](concept.png)\
+```{r simp-ackage, message = FALSE, warning = FALSE}
+devtools::install_github("cct663/simcoRt")
+```
 
-***Figure 1.** Conceptual illustration of the structure of the simulation.
+```{r concept-fig, echo = FALSE, message = FALSE, warning = FALSE, out.width = '90%', fig.cap = "Conceptual illustration of the structure of the simulation.
 For each simulated animal, seven parameters are sampled from a multivariate normal distribution.
@@
 The mean and standard deviation for each parameter can be set along with the degree of covariation between each pair of parameters.
-Note that the simulation can easily be simplified as desired by setting some parameter mean or standard deviations to zero.*
+Note that the simulation can easily be simplified as desired by setting some parameter mean or standard deviations to zero."}
+knitr::opts_chunk$set(fig.pos = "!H")
+
+knitr::include_graphics("concept.png")
+
+```
 
 The simulation is constructed as two main functions with several minor functions for downstream analysis.
-I have included a detailed description of the functions and arguments in the appendix along with a full vignette detailing the process of specifying values for each function (https://github.com/cct663/speed_vs_scope).
-Briefly, function `cort_sim1` samples the parameters shown in Figure 1 from an arbitrary number of animals.
+Detailed descriptions of the arguments to each function and a full vignette are included with the package documentation.
+Briefly, function `cort_sim1` samples the parameters shown in Figure \@ref(fig:concept-fig) from an arbitrary number of animals.
 These parameters are sampled from a multivariate normal distribution with user specified mean, variance, and covariance for each parameter.
@@
 First, within-individual variation in expression is represented by specifying what amount of variation in the observation of each parameter is determined by the true value and what amount is determined by an additional randomly sampled response, based on the population parameters (this additional sampling maintains the user specified covariance structure of the population).
-Second, additional noise can be added to all measurements to represent measurement error (e.g., assay error, inaccurate pipetting, etc).
 After sampling the parameters, values are interpolated for each one minute time point and a localized regression is fit to create a smoothed curve that represents the observed glucocorticoid response.
+From this expressed response, individual data points are then collected at user specified times that would reflect an empirical study design (e.g., 2, 30, and 60 minutes).
+Additional noise is added to these data points to represent measurement error (e.g., assay error).
 
 The function also generates a simulated performance (e.g., fitness) measure, based on the underlying true values.
-In addition to returning the full curve, a down sampled dataset is returned that retains measurements made only at the user specified fixed time points (e.g., only at 2 and 30 minutes, as is typical in many field based studies).
-Data generated from this function can then be used in downstream analyses with any standard statistical approaches or software.
-For example, a user could perform an analysis to ask whether a known relationship between fitness and a particular true parameter is recovered in an analysis that includes only measures taken at particular time points.
-An additional convenience function summarizes the output of a simulation run in a multi-panel plot (Figure 2).
+Data reflecting the true phenotypic values, the repeated expression of acute responses, and the observed time points can then be used in downstream analyses with any standard statistical approaches or software.
+For example, a user could perform an analysis to ask whether a known relationship between fitness and a particular true parameter is recovered in a study that includes only measures taken at particular time points.
+An additional convenience function summarizes the output of a simulation run in a multi-panel plot (Figure \@ref(fig:initial-demo)).
 
-![](demo.png)\
-
-***Figure 2.** Example of simulation output with default settings.
-Panel A shows the down sampled data set for this run with samples collected at 1, 15, and 30 minutes in this case.
+```{r initial-demo, results = FALSE, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10.2, fig.height = 7.5, fig.cap = "Example of simulation output with default settings.
+Panel A shows the downsampled data set for this run with samples collected at 1, 15, and 30 minutes in this case.
 Panel B shows the full observed response curve for each animal.
 Panel C shows the rank order of glucocorticoid level at each time point for each animal.
-In each panel, the vertical dashed lines represent the three time points that might have been measured in a typical empirical study.*
+In each panel, the vertical dashed lines represent the three time points that might have been measured in a typical empirical study.
+Note that individuals in the top panels do not match perfectly because measurement error is added to the downsampled dataset in panel A."}
+  # Make plot with default settings from package 
+     set.seed(100)   # makes it reproducible by using same random seed
+      demo <- cort_sim2()
+      p <- plot_cort_sim(demo)
+      p
+```
 
 Finally, given recent interest in estimating the repeatability of glucocorticoid regulation [@taff2018; @cockrem2013; @hau2016], I also included a function that takes input from `cort_sim2` and calculates the observed repeatability of several measures using package `rptR` [@stoffel2017].
-Full details are included in the appendix, but this function returns repeatability for each individual time point specified in the down sampled data set, profile repeatability [@reed2019], and for area under the curve calculated as both increase (AUC~I~) and ground (AUC~G~) approaches [@pruessner2003].
-For each AUC measure, I return repeatability for the full time course, for an estimate using only the observed values in the down sampled data set, and for the full data set constrained to the time period encompassing the observed data points.
+Full details are included in the package documentation, but this function returns repeatability for each individual time point specified in the down sampled data set, profile repeatability [@reed2019], and repeatability for area under the curve calculated as both increase (AUC~I~) and ground (AUC~G~) approaches [@pruessner2003].
+For each AUC measure, the function returns repeatability for the full time course, for an estimate using only the observed values in the down sampled data set, and for the full data set constrained to the time period encompassing the observed data points.
+Simple plots illustrating repeated samples from the same individuals are also returned by default.
 
@@
 
+In order for simulation to be useful, we should be able to create artificial datasets that have similar characteristics to empirical data for different systems.
+Simulating realistic data provides a starting point for evaluating different study designs and the consequences of changes in different assumptions or parameters.
+Simulating realistic data is also useful because it can aid in study design or be used as a basis for pre-registered reports that demonstrate the feasibility of a planned study before data are ever collected.
+Simulated data can be created and entered in a complete analysis pipeline, with empirical data substituted later.
+In addition to helping to design better studies, this approach has the advantage of increasing the transparency and reliability for studies of physiological flexibility, by making analysis choices and predictions clear before data are collected.
+
+Here, I compare simulated data to an empirical dataset from red-winged blackbirds (*Agelaius phoeniceus*) to demonstrate how `simcoRt` can create synthetic datasets similar to observed data.
+This dataset is included in the `simcoRt` package and is fully described in the documentation.
+Briefly, acute glucocorticoid responses were measured using a 7-point sampling scheme, with corticosterone measured at 1, 3, 7, 15, 30, 45, and 60 minutes post capture.
+I simulated data using the functions described above starting with the input values calculated directly from the empirical data.
+The simulation creates a new dataset that has similar variation and patterns to empirical data (Figure \@ref(fig:rwbb-demo)A) along with a population wide corticosterone response curve shape that closely matches the empirical data (Figure \@ref(fig:rwbb-demo)B).
+In this case, the plotted simulation data include the same number of animals sampled at the same time points as the empirical data, but these sampling points and total sample size can easily be changed as desired.
+The parameterized simulation can now be used to test the sensitivity of any number of experimental designs before additional data is collected.
+
+```{r rwbb-demo, echo = FALSE, fig.height = 3.5, fig.width = 8, message = FALSE, warning = FALSE, fig.cap = "Panel A shows the acute corticosterone response for measured (blue) or simulated (orange) red-winged blackbirds measured at seven time points.
+Panel B shows the mean and standard error of the two datasets."}
+      rwde <- subset(rwbb, rwbb$lh_stg == "early-breeding")  
+      rwde <- subset(rwde, rwde$id != "452")   # This one has an unusually high cort response
+    set.seed(165)
+        sim_rwd <- cort_sim2(cort_sim1(
+       n = 23,
+       base_mu = 3.5,
+       base_sd = 3.5,
+       slope_mu = 17,
+       slope_sd = 0.7,
+       fastpct_mu = 0.7,
+       fastpct_sd = 0.03,
+       max_mu = log(56),
+       max_sd = 0.5,
+       speed_mu = 65,
+       speed_sd = 5,
+       return_mu = 120,
+       maxtime_mu = 25,
+       cor_max_slope = -0.7
+     ), 
+     sample_times = 1,
+     bleed_times = c(1, 3, 8, 15, 30, 45, 60),
+     assay_error = 0.2,
+     sm_span = 0.2)
+
+     sim_rwd$simulated_dataset_long$type <- "simulated"
+     sim_rwd2 <- sim_rwd$simulated_dataset_long[, c("animal_sample", "time", "cort", "type")]
+     rwde2 <- rwde[, c("id", "lat_sec", "cort")]
+     rwde2$lat_sec <- rwde2$lat_sec / 60
+     rwde2$type <- "red-winged blackbird"
+     colnames(rwde2) <- c("animal_sample", "time", "cort", "type")
+
+     comp_sim <- rbind(rwde2, sim_rwd2)
+
+     p1 <- ggplot(comp_sim, mapping = aes(x = time, y = cort, by = animal_sample, color = type)) +
+       geom_line(alpha = 0.7, size = 0.7) +
+       theme_classic() +
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       scale_color_manual(values = c("slateblue", "orange")) +
+       guides(color = FALSE) +
+       theme(axis.title = element_text(size = 14)) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A")
+
+     p2 <- ggplot(comp_sim, mapping = aes(x = time, y = cort, color = type, fill = type)) +
+       geom_smooth() +
+       theme_classic() + 
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       scale_color_manual(values = c("slateblue", "orange")) +
+       scale_fill_manual(values = c("slateblue", "orange")) +
+       theme(legend.position = c(0.7, 0.15), axis.title = element_text(size = 14)) +
+       guides(color = guide_legend(title = ""), fill = guide_legend(title = "")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B")
+
+      ggpubr::ggarrange(p1, p2, nrow = 1) 
+```
+
+### *Accurately measuring a single glucocorticoid trait*
+
+Single time point measures of glucocorticoids are often interpreted as representing meaningful variation between individuals.
+For example, variation in the level of glucocorticoids after 30 minutes of standardized restraint is typically interpreted as variation in the magnitude of the stress response [@taff2019].
+However, this interpretation rests on assumptions that are rarely explicitly tested with empirical data.
+For example, the time chosen to take a stress-induced sample is often assumed to be either at the species peak or during a plateau period after the species peak.
+In some early studies, great care was taken to determine an average population level peak time [CITE], but many studies adopt the widely used 'standard' time of 30 minutes post capture without extensive validation.
+While there is a general assumption that sampling later than the peak is acceptable (and perhaps preferable) because animals will be sampled during a relatively stable high plateau, there is little empirical data to evaluate this assertion or to determine how much under or overshooting the species peak timing might influence inferences.
+Furthermore, even when the average peak timing is well established, differences in the amount of between-individual variation in the time to reach the peak or in peak values are common across species and even in different life history stages within species [CITE].
+The combinations of these patterns of variation could have major consequences on the accuracy of single point estimates taken at 30 minutes, but these questions cannot be addressed directly with empirical datasets where the true underlying values of each individual are unknown.
+
+Here, I simulate a simple scenario exploring the consequences of variation in each of these parameters on the accuracy of estimating between individual differences in maximally expressed glucocorticoids during an acute response.
+For purposes of this illustration, I consider a single study design in which animals are sampled at 30 minutes.
+Using this design as a starting point, I systematically vary i) the timing of the population average peak (15, 30, or 45 minutes), ii) the amount of variation in maximum glucocorticoid levels reached, iii) and the amount of variation in the number of minutes taken to reach peak levels.
+All other variables in the simulation are constrained to be invariant between individuals in the population (e.g., all individuals have identical baseline glucocorticoids in this case), though I consider cases in which multiple aspects of the rapid response are correlated with each other in the next section.
+I included moderate within-individual variability and a small amount of assay error across all iterations.
+For each combination of parameters, I simulated 200 animals and estimated the R^2^ value from a regression of the observed estimates of glucocorticoid levels at 30 minutes to the true known values.
+This simulation is likely a best case scenario because it eliminates many sources of variation or noise that would be present in real data, but it illustrates the effect of variation in these three key parameters even when the exact same sampling design is employed.
+
+```{r max-sim, echo = FALSE, warning = FALSE, message = FALSE, fig.pos = "!h", fig.align = "center", fig.width = 8, fig.height = 3, fig.cap = "Results of simulation runs with different amounts of between-individual variation in the time to reach maximum glucocorticoid levels and in the maximum level reached.
+Simulations are run with samples taken at 30 minutes on populations with an average peak time of 15 minutes (left), 30 minutes (center), or 45 minutes (right).
+Each grid cell is the R$^2$ value from the regression of observed glucocorticoids at 30 minutes to true maximum levels in a simulation of 200 individuals."}
+# set.seed(901)
+# grid.size <- 15
+# 
+# max_sd <- seq(0.01, 0.26, length.out = grid.size)
+# time_sd <- seq(1, 20, length.out = grid.size)
+# #plateau <- c(2, 20)
+# speeds <- c(15, 30, 45)
+# 
+# storer <- data.frame(max_sd = rep(rep(max_sd, grid.size), 3),
+#                      time_sd = rep(rep(time_sd, each = grid.size), 3),
+#                      rsq = rep(rep(NA, grid.size*grid.size), 3),
+#                      slp = rep(rep(NA, grid.size*grid.size), 3),
+#                      #plateau = rep(c("Short plateau", "Long plateau"), each = 300),
+#                      speeds = rep(c("Mean max 15m", "Mean max 30m", "Mean max 45m"), each = grid.size*grid.size))
+# 
+# counter <- 0
+# 
+# #for(u in 1:length(plateau)){
+#   for(k in 1:length(speeds)){
+#     for(i in 1:length(time_sd)){
+#         for(j in 1:length(max_sd)){
+#           dm <- cort_sim2(data = cort_sim1(n = 200, base_mu = 0, base_sd = 0, slope_sd = 0, slope_mu = 5, fastpct_mu = 0.4, fastpct_sd = 0, maxtime_sd = 0, return_sd = 0,
+#                                          speed_mu = speeds[k], 
+#                                          speed_sd = time_sd[i],
+#                                          max_mu = log(45),
+#                                          max_sd = max_sd[j],
+#                                          maxtime_mu = 10,
+#                                          return_mu = speeds[k] + 60
+#                                          ),
+#                         bleed_times = c(2, 30),
+#                         sample_times = 1)
+#                         #base_error = 0, speed_error = 0, max_error = 0, maxtime_error = 0, return_error = 0, slope_error = 0, fastpct_error = 0)
+#         
+#           dm2 <- data.frame(max30 = subset(dm$simulated_dataset_long$cort, dm$simulated_dataset_long$time == 30),
+#                           max_true = dm$true_values$max)
+#           m <- lm(max_true ~ max30, data = dm2)
+#           rs <- summary(m)$r.squared
+#           
+#           storer[counter + j, 3] <- rs
+#           storer[counter + j, 4] <- coef(m)[2]
+#         }
+#       counter <- counter + grid.size
+#       print(paste(i, "of", length(time_sd), sep = " "))
+#     }
+#     print(paste(k, "of", length(speeds)))
+#   }
+# #  print(paste(u, "of", length(plateau)))
+# #}
+
+#saveRDS(storer, here::here("5_other_outputs/max_sim.rds"))
+storer <-readRDS(here::here("5_other_outputs/max_sim.rds"))
+
+# Making bins for r squared values
+  storer$rsq2 <- cut(storer$rsq, breaks = seq(0, 1, 0.2))
+  new_bins <- data.frame(rsq2 = levels(storer$rsq2),
+                         rsq3 = c("0.0 - 0.2", "0.2 - 0.4", "0.4 - 0.6", "0.6 - 0.8", "0.8 - 1.0"))
+  storer <- plyr::join(storer, new_bins, "rsq2")
+
+# Putting max sd onto interpretable scale
+  storer$max_sd2 <- (exp(storer$max_sd + rep(log(45), nrow(storer))) - exp(rep(log(45), nrow(storer)) - storer$max_sd)) / 2
+
+ggplot(data = storer, mapping = aes(x = time_sd, y = max_sd2, fill = rsq3)) +
+  geom_tile() +
+  scale_fill_viridis(option = "plasma", discrete = TRUE) +
+  theme_bw() +
+  theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
+  ylab("Variation in max (SD ng/\U00B5l)") +
+  xlab("Variation in time to reach max (SD minutes)") +
+  facet_wrap(~ speeds, nrow = 1) +
+  #facet_grid(rows = vars(plateau), cols = vars(speeds)) +
+  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12)) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  labs(fill = expression(paste("R"^{2})))
+
+```
+
+Results of this simulation are summarized in figure \@ref(fig:max-sim).
+The amount of between-individual variation in the maximum glucocorticoid value has a profound effect on the ability to detect true maximal levels with samples taken at 30 minutes.
+In one sense, this result is unsurprising because it is intuitive that large differences would be easier to detect, but there are important consequences of this fact for interpreting studies that seek to link between-individual variation in the magnitude of the stress response with other traits.
+For example, the magnitude of the acute stress response often varies substantially across life history stages [CITE].
+Even if study designs are identical and maximum glucocorticoids are associated with performance, it will be easier to detect those patterns during life history stages with greater variation (see section on detecting fitness associations below).
+There is a weaker, but still substantial impact of variation in the time taken to reach maximum values on the accuracy of estimates in this simulation.
+Greater variation in time reduces the accuracy of estimates of maximal values.
+Finally, the timing of sampling relative to the average population peak timing also influences accuracy.
+Measuring after the average peak time results in the most accurate estimates across a range of parameter values, while measuring before the average peak time produces the least accurate measures, particularly when there is also high variation in the time to reach maximum values between individuals.
+This simple example demonstrates clearly that the same experimental design will perform better or worse depending on the combination of glucocorticoid regulation parameters in the population being studied.
+
 ### *Exploring covariance between response components*
 
-### *Comparing repeatability of different measures*
+In reality, fully characterizing the acute glucocorticoid response requires more than identifying just the maximum value reached.
+Individuals may differ in baseline levels, rate of initial increase, the speed of reaching the maximum level, time spent at maximum, and the speed of return to baseline.
+Moreover, each of these components of the endocrine response could be positively or negatively correlated with each other within and between individuals.
+In these cases, measurements taken at particular time points contain information about multiple aspects of the response and without additional information it may be difficult to know what trait is being measured.
+The fact that each of these traits might be important and that they might covary has been discussed in a general sense [CITE], but simulations are uniquely powerful for exploring under exactly what conditions time point measure of glucocorticoids can or cannot be used as indicators of these traits.
+
+```{r cor-ex1, echo = FALSE, fig.height = 2.8, fig.width = 7.3, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Simulated glucocorticoid responses in which the maximum value and response speed are positively correlated (A), uncorrelated (B), or negatively correlated (C).
+Orange curves show the full response for each individual.
+Blue points show the maximum value and time to reach maximum for each individual.
+Blue lines are simple linear regressions of speed and maximum value for each group.
+For clarity, only the first 40 individuals in each simulated dataset are plotted."}
+# 
+# set.seed(431)
+# 
+# spd_sd <- 12
+# mx_sd <- 0.21
+# 
+# c_pos <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = -0.6, speed_mu = 35, speed_sd = spd_sd, base_sd = 0,
+#                                     slope_mu = 8, slope_sd = 0.1, fastpct_sd = 0, max_sd = mx_sd, speed_min = 8),
+#                      bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+# c_zer <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = 0, speed_mu = 35, speed_sd = spd_sd, base_sd = 0,
+#                                     slope_mu = 8, slope_sd = 0.1, fastpct_sd = 0, max_sd = mx_sd, speed_min = 8),
+#                      bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+# c_neg <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = 0.6, speed_mu = 35, speed_sd = spd_sd, base_sd = 0,
+#                                     slope_mu = 8, slope_sd = 0.1, fastpct_sd = 0, max_sd = mx_sd, speed_min = 8),
+#                     bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+# 
+# c1 <- c_pos$timecourse_long
+# c1$run <- "faster = higher"
+# c2 <- c_zer$timecourse_long
+# c2$run <- "uncorrelated"
+# c3 <- c_neg$timecourse_long
+# c3$run <- "faster = lower"
+# 
+# c_comb <- rbind(c1[1:(50 * 171), ], c2[1:(50 * 171), ], c3[1:(50 * 171), ])
+# c_comb$as_run <- paste(c_comb$animal, c_comb$run, sep = "_")
+# c_comb_tem <- c_comb[!duplicated(c_comb$as_run), ]
+# 
+# c_comb$run <- factor(c_comb$run, levels = c("faster = higher", "uncorrelated", "faster = lower"))
+# 
+# for(i in 1:nrow(c_comb_tem)){
+#   sub <- subset(c_comb, c_comb$as_run == c_comb_tem$as_run[i])
+#   maxc <- sub[which.max(sub$cort), ]
+#   if(i == 1){max_vals <- maxc}
+#   if(i > 1){max_vals <- rbind(max_vals, maxc)}
+# }
+# 
+# # Add text with dataframe that has a/b/c
+# 
+# text_labs <- data.frame(time = rep(1, 3),
+#                         cort = rep(77, 3),
+#                         run = levels(c_comb$run),
+#                         labs = c("A", "B", "C"),
+#                         animal = rep(NA, 3))
+# 
+# text_labs$run <- factor(text_labs$run, levels = c("faster = higher", "uncorrelated", "faster = lower"))
+# 
+# #Bookdown wasn't rendering this figure properly so I've saved and loaded it here.
+# pcor <- ggplot(data = c_comb, mapping = aes(x = time, y = cort)) +
+#   geom_line(color = "orange", alpha = 0.5, mapping = aes(by = animal)) + theme_bw() +
+#   facet_wrap(~ run) +
+#   geom_smooth(data = max_vals, se = FALSE, color = "slateblue", alpha = 0.6, method = "lm") +
+#   coord_cartesian(xlim = c(0, 80)) +
+#   geom_point(data = max_vals, color = "slateblue", alpha = 0.6) +
+#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   xlab("Time (minutes)") + ylab("Glucocorticoids (ng/\U00B5l)") +
+#   theme(axis.title = element_text(size = 12)) +
+#   geom_text(data = text_labs, mapping = aes(label = labs), size = 5)
+
+#saveRDS(pcor, here::here("5_other_outputs/example_corr.rds"))
+
+pcor <- readRDS(here::here("5_other_outputs/example_corr.rds"))
+pcor
+
+```
+
+To illustrate this point, I explored the consequences of variation in the correlation between and relative amount of variation in just two aspects of the acute stress response: the maximum glucocorticoid level reached and the time required to reach the maximum level.
+For simplicity, I refer to the 'speed' of the response, but note that other aspects, such as the rate of initial increase, could also be considered as variation in the speed of response.
+When considering these two traits, a population of animals could plausibly display one of three patterns.
+Individuals that reach their maximum value faster might also reach higher values (figure \@ref(fig:cor-ex1)A; simulation correlation = -0.6).
+Alternatively, the speed and maximum values might vary independently (figure \@ref(fig:cor-ex1)B; correlation = 0).
+Finally, individuals that are faster might max out at lower glucocorticoid values (figure \@ref(fig:cor-ex1)C; correlation = 0.6).
+While many researchers in this field might have intuitions about which of these scenarios is most likely to prevail, there is very little empirical data available to actually determine which is most common.
+Moreover, regardless of the specifics for this particular correlation, the general pattern and considerations presented here will apply in similar ways to correlations between other aspects of the acute stress response.
+
+Using these three simulated populations as a starting point, I asked how well glucocorticoid values measured at one timepoint reflected true trait values.
+For each population I set an average population level speed of 30 minutes with other values in the simulation set at their default value.
+For every time point from 0 to 35 minutes I fit two simple linear regressions of the measured value on the true speed and maximum value and extracted the R^2^ value from the model.
+I repeated this simulation for all populations 50 times with 100 individuals sampled from the population each time.
+Finally, I repeated the entire set of simulations with each combination of low and high between-individual variation in the speed or maximum values (variation in speed: low = 2 minute SD, high = 12 minute SD; variation in maximum: low = 1ng/$\mu$l SD, high = 10ng/$\mu$l SD).
+
+```{r cor-ex2, echo = FALSE, fig.height = 6, fig.width = 7, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Relationship between single time point measures of glucocorticoids and the true value of either maximum level (solid lines) or the speed of the glucocorticoid response (dashed lines).
+Panels show results when the overall variation in maximum values and speed are both low (A), when one is low while the other is high (B and C), and when both are high (D).
+In each panel, three different simulation scenarios illustrate the patterns when speed and maximum value are positively correlated (purple), uncorrelated (teal), or negatively correlated (yellow).
+Faded lines show the results from each of 50 separate simulation runs and thick lines are the averages across all runs."}
+
+# spd_sd <- c(2, 2, 12, 12)
+# mx_sd <- c(0.02, 0.21, 0.02, 0.21)
+# 
+# for(u in 1:4){
+#   for(k in 1:50){
+#     # Create simulations that have positive, no, or negative correlatoins b/w speed and scope
+#       c_pos <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = -0.6, speed_mu = 30, speed_sd = spd_sd[u], 
+#                                           slope_mu = 8, max_sd = mx_sd[u], speed_min = 8), 
+#                            bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+#       c_zer <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = 0, speed_mu = 30, speed_sd = spd_sd[u],
+#                                           slope_mu = 8, max_sd = mx_sd[u], speed_min = 8), 
+#                            bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+#       c_neg <- cort_sim2(data = cort_sim1(n = 100, cor_speed_max = 0.6, speed_mu = 30, speed_sd = spd_sd[u],
+#                                           slope_mu = 8, max_sd = mx_sd[u], speed_min = 8),
+#                           bleed_times = c(1, 5, 10, 15, 30), sample_times = 1)
+#     
+#     # Determine correlation between each timepoint measure and speed and max for each simulation and combine
+#           times1 <- data.frame(time = unique(c_pos$timecourse_long$time),
+#                         cor_spd = NA,
+#                         cor_max = NA,
+#                         run = "faster = higher")
+#           for(i in 1:50){
+#             dat <- subset(c_pos$timecourse_long, c_pos$timecourse_long$time == i)
+#             dat <- plyr::join(dat, c_pos$true_values, "animal")
+#             times1$cor_spd[i] <- cor(dat$cort, dat$speed) ^ 2
+#             times1$cor_max[i] <- cor(dat$cort, dat$max) ^ 2
+#             
+#           
+#           }
+#           
+#           times2 <- data.frame(time = unique(c_zer$timecourse_long$time),
+#                               cor_spd = NA,
+#                               cor_max = NA,
+#                               run = "uncorrelated")
+#           for(i in 1:50){
+#             dat <- subset(c_zer$timecourse_long, c_zer$timecourse_long$time == i)
+#             dat <- plyr::join(dat[, c("animal", "time", "cort")], c_zer$true_values, "animal")
+#             times2$cor_spd[i] <- cor(dat$cort, dat$speed) ^ 2
+#             times2$cor_max[i] <- cor(dat$cort, dat$max) ^ 2
+#             ggplot(data = dat, mapping = aes(x = cort, y = speed)) + geom_smooth(method = "lm") + geom_point()
+#             cor(dat$cort, dat$speed) ^2
+#           }
+#           
+#           times3 <- data.frame(time = unique(c_neg$timecourse_long$time),
+#                               cor_spd = NA,
+#                               cor_max = NA,
+#                               run = "faster = lower")
+#           for(i in 1:50){
+#             dat <- subset(c_neg$timecourse_long, c_neg$timecourse_long$time == i)
+#             dat <- plyr::join(dat, c_neg$true_values, "animal")
+#             times3$cor_spd[i] <- cor(dat$cort, dat$speed) ^ 2
+#             times3$cor_max[i] <- cor(dat$cort, dat$max) ^ 2
+#           }
+#           
+#           times <- rbind(times1, times2, times3)
+#           times$spd_sd <- rep(spd_sd[u], nrow(times))
+#           times$max_sd <- rep(mx_sd[u], nrow(times))
+#           times$run_num <- paste(u, k, sep = "_")
+#           
+#           if(k == 1){if(u == 1){time_out <- times}}
+#           if(k == 1){if(u > 1){time_out <- rbind(time_out, times)}}
+#           if(k > 1){time_out <- rbind(time_out, times)}
+#           
+#           print(paste(k, "of 50 run", u, "of 4", sep = " "))
+#     
+#   }
+#   
+# }
+# 
+# time_out2 <- pivot_longer(time_out, cols = cor_spd:cor_max, names_to = "type", values_to = "correl")
+# time_out2$run <- factor(time_out2$run, levels = c("faster = higher", "uncorrelated", "faster = lower"))
+# 
+# mx <- data.frame(max_sd = c(0.02, 0.21),
+#                  max_sd2 = c("Low variation in max", "High variation in max"))
+# sp <- data.frame(spd_sd = c(2, 12),
+#                  spd_sd2 = c("Low variation in speed", "High variation in speed"))
+# 
+# time_out2 <- plyr::join(time_out2, mx, "max_sd")
+# time_out2 <- plyr::join(time_out2, sp, "spd_sd")
+# 
+# time_out2$max_sd2 <- factor(time_out2$max_sd2, levels = c("Low variation in max", "High variation in max"))
+# time_out2$spd_sd2 <- factor(time_out2$spd_sd2, levels = c("Low variation in speed", "High variation in speed"))
+# 
+# time_out2$type <- gsub("cor_max", "maximum", time_out2$type)
+# time_out2$type <- gsub("cor_spd", "speed", time_out2$type)
+# 
+# to_lab <- data.frame(labs = c("A", "B", "C", "D"),
+#                      run = rep("uncorrelated", 4),
+#                      correl = rep(0.98, 4),
+#                      type = rep(NA, 4),
+#                      time = rep(0, 4),
+#                      run_num = rep(NA, 4),
+#                      max_sd = c(0.02, 0.02, 0.21, 0.21),
+#                      spd_sd = c(2, 12, 2, 12),
+#                      max_sd2 = c(rep("Low variation in max", 2), rep("High variation in max", 2)),
+#                      spd_sd2 = rep(c("Low variation in speed", "High variation in speed"), 2))
+# 
+# to_lab$max_sd2 <- factor(to_lab$max_sd2, levels = c("Low variation in max", "High variation in max"))
+# to_lab$spd_sd2 <- factor(to_lab$spd_sd2, levels = c("Low variation in speed", "High variation in speed"))
+# 
+# p4 <- ggplot(data = time_out2, mapping = aes(x = time, y = correl, color = run, linetype = type, fill = run)) +
+#   geom_line(mapping = aes(by = run_num), alpha = 0.1) +
+#   geom_smooth(alpha = 0.4, se = FALSE) +
+#   scale_color_viridis(discrete = TRUE) +
+#   scale_fill_viridis(discrete = TRUE) +
+#   theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   coord_cartesian(xlim = c(0, 35), ylim = c(0, 1)) +
+#   theme(axis.title = element_text(size = 13)) +
+#   #geom_line(mapping = aes(y = cor_max), linetype = "dashed", size = 2) +
+#   xlab("Time (minutes)") + ylab(expression(paste("R"^{2}, " measured vs. true trait"))) +
+#   facet_grid(rows = vars(max_sd2), cols = vars(spd_sd2)) +
+#   geom_text(data = to_lab, mapping = aes(label = labs), color = "black", size = 5) +
+#   labs(linetype = "Trait", color = "Correlation") + guides(fill = FALSE)
+#   
+# 
+# saveRDS(p4, here::here("5_other_outputs/example_corr2.rds"))
+
+p4 <- readRDS(here::here("5_other_outputs/example_corr2.rds"))
+p4 + theme(legend.position = "top")
+```
+
+The time that samples were taken at, relative amount of variation in speed and maximum, and degree of correlation between the speed and maximum all had substantial impacts on the ability to infer true trait values from single time point glucocorticoid measures (figure \@ref(fig:cor-ex2)).
+While these scenarios do not explore all possible parameter space, there are several clear conclusions that can be made.
+First, neither speed or maximum traits could be assessed accurately when between-individual variation in both traits was low (figure \@ref(fig:cor-ex2)A).
+This is potentially important for interpreting apparent differences in glucocorticoid fitness relationships because between-individual variation is known to differ across life history stages [CITE].
+Second, accurately assessing variation in speed was much harder-if not impossible-with single measures.
+
+Only when speed varied between individuals with very low variation in maximum was it possible to accurately measure speed, but this situation may be rare in natural populations.
+When speed was tightly correlated with maximum (figure \@ref(fig:cor-ex2)D) it was sometimes possible to attain reasonable estimates of speed (figure \@ref(fig:cor-ex2)C-D), but when speed was not correlated with maximum single measures were not good indicators of variation in speed (figure \@ref(fig:cor-ex2)A, C-D).
+Finally, measuring variation in maximum values was much easier under many conditions (figure \@ref(fig:cor-ex2)C-D), but the accuracy of assessment of maximum values was also negative impacted by variation in speed and the degree of this impact differed depending on the correlation between the two traits (figure \@ref(fig:cor-ex2)).
+Beyond the specifics of this particular example, what these results demonstrate clearly is that understanding what aspect of the glucocorticoid response is being measured by any particular study design depends on extensive knowledge of the overall shape and amount of variation in different aspects of the acute stress response.
+
+### *Detecting links between fitness and responses*
+
+A common goal of recent studies is to establish whether variation in glucocorticoids is associated with fitness or some proxy for fitness [@schoenle2020].
+While there has been a great deal of discussion about the extent to which these relationships might differ with life history characteristics or between breeding stages [CITE], there has been relatively little consideration of the way that methodological limitations might limit the ability to detect these relationships even when they exist.
+
+```{r fitness-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Relationship between observed maximum glucocorticoid values and fitness for simulated populations that have low between-individual variation (blue) or high between-individual variation (red).
+Each point is the result of a separate simulation of 50 individuals using the settings described in the text.
+Filled circles are simulations in which observed glucocorticoid values at 30 minutes were significantly correlated with fitness and crosses are simulations in which the relationship was not significant.
+Panels illustrate conditions with low measurement error (A, C) versus high measurement error (B, D) and low within-individual variation (A, B) versus high within-individual variation (C, D).
+For each simulation, the correlation between true maximum glucocorticoids fitness is plotted on the y-axis and the correaltion with observed values is plotted on the x-axis."}
+#draw figure ----
+# n_sam <- 50
+# fit_dat <- data.frame(f1b_true = rep(NA, n_sam), f1b_obs = rep(NA, n_sam), f1b_p = rep(NA, n_sam),
+#                       f2b_true = rep(NA, n_sam), f2b_obs = rep(NA, n_sam), f2b_p = rep(NA, n_sam))
+# 
+# for(i in 1:n_sam){
+#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.1, max_error = 0.2,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
+#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
+#     f1b <- subset(f1b, f1b$time == 30)
+#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
+#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
+#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
+#     
+#     
+#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.1, max_error = 0.2)
+#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
+#     f2b <- subset(f2b, f2b$time == 30)
+#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
+#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
+#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
+#     print(paste(i, "of", n_sam, sep = " "))
+# }
+# 
+# fd1 <- fit_dat[, 1:3]
+# fd2 <- fit_dat[, 4:6]
+# colnames(fd1) <- c("true", "obs", "p")
+# colnames(fd2) <- c("true", "obs", "p")
+# 
+# fd1$dif <- fd1$obs - fd1$true
+# fd2$dif <- fd2$obs - fd2$true
+# 
+# fd1$group <- "High"
+# fd2$group <- "Low"
+# 
+# f3 <- rbind(fd1, fd2)
+# 
+# f3$group <- as.factor(f3$group)
+# f3$shp <- NA
+# 
+# for(i in 1:nrow(f3)){
+#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
+# }
+# f3a <- f3
+# pa <- ggplot(data = f3a, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
+#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
+#   #geom_smooth(method = "lm", alpha = 0.3) +
+#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
+#   scale_color_manual(values = c("coral3", "slateblue")) +
+#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   theme_bw() +
+#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   ggtitle("Low measurement error \nLow within-individual variation") +
+#   xlab("") + ylab("True correlation") +
+#   theme(legend.position = c(0.75, 0.15)) +
+#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
+#   guides(fill = FALSE, color = FALSE, shape = FALSE) +
+#   scale_shape_manual(values = c(21, 3), guide = "none") +
+#   #theme(legend.background = element_rect(fill = alpha("white", 0))) +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A", size = 8) +
+#   theme(axis.title = element_text(size = 13)) 
+#   #annotate(geom = "text", x = -0.1, y = 0.52, label = "Low measurement error \nLow within-individual variation", hjust = 0)
+# 
+# pA <- ggExtra::ggMarginal(pa, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
+#  
+# ## Low assay, high within
+# 
+# for(i in 1:n_sam){
+#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.4, max_error = 0.1,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
+#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
+#     f1b <- subset(f1b, f1b$time == 30)
+#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
+#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
+#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
+#     
+#     
+#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.4, max_error = 0.1)
+#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
+#     f2b <- subset(f2b, f2b$time == 30)
+#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
+#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
+#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
+#     print(paste(i, "of", n_sam, sep = " "))
+# }
+# 
+# fd1 <- fit_dat[, 1:3]
+# fd2 <- fit_dat[, 4:6]
+# colnames(fd1) <- c("true", "obs", "p")
+# colnames(fd2) <- c("true", "obs", "p")
+# 
+# fd1$dif <- fd1$obs - fd1$true
+# fd2$dif <- fd2$obs - fd2$true
+# 
+# fd1$group <- "High"
+# fd2$group <- "Low"
+# 
+# f3 <- rbind(fd1, fd2)
+# 
+# f3$group <- as.factor(f3$group)
+# f3$shp <- NA
+# for(i in 1:nrow(f3)){
+#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
+# }
+# f3b <- f3
+# pb <- ggplot(data = f3b, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
+#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
+#   #geom_smooth(method = "lm", alpha = 0.3) +
+#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
+#   scale_color_manual(values = c("coral3", "slateblue")) +
+#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   theme_bw() +
+#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   ggtitle("High measurement error \nLow within-individual variation") +
+#   xlab("") + ylab("") +
+#   #theme(legend.position = c(0.82, 0.15)) +
+#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
+#   guides(fill = "none", color = "none") +
+#   scale_shape_manual(values = c(21, 3), guide = "none") +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B", size = 8) +
+#   theme(axis.title = element_text(size = 13)) 
+#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "High measurement error \nLow within-individual variation", hjust = 0)
+# 
+# pB <- ggExtra::ggMarginal(pb, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both") 
+# 
+# ## Low assay, high within
+# 
+# for(i in 1:n_sam){
+#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.1, max_error = 0.6,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
+#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
+#     f1b <- subset(f1b, f1b$time == 30)
+#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
+#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
+#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
+#     
+#     
+#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.1, max_error = 0.6)
+#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
+#     f2b <- subset(f2b, f2b$time == 30)
+#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
+#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
+#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
+#     print(paste(i, "of", n_sam, sep = " "))
+# }
+# 
+# fd1 <- fit_dat[, 1:3]
+# fd2 <- fit_dat[, 4:6]
+# colnames(fd1) <- c("true", "obs", "p")
+# colnames(fd2) <- c("true", "obs", "p")
+# 
+# fd1$dif <- fd1$obs - fd1$true
+# fd2$dif <- fd2$obs - fd2$true
+# 
+# fd1$group <- "High"
+# fd2$group <- "Low"
+# 
+# f3 <- rbind(fd1, fd2)
+# 
+# f3$group <- as.factor(f3$group)
+# f3$shp <- NA
+# for(i in 1:nrow(f3)){
+#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
+# }
+# f3c <- f3
+# pc <- ggplot(data = f3c, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
+#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
+#   #geom_smooth(method = "lm", alpha = 0.3) +
+#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
+#   scale_color_manual(values = c("coral3", "slateblue")) +
+#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   theme_bw() +
+#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   ggtitle("Low measurement error \nHigh within-individual variation") +
+#   xlab("Observed correlation") + ylab("True correlation") +
+#   #theme(legend.position = c(0.82, 0.15)) +
+#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
+#   guides(fill = "none", color = "none") +
+#   scale_shape_manual(values = c(21, 3), guide = "none") +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C", size = 8) +
+#   theme(axis.title = element_text(size = 13)) 
+#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "Low measurement error \nHigh within-individual variation", hjust = 0)
+# 
+# pC <- ggExtra::ggMarginal(pc, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
+# 
+# ## high assay, high within
+# 
+# for(i in 1:n_sam){
+#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.3, max_error = 0.6,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
+#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
+#     f1b <- subset(f1b, f1b$time == 30)
+#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
+#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
+#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
+#     
+#     
+#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#                                 bleed_times = c(2, 30), sample_times = 1,
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.3, max_error = 0.6)
+#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
+#     f2b <- subset(f2b, f2b$time == 30)
+#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
+#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
+#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
+#     print(paste(i, "of", n_sam, sep = " "))
+# }
+# 
+# fd1 <- fit_dat[, 1:3]
+# fd2 <- fit_dat[, 4:6]
+# colnames(fd1) <- c("true", "obs", "p")
+# colnames(fd2) <- c("true", "obs", "p")
+# 
+# fd1$dif <- fd1$obs - fd1$true
+# fd2$dif <- fd2$obs - fd2$true
+# 
+# fd1$group <- "High"
+# fd2$group <- "Low"
+# 
+# f3 <- rbind(fd1, fd2)
+# 
+# f3$group <- as.factor(f3$group)
+# f3$shp <- NA
+# for(i in 1:nrow(f3)){
+#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
+# }
+# f3d <- f3
+# pd <- ggplot(data = f3d, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
+#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
+#   #geom_smooth(method = "lm", alpha = 0.3) +
+#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
+#   scale_color_manual(values = c("coral3", "slateblue")) +
+#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   theme_bw() +
+#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   ggtitle("High measurement error \nHigh within-individual variation") +
+#   xlab("Observed correlation") + ylab("") +
+#   #theme(legend.position = c(0.82, 0.15)) +
+#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
+#   guides(fill = "none", color = "none") +
+#   scale_shape_manual(values = c(21, 3), guide = "none") +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "D", size = 8) +
+#   theme(axis.title = element_text(size = 13)) 
+#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "High measurement error \nHigh within-individual variation", hjust = 0)
+# 
+# pD <- ggExtra::ggMarginal(pd, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
+# 
+# pAD <- ggpubr::ggarrange(pA, pB, pC, pD, nrow = 2, ncol = 2)
+# 
+# saveRDS(pAD, here::here("5_other_outputs/fitness-ex.rds"))
+
+# reload and print ----
+pAD <- readRDS(here::here("5_other_outputs/fitness-ex.rds"))
+pAD
+
+```
+
+Here, I imagine a simple scenario in which the 'true' maximum glucocorticoid level during an acute response explains 80% of the variation in fitness (clearly this is unrealistically high, but it is chosen for illustration only).
+I next construct a study in which researchers measure 50 individuals using a typical stress-induced (30 minute) sampling protocol.
+For simplicity, I set the other parameters in the simulation at their default values.
+Keeping the study design constant, I ask whether the glucocorticoid-fitness relationship can be recovered for two hypothetical populations that have low or high between-individual variation in maximum glucocorticoid levels.
+For each of these two populations, I ask how the ability to detect glucocorticoid-fitness relationships changes with different amounts of within-individual variation in acute response expression and with differing amounts of measurement error.
+For each combination of parameters, I simulated 50 populations and fit a simple linear regression model with observed glucocorticoid levels at 30 minutes as a predictor of fitness to ask whether the true glucocorticoid-fitness relationship was recovered.
+
+Several patterns can be identified by examining the results of this simulation.
+First, the correlation between the true maximum glucorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex) A-D).
+In all cases, however, the observed correlation is always lower than the true correlation and always lowest in the population with low between-individual variation.
+The ubquity of this pattern is a product of the simulation structure, because adding measurement error or within-individual variation effecively add noise to the true correlation.
+It is important to note that in the real world, it is unlikely that this pattern would be so universal, because unmeasured variables could influence both fitness and glucocorticoids.
+For example, if habitat quality directly alters fitness and glucocorticoids, the observed correlation could be stronger than the 'true' correlation.
+Thus, interpretation of these results should be made cautiosly in light of the simplicity of the simulation compared to real world conditions.
+
+Nevertheless, general patterns illustrated by the simulation are likely to pertain across a wide range of conditions.
+In this case, it is easiest to detect significant glucocorticoid-fitness relationships when both measurement error and within-individual variation are low (figure \@ref(fig:fitness-ex) A).
+It becomes harder to detect these true relationships when either measurement error (figure \@ref(fig:fitness-ex) B) or within-individual variation (figure \@ref(fig:fitness-ex) C) are high, but even in these more challenging situations the relationship can be detected the majority of the time if between-individual variation in maximum levels is high.
+When both measurement error and within-individual variation are high, it is nearly impossible to detect glucocorticoid-fitness relationships with low-between individual variation, but in populations with high between-individual variation the relationship is still detected in about half of the simulations.
+
+The fact that low between-individual variation in maximum glucocorticoids makes it harder to detect true glucocorticoid-fitness relationships across a wide range of conditions has important consequences for interpreting empirical results.
 
@@
 
-### *Detecting links between fitness and responses*
+One of the major benefits of simulating glucocorticoid response curves will be the ability to design optimal sampling strategies before data are collected.
+A simulation can be constrained to match any real world limitations (e.g., maximum number of samples possible per individual) and then explored to determine how to best allocate sampling resources.
+The specifics of this task will vary considerably with the study system and question being addressed, but here I illustrate one possible application.
+Consider an experiment in which the acute glucocorticoid response of a treatment group and control group are compared after some experimental manipulation.
+The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group (figure \@ref(fig:prediction-ex)).
+
+For constraints, we can only sample a maximum of 20 individuals per group, we can only sample each individual once post-treatment, and during that single sampling event we can take a blood sample at a maximum of two different time points, resulting in a total of 40 data points per treatment group.
+Given these constraints, we can design a simulation to ask how well several different sampling schemes will perform in estimating the shape of the response curve for each group.
+I compare three different sampling designs: i) a study in which every animal is sampled at 1 minute and at 30 minutes, ii) a study in which two sampling times between 1 and 60 minutes are randomly chosen for every animal, iii) a study in which sampling times are randomly chosen for each animal, but weighted more heavily around 30 minutes, which is the expected average maximum glucocorticoid time.
+In the random sampling schemes, I constrain the randomization so that two samples from the same animal must be taken at least 5 minutes apart.
+To evaluate the performance of these sampling schemes, I simulated each one 50 times and fit a generalized additive model for each simulation with glucocorticoid level as the response and a treatment by smoothed time interaction as a predictor.
+Using the fit models, I asked how well each scheme estimated the time of maximum glucocorticoid and also how well the entire shape of the glucocorticoid response between 0 and 60 minutes was estimated.
+
+```{r prediction-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "caption"}
+
+#plot(1,1)
+
+```
+
+```{r treatment-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "caption"}
+
+```
+
+This simulation is particular to the specific scenario, but a similar scenario could be designed for any number of studies and any number of predictions about how the speed, scope, or other attributes of the glucocorticoid response are expected to change with a treatment.
+Clearly, when estimating the timing of peak glucocorticoids, a simple baseline plus induced sampling scheme is sub optimal, but this scheme may be perform well in other situations where the maximum value is the target and there is relatively little variation in response time.
+Creating simulations like this before studies are conducted has the potential increase the efficient use of researches time and funds, but also forces researches to think explicitly about quantitative predictions ahead of time.
+These simulations could be included as part of a study pre-registration, grant application, or registered report to demonstrate exactly what data collection and analysis approaches are planned and to justify those decisions.
+
+**Still working on thiis section. Will have two treatment groups with different response shapes and evaluate how well different sampling schemes recover those shapes.**
+
+### Repeatability?
+
+The other example I had thought about including was something showing the repeatability functions and how repeatability of different measures might differ, but I'm leaning towards not including this since there is already a lot here.
 
@@
 
+More focus on measuring characteristics of the response other than just base/stress for all individuals. Group comparisons with treatments could look at shape changes.
+
 Based on the brief scenarios developed above:
+
+Goal is for this to be a pretty short section highlighting some main insights from above examples and other ways simulations could be used.
 
@@
 I would like to thank John Wingfield for discussions about the ideas presented in this project and for providing an insightful overview of the history of study design in field based studies of acute stress responses.
-I also thank Maren Vitousek and all members of the Vitousek Lab for feedback and discussion on early versions of this project.
+I also thank members of the Vitousek Lab for feedback and discussion on early versions of this project.
 

### 2021-07-27 3_r_scripts/2_cortResponse_script.R

@@ -1,698 +1,9 @@
 ### Script for speed vs. scope manuscript
 # Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
 
-
 ## Load packages ----
-    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra, sp, rptR)
+    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra, rptR, simcoRt, sp)
 
-## Create simulation function ----
-  # The first function simulates cort response parameters for a population of animals of size n. We
-    # can think of these as the 'true' phenotypes of these individuals. For purposes of this simulation
-    # the cort response is described with five simulated parameters that can covary or not.
-    # The parameters estimated are: 
-          #base cort
-          #time to reach max cort
-          #max cort
-          #time at max cort level
-          #time of return to base cort level 
-          cort_sim1 <- function(n = 20, 
-                               base_mu = 5, base_sd = 2, base_min = 0.5,
-                               slope_mu = 15, slope_sd = 2,
-                               fastpct_mu = 0.55, fastpct_sd = 0.08, 
-                               speed_mu = 25, speed_sd = 5, speed_min = 10,
-                               max_mu = log(45), max_sd = 0.21, 
-                               maxtime_mu = 10, maxtime_sd = 3, maxtime_min = 5,
-                               return_mu = 90, return_sd = 15, 
-                               extra_n = 500,
-                               cor_base_speed = 0, cor_base_max = 0, cor_base_maxtime = 0, cor_base_return = 0, cor_base_slope = 0, cor_base_fastpct = 0,
-                               cor_speed_max = 0, cor_speed_maxtime = 0, cor_speed_return = 0, cor_speed_slope = 0, cor_speed_fastpct = 0,
-                               cor_max_maxtime = 0, cor_max_return = 0, cor_max_slope = 0, cor_max_fastpct = 0,
-                               cor_maxtime_return = 0, cor_maxtime_slope = 0, cor_maxtime_fastpct = 0,
-                               cor_return_slope = 0, cor_return_fastpct = 0,
-                               cor_slope_fastpct = 0)
-              {
-                  # Build a full variance-covariance matrix based on the values for each parameter and correlation between parameters
-                      cort_vcov <- matrix(nrow = 7, ncol = 7)
[…]

### 2021-09-03 3_r_scripts/Manuscript.Rmd
Changed the data simulation example from blackbirds to rats.

@@
 output: 
+  bookdown::pdf_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependencies: ["flafter"]
   bookdown::word_document2:
@@
       extra_dependcies: ["flafter"]
-  bookdown::pdf_document2:
-      number_sections: FALSE
-      toc: FALSE
-      extra_dependencies: ["flafter"]
 bibliography: references.bib
@@
 Using this simulation, I explore several possible scenarios to highlight areas where simulation might either provide new insight into physiological flexibility directly or aid in designing empirical studies that are better able to test the hypotheses of interest.
-The package includes a vignette with details on use and code to reproduce all of the examples explored in this paper.
+All of the code required to reproduce the examples presented here is included in a public GitHub repository.
 
@@
 The simulation is constructed as two main functions with several minor functions for downstream analysis.
-Detailed descriptions of the arguments to each function and a full vignette are included with the package documentation.
+Detailed descriptions of the arguments to each function are included with the package documentation.
 Briefly, function `cort_sim1` samples the parameters shown in Figure \@ref(fig:concept-fig) from an arbitrary number of animals.
@@
 Within each scenario, I have illustrated how the simulation functions might be used to address the particular question of interest, but I have not fully explored all the possible permutations of parameters systematically, because these will depend to a large extent on the empirical details of the system being studied.
+A complete set of reproducible code to create all of the examples presented in this paper is available on GitHub (https://github.com/cct663/speed_vs_scope).
 
@@
 
-Here, I compare simulated data to an empirical dataset from red-winged blackbirds (*Agelaius phoeniceus*) to demonstrate how `simcoRt` can create synthetic datasets similar to observed data.
-This dataset is included in the `simcoRt` package and is fully described in the documentation.
-Briefly, acute glucocorticoid responses were measured using a 7-point sampling scheme, with corticosterone measured at 1, 3, 7, 15, 30, 45, and 60 minutes post capture.
-I simulated data using the functions described above starting with the input values calculated directly from the empirical data.
-The simulation creates a new dataset that has similar variation and patterns to empirical data (Figure \@ref(fig:rwbb-demo)A) along with a population wide corticosterone response curve shape that closely matches the empirical data (Figure \@ref(fig:rwbb-demo)B).
-In this case, the plotted simulation data include the same number of animals sampled at the same time points as the empirical data, but these sampling points and total sample size can easily be changed as desired.
-The parameterized simulation can now be used to test the sensitivity of any number of experimental designs before additional data is collected.
+```{r kool-demo, echo = FALSE, fig.height = 3.5, fig.width = 8, message = FALSE, warning = FALSE, fig.cap = "Panel A shows the acute corticosterone response for measured (blue) or simulated (orange) rats measured at five time points.
+Panel B shows the mean and standard error of the two datasets.
+Empirical data are extracted from Koolhaas et al.
+2010 Figure 6 using the WebPlotDigitizer tool."}
+     # Load and set up data from Koolhaas et al 2010
+#   kool <- read.delim(here::here("1_raw_data/koolhaas_2010_figure6_data.txt"))
+#   kool <- subset(kool, kool$time > -2)
+#   kool$type <- "Koolhaas et al. 2010"
+# 
+# # Create simulated dataset with similar characteristics  
+#     set.seed(175)
+#           sim_kool <- cort_sim2(cort_sim1(
+#            n = 14,
+#            base_mu = 11.6,
+#            base_sd = 3.65,
+#            slope_mu = 3.5,
+#            slope_sd = 0.9,
+#            fastpct_mu = 0.9,
+#            fastpct_sd = 0.1,
+#            max_mu = log(24),
+#            max_sd = 0.1,
+#            speed_mu = 12,
+#            speed_sd = 2.5,
+#            return_mu = 42,
+#            maxtime_mu = 1,
+#            maxtime_sd = 0.5
+#          ), 
+#          sample_times = 1,
+#          bleed_times = c(0, 1, 5, 15, 30, 60),
+#          assay_error = 0.1,
+#          sm_span = 0.2)  
+#       
+# # Combine simulated and real data
+#     sim_kool$simulated_dataset_long$type <- "simulated"
+#      sim_kool2 <- sim_kool$simulated_dataset_long[, c("animal_sample", "time", "cort", "type")]
+#      kool2 <- sim_kool2[, c("animal_sample", "time", "cort")]
+#      kool2$type <- "simulated"
+#      colnames(kool2) <- c("animal_sample", "time", "cort", "type")
+#      
+#      kool <- kool[, c("animal", "time", "cort", "type")]
+#      colnames(kool)[1] <- "animal_sample"
+#      comp_kool <- rbind(kool, kool2)
 
-```{r rwbb-demo, echo = FALSE, fig.height = 3.5, fig.width = 8, message = FALSE, warning = FALSE, fig.cap = "Panel A shows the acute corticosterone response for measured (blue) or simulated (orange) red-winged blackbirds measured at seven time points.
-Panel B shows the mean and standard error of the two datasets."}
-      rwde <- subset(rwbb, rwbb$lh_stg == "early-breeding")  
-      rwde <- subset(rwde, rwde$id != "452")   # This one has an unusually high cort response
-    set.seed(165)
-        sim_rwd <- cort_sim2(cort_sim1(
-       n = 23,
-       base_mu = 3.5,
-       base_sd = 3.5,
-       slope_mu = 17,
-       slope_sd = 0.7,
-       fastpct_mu = 0.7,
-       fastpct_sd = 0.03,
-       max_mu = log(56),
-       max_sd = 0.5,
-       speed_mu = 65,
-       speed_sd = 5,
-       return_mu = 120,
-       maxtime_mu = 25,
-       cor_max_slope = -0.7
-     ), 
-     sample_times = 1,
-     bleed_times = c(1, 3, 8, 15, 30, 45, 60),
-     assay_error = 0.2,
-     sm_span = 0.2)
+# save the data to file and load back in
+     #saveRDS(comp_kool, here::here("5_other_outputs/comp_kool.rds"))
+     comp_kool <- readRDS(here::here("5_other_outputs/comp_kool.rds"))
 
-     sim_rwd$simulated_dataset_long$type <- "simulated"
-     sim_rwd2 <- sim_rwd$simulated_dataset_long[, c("animal_sample", "time", "cort", "type")]
-     rwde2 <- rwde[, c("id", "lat_sec", "cort")]
-     rwde2$lat_sec <- rwde2$lat_sec / 60
-     rwde2$type <- "red-winged blackbird"
-     colnames(rwde2) <- c("animal_sample", "time", "cort", "type")
-
-     comp_sim <- rbind(rwde2, sim_rwd2)
-
-     p1 <- ggplot(comp_sim, mapping = aes(x = time, y = cort, by = animal_sample, color = type)) +
+# Plot the data
+    p1 <- ggplot(comp_kool, mapping = aes(x = time, y = cort, by = animal_sample, color = type)) +
        geom_line(alpha = 0.7, size = 0.7) +
        theme_classic() +
-       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
-       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 45)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (\U00B5g/dl)") +
        scale_color_manual(values = c("slateblue", "orange")) +
-       guides(color = FALSE) +
+       guides(color = "none") +
        theme(axis.title = element_text(size = 14)) +
@@
 
-     p2 <- ggplot(comp_sim, mapping = aes(x = time, y = cort, color = type, fill = type)) +
+     p2 <- ggplot(comp_kool, mapping = aes(x = time, y = cort, color = type, fill = type)) +
        geom_smooth() +
        theme_classic() + 
-       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
-       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 45)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (\U00B5g/dl)") +
        scale_color_manual(values = c("slateblue", "orange")) +
@@
 ```
+
+To demonstrate this utility, I have redrawn data from Koolhaas et al.
+[-@koolhaas2010].
+As part of that study, a series of corticosterone measurements were collected during and after an acute stressor from 14 laboratory rats *Rattus norvegicus* using permanently implanted jugular vein canulae.
+I next simulated data using the functions described above starting with the input values calculated directly from the empirical data.
+The simulation creates a new dataset that has similar variation and patterns to the empirical data (Figure \@ref(fig:kool-demo)A) along with a population wide corticosterone response curve shape that closely matches the empirical data (Figure \@ref(fig:kool-demo)B).
+In this case, the plotted simulation data include the same number of animals sampled at the same time points as the empirical data, but these sampling points and total sample size can easily be changed as desired.
+The parameterized simulation can now be used to test the sensitivity of any number of experimental designs before additional data is collected.
 
@@
 
-Using these three simulated populations as a starting point, I asked how well glucocorticoid values measured at one timepoint reflected true trait values.
-For each population I set an average population level speed of 30 minutes with other values in the simulation set at their default value.
-For every time point from 0 to 35 minutes I fit two simple linear regressions of the measured value on the true speed and maximum value and extracted the R^2^ value from the model.
-I repeated this simulation for all populations 50 times with 100 individuals sampled from the population each time.
-Finally, I repeated the entire set of simulations with each combination of low and high between-individual variation in the speed or maximum values (variation in speed: low = 2 minute SD, high = 12 minute SD; variation in maximum: low = 1ng/$\mu$l SD, high = 10ng/$\mu$l SD).
-
 ```{r cor-ex2, echo = FALSE, fig.height = 6, fig.width = 7, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Relationship between single time point measures of glucocorticoids and the true value of either maximum level (solid lines) or the speed of the glucocorticoid response (dashed lines).
@@
 ```
+
+Using these three simulated populations as a starting point, I asked how well glucocorticoid values measured at one timepoint reflected true trait values.
+For each population I set an average population level speed of 30 minutes with other values in the simulation set at their default value.
+For every time point from 0 to 35 minutes I fit two simple linear regressions of the measured value on the true speed and maximum value and extracted the R^2^ value from the model.
+I repeated this simulation for all populations 50 times with 100 individuals sampled from the population each time.
+Finally, I repeated the entire set of simulations with each combination of low and high between-individual variation in the speed or maximum values (variation in speed: low = 2 minute SD, high = 12 minute SD; variation in maximum: low = 1ng/$\mu$l SD, high = 10ng/$\mu$l SD).
 
@@
 Several patterns can be identified by examining the results of this simulation.
-First, the correlation between the true maximum glucorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex) A-D).
+First, the correlation between the true maximum glucocorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex) A-D).
 In all cases, however, the observed correlation is always lower than the true correlation and always lowest in the population with low between-individual variation.
-The ubquity of this pattern is a product of the simulation structure, because adding measurement error or within-individual variation effecively add noise to the true correlation.
+The ubiquity of this pattern is a product of the simulation structure, because adding measurement error or within-individual variation effectively adds noise to the true correlation.
 It is important to note that in the real world, it is unlikely that this pattern would be so universal, because unmeasured variables could influence both fitness and glucocorticoids.
@@
 The fact that low between-individual variation in maximum glucocorticoids makes it harder to detect true glucocorticoid-fitness relationships across a wide range of conditions has important consequences for interpreting empirical results.
+Many studies have demonstrated different relationships (or lack thereof) between corticosterone and fitness at different life history stages [add our FEE & Bonier papers CITE], but it is also well known that the absolute amount of between individual variation in glucocorticoid traits varies considerably at different stages [CITE].
+Our simulation demonstrates that the power to detect true relationships will differ drastically across these conditions even with similar study designs and samples sizes, suggesting that great care is needed to conclusively demonstrate different glucocorticoid-fitness relationships across contexts.
 
@@
 Consider an experiment in which the acute glucocorticoid response of a treatment group and control group are compared after some experimental manipulation.
-The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group (figure \@ref(fig:prediction-ex)).
+The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group, but will not differ in the maximum value itself (figure \@ref(fig:prediction-ex)).
 
@@
 
-#plot(1,1)
+set.seed(295)
+
+# Simulate group A
+        sim_grpa <- cort_sim2(cort_sim1(
+       n = 20,
+       slope_mu = 9,
+       fastpct_mu = 0.7,
+       max_sd = 0.14,
+       speed_mu = 20,
+       speed_sd = 4
+     ), 
+     sample_times = 1,
+     sm_span = 0.2,
+     bleed_times = c(2, 30))
+
+# Simulate group B
+        sim_grpb <- cort_sim2(cort_sim1(
+       n = 20,
+       slope_mu = 11,
+       fastpct_mu = 0.55,
+       max_sd = 0.14,
+       speed_mu = 32,
+       speed_sd = 4
+     ), 
+     sample_times = 1,
+     sm_span = 0.2,
+     bleed_times = c(2, 30))
+
+# Combine the two datasets
+      sim_grpa$timecourse_long$type <- "Treatment"
+      sim_a2 <- sim_grpa$timecourse_long[, c("animal_sample", "time", "cort", "type")]
+      sim_grpb$timecourse_long$type <- "Control"
+      sim_b2 <- sim_grpb$timecourse_long[, c("animal_sample", "time", "cort", "type")]
+
+      comp_sim2 <- rbind(sim_a2, sim_b2)
+
+     p1 <- ggplot(comp_sim2, mapping = aes(x = time, y = cort, by = animal_sample, color = type)) +
+       geom_line(alpha = 0.3, size = 0.7) +
+       theme_classic() +
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 75)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       scale_color_manual(values = c("slateblue", "orange")) +
+       #guides(color = "none") +
+       theme(axis.title = element_text(size = 14)) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A") +
+       geom_smooth(se = FALSE, size = 1.5, mapping = aes(by = type, color = type)) +
+       theme(legend.title = element_blank())
+
+      sim_grpa$true_values$type <- "Treatment"
+      sim_a22 <- sim_grpa$true_values[, c("baseline", "slope", "fastpct", "speed", "max", "atmax", "type")]
+      sim_grpb$true_values$type <- "Control"
+      sim_b22 <- sim_grpb$true_values[, c("baseline", "slope", "fastpct", "speed", "max", "atmax", "type")]
+
+      comp_sim3 <- rbind(sim_a22, sim_b22)
+      comp_sim4 <- pivot_longer(comp_sim3, cols = c("baseline", "slope", "fastpct", "speed", "max", "atmax"), names_to = "trait", values_to = "value")
+
+      ggplot(comp_sim4, mapping = aes(x = type, y = value, color = type, fill = type)) +
+        geom_boxplot(alpha = 0.5) +
+        scale_color_manual(values = c("slateblue", "orange")) +
+       scale_fill_manual(values = c("slateblue", "orange")) +
+        facet_wrap(~ trait, nrow = 1, scales = "free") +
+        xlab("") +
+        theme_bw() +
+        theme(axis.text.x = element_blank())
+
+     p2 <- ggplot(comp_sim2, mapping = aes(x = time, y = cort, color = type, fill = type)) +
+       geom_smooth() +
+       theme_classic() + 
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
+       xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
+       scale_color_manual(values = c("slateblue", "orange")) +
+       scale_fill_manual(values = c("slateblue", "orange")) +
+       theme(legend.position = c(0.7, 0.15), axis.title = element_text(size = 14)) +
+       guides(color = guide_legend(title = ""), fill = guide_legend(title = "")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B")
+
+      ggpubr::ggarrange(p1, p2, nrow = 1)
 
@@
 
-**Still working on thiis section. Will have two treatment groups with different response shapes and evaluate how well different sampling schemes recover those shapes.**
-
-### Repeatability?
-
-The other example I had thought about including was something showing the repeatability functions and how repeatability of different measures might differ, but I'm leaning towards not including this since there is already a lot here.
-
 ## DISCUSSION
@@
 Empirical data is so limited and hard to collect, that it makes sense to proceed with simulations to guide and focus empirical work and to determine whether empirical study designs actually have the ability in principle to support hypotheses of interest given reasonable effect sizes.
+
+How to interpret some of the empirical studies that are published in light of simulations?
 

### 2021-09-03 3_r_scripts/Manuscript.Rmd
Updated study design and conclusion sections.

@@
 output: 
+  bookdown::word_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependcies: ["flafter"]
   bookdown::pdf_document2:
@@
       extra_dependencies: ["flafter"]
-  bookdown::word_document2:
-      number_sections: FALSE
-      toc: FALSE
-      extra_dependcies: ["flafter"]
 bibliography: references.bib
@@
 Consider an experiment in which the acute glucocorticoid response of a treatment group and control group are compared after some experimental manipulation.
-The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group, but will not differ in the maximum value itself (figure \@ref(fig:prediction-ex)).
-
-For constraints, we can only sample a maximum of 20 individuals per group, we can only sample each individual once post-treatment, and during that single sampling event we can take a blood sample at a maximum of two different time points, resulting in a total of 40 data points per treatment group.
-Given these constraints, we can design a simulation to ask how well several different sampling schemes will perform in estimating the shape of the response curve for each group.
-I compare three different sampling designs: i) a study in which every animal is sampled at 1 minute and at 30 minutes, ii) a study in which two sampling times between 1 and 60 minutes are randomly chosen for every animal, iii) a study in which sampling times are randomly chosen for each animal, but weighted more heavily around 30 minutes, which is the expected average maximum glucocorticoid time.
-In the random sampling schemes, I constrain the randomization so that two samples from the same animal must be taken at least 5 minutes apart.
-To evaluate the performance of these sampling schemes, I simulated each one 50 times and fit a generalized additive model for each simulation with glucocorticoid level as the response and a treatment by smoothed time interaction as a predictor.
-Using the fit models, I asked how well each scheme estimated the time of maximum glucocorticoid and also how well the entire shape of the glucocorticoid response between 0 and 60 minutes was estimated.
-
-```{r prediction-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "caption"}
-
-set.seed(295)
-
-# Simulate group A
-        sim_grpa <- cort_sim2(cort_sim1(
-       n = 20,
-       slope_mu = 9,
-       fastpct_mu = 0.7,
-       max_sd = 0.14,
-       speed_mu = 20,
-       speed_sd = 4
-     ), 
-     sample_times = 1,
-     sm_span = 0.2,
-     bleed_times = c(2, 30))
-
-# Simulate group B
-        sim_grpb <- cort_sim2(cort_sim1(
-       n = 20,
-       slope_mu = 11,
-       fastpct_mu = 0.55,
-       max_sd = 0.14,
-       speed_mu = 32,
-       speed_sd = 4
-     ), 
-     sample_times = 1,
-     sm_span = 0.2,
-     bleed_times = c(2, 30))
-
-# Combine the two datasets
-      sim_grpa$timecourse_long$type <- "Treatment"
-      sim_a2 <- sim_grpa$timecourse_long[, c("animal_sample", "time", "cort", "type")]
-      sim_grpb$timecourse_long$type <- "Control"
-      sim_b2 <- sim_grpb$timecourse_long[, c("animal_sample", "time", "cort", "type")]
-
-      comp_sim2 <- rbind(sim_a2, sim_b2)
+The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group, but will not differ in the maximum value itself.
+I have implemented this difference by simulating two populations in which the treatment group has a steeper initial slope and also reaches the maximum value faster (figure \@ref(fig:prediction-ex)).
+
+```{r prediction-ex, echo = FALSE, fig.height = 3.5, fig.width = 5, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Simulated data for a hypothetical control (blue) and treatment (orange) group.
+Faded thin lines show the acute response for each individual simulated (20 per group) and thick lines show the average response curve for each group."}
+
+# set.seed(295)
+# 
+# # Simulate group A
+#         sim_grpa <- cort_sim2(cort_sim1(
+#        n = 20,
+#        slope_mu = 9,
+#        fastpct_mu = 0.7,
+#        max_sd = 0.14,
+#        speed_mu = 20,
+#        speed_sd = 4
+#      ),
+#      sample_times = 1,
+#      sm_span = 0.2,
+#      bleed_times = seq(1, 60, 1))
+# 
+# # Simulate group B
+#         sim_grpb <- cort_sim2(cort_sim1(
+#        n = 20,
+#        slope_mu = 11,
+#        fastpct_mu = 0.55,
+#        max_sd = 0.14,
+#        speed_mu = 32,
+#        speed_sd = 4
+#      ),
+#      sample_times = 1,
+#      sm_span = 0.2,
+#      bleed_times = seq(1, 60, 1))
+# 
+# # Combine the two datasets
+#       sim_grpa$timecourse_long$type <- "Treatment"
+#       sim_a2 <- sim_grpa$timecourse_long[, c("animal_sample", "time", "cort", "type")]
+#       sim_grpb$timecourse_long$type <- "Control"
+#       sim_b2 <- sim_grpb$timecourse_long[, c("animal_sample", "time", "cort", "type")]
+# 
+#       comp_sim2 <- rbind(sim_a2, sim_b2)
+
+# save group and read it in
+      #saveRDS(comp_sim2, here::here("5_other_outputs/comp_sim2.rds"))
+      comp_sim2 <- readRDS(here::here("5_other_outputs/comp_sim2.rds"))
 
@@
 
-      sim_grpa$true_values$type <- "Treatment"
-      sim_a22 <- sim_grpa$true_values[, c("baseline", "slope", "fastpct", "speed", "max", "atmax", "type")]
-      sim_grpb$true_values$type <- "Control"
-      sim_b22 <- sim_grpb$true_values[, c("baseline", "slope", "fastpct", "speed", "max", "atmax", "type")]
-
-      comp_sim3 <- rbind(sim_a22, sim_b22)
-      comp_sim4 <- pivot_longer(comp_sim3, cols = c("baseline", "slope", "fastpct", "speed", "max", "atmax"), names_to = "trait", values_to = "value")
-
-      ggplot(comp_sim4, mapping = aes(x = type, y = value, color = type, fill = type)) +
-        geom_boxplot(alpha = 0.5) +
-        scale_color_manual(values = c("slateblue", "orange")) +
-       scale_fill_manual(values = c("slateblue", "orange")) +
-        facet_wrap(~ trait, nrow = 1, scales = "free") +
-        xlab("") +
-        theme_bw() +
-        theme(axis.text.x = element_blank())
-
-     p2 <- ggplot(comp_sim2, mapping = aes(x = time, y = cort, color = type, fill = type)) +
-       geom_smooth() +
+     p1
+
+```
+
+Next, we can ask how well different study designs can detect this difference.
+Here we can impose any logistical constraints relevant to the study system.
+As an example, in this case we can only sample a maximum of 20 individuals per group, we can only sample each individual once post-treatment, and during that single sampling event we can take a blood sample at a maximum of two different time points, resulting in a total of 80 data points.
+Given these constraints, I compare three different sampling designs: i) a study in which every animal is sampled at 1 minute, 30 minutes, and 60 minutes, ii) a study in which two sampling times between 1 and 60 minutes are randomly chosen for every animal, iii) a study in which two sampling times are randomly chosen for each animal, but weighted more heavily around the range of times when maximum levels are expected to be reached for the population.
+Note that the first sampling scheme closely mirrors the most common empirical design and in this case I have allowed an extra, third sample at 60 minutes, such that it includes 120, rather than 80, data points.
+For illustration purposes I sampled directly from the 'true' response curves in this example so that there is no additional measurement error added.
+To evaluate these schemes I compare estimates of the acute response curve for each group to the 'true' known curves shown in figure \@ref(fig:prediction-ex).
+Note that a more complete analysis of a sampling schemes performance should include many more iterations and full statistical comparisons, but the details here will be highly dependent on the study system and goals, so I provide this simple example to illustrate the approach rather than to make any more widely applicable conclusions.
+
+```{r treatment-ex, echo = FALSE, fig.height = 2.8, fig.width = 7.3, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Three possible sampling schemes to compare two groups.
+For standard sampling (A), every individual is sampled at exactly 2, 30, and 60 minutes.
+For random sampling (B) each individual is sampled at two random points between 1 and 60 minutes.
+For weighted sampling (C) two sampling times are chosen for each individual from a normal distribution with mean of 32 and sd of 9 minutes.
+In all three panels, solid lines are the true group averages, dashed lines are the estimates based on samples, and points are individual samples collected."}
+
+# save group and read it in
+    set.seed(61)
+      #saveRDS(comp_sim2, here::here("5_other_outputs/comp_sim2.rds"))
+      comp_sim2 <- readRDS(here::here("5_other_outputs/comp_sim2.rds"))
+
+     p1 <- ggplot(comp_sim2, mapping = aes(x = time, y = cort, by = animal_sample, color = type)) +
        theme_classic() +
-       coord_cartesian(xlim = c(0, 60), ylim = c(0, 125)) +
+       coord_cartesian(xlim = c(0, 60), ylim = c(0, 75)) +
        xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
        scale_color_manual(values = c("slateblue", "orange")) +
+       guides(color = "none") +
+       theme(axis.title = element_text(size = 14)) +
+       geom_smooth(se = FALSE, size = 1.1, mapping = aes(by = type, color = type)) +
+       theme(legend.title = element_blank())
+
+# sampling scheme 1
+     dd1 <- subset(comp_sim2, comp_sim2$time == 1 | comp_sim2$time == 30 | comp_sim2$time == 60)
+
+     p1a <- p1 +
+       #geom_line(data = dd1, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
+       geom_smooth(data = dd1, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed") +
        scale_fill_manual(values = c("slateblue", "orange")) +
-       theme(legend.position = c(0.7, 0.15), axis.title = element_text(size = 14)) +
-       guides(color = guide_legend(title = ""), fill = guide_legend(title = "")) +
-       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B")
-
-      ggpubr::ggarrange(p1, p2, nrow = 1)
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A") +
+       guides(fill = "none") +
+       geom_point(data = dd1, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
+       ggtitle("Standard sampling")
+
+# sampling scheme 2
+     dd2 <- data.frame(animal_sample = NA, time = NA, cort = NA, type = NA, an2 = NA)
+     comp_sim2$an2 <- paste(comp_sim2$animal_sample, comp_sim2$type, sep = "_")
+     ans <- data.frame(animal_sample = unique(comp_sim2$an2))
+     cs2 <- subset(comp_sim2, comp_sim2$time > 0 & comp_sim2$time < 61)
+     for(i in 1:nrow(ans)){
+       sub <- subset(cs2, cs2$an2 == ans$animal_sample[i])
+       subs <- sub[sample(seq(1, 60, 1), 2, replace = FALSE), ]
+       dd2 <- rbind(dd2, subs)
+     }
+     dd2 <- subset(dd2, is.na(dd2$an2) == FALSE)
+
+     p1b <- p1 +
+       #geom_line(data = dd2, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
+       geom_smooth(data = dd2, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed") +
+       scale_fill_manual(values = c("slateblue", "orange")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B") +
+       guides(fill = "none") +
+       ylab("") +
+       geom_point(data = dd2, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
+       ggtitle("Random sampling")
+
+# sampling scheme 3
+     dd3 <- data.frame(animal_sample = NA, time = NA, cort = NA, type = NA, an2 = NA)
+     comp_sim2$an2 <- paste(comp_sim2$animal_sample, comp_sim2$type, sep = "_")
+     ans <- data.frame(animal_sample = unique(comp_sim2$an2))
+     cs2 <- subset(comp_sim2, comp_sim2$time > 0 & comp_sim2$time < 61)
+     for(i in 1:nrow(ans)){
+       choose <- round(rnorm(2, 32, 9))
+       sub <- subset(cs2, cs2$an2 == ans$animal_sample[i])
+       subs <- sub[choose, ]
+       dd3 <- rbind(dd3, subs)
+     }
+     dd3 <- subset(dd3, is.na(dd3$an2) == FALSE)
+
+     p1c <- p1 +
+       #geom_line(data = dd3, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
+       geom_smooth(data = dd3, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed", span = 0.9) +
+       scale_fill_manual(values = c("slateblue", "orange")) +
+       scale_color_manual(values = c("slateblue", "orange")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C") +
+       #theme(legend.position = c(.6, .2)) +
+       guides(fill = "none") +
+       ylab("") +
+       geom_point(data = dd3, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
+       ggtitle("Weighted sampling")
+
+     ggpubr::ggarrange(p1a, p1b, p1c, nrow = 1)
 
@@
 
-```{r treatment-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "caption"}
-
-```
-
-This simulation is particular to the specific scenario, but a similar scenario could be designed for any number of studies and any number of predictions about how the speed, scope, or other attributes of the glucocorticoid response are expected to change with a treatment.
+In this case, the standard sampling scheme performs poorly (figure \@ref(fig:treatment-ex) A), with no differences detectable between the two groups, despite the fact that the treatment group reaches it's maximum value on average 12 minutes (~40%) faster than the control group.
+In contrast, both the random sampling and weighted sampling schemes detect differences in the shape of the acute response (figure \@ref(fig:treatment-ex) B & C).
+In this particular scenario, there is no clear difference between these two approaches.
+A few clear takeaways can be derived from these results.
+First, while strict standardization of the timing of samples has some clear advantages, it also comes with costs and likely makes it nearly impossible to detect certain types of variation between groups or individuals.
+In this case, standardized sampling performed much worse than the other two approaches despite the fact that the analysis included 50% more data and it should be clear that no amount of additional sampling would allow that approach to detect the differences between these groups.
+Second, while it may be very difficult to accurately estimate the full shape of the acute stress response for *individuals*, the sampling schemes shown here demonstrate that it should be possible to describe these shapes accurately for groups (e.g., treatments, species, different contexts) even without extraordinarily large sample sizes.
+A similar argument about the power of randomly timed sampling has been put forward in the function valued trait literature [CITE], but this type of sampling scheme is rarely used in evolutionary endocrinology research.
+
+This simulation is particular to a single very specific scenario, but a similar scenario could be designed for any number of studies and any number of predictions about how the speed, scope, or other attributes of the glucocorticoid response are expected to change with a treatment or between different groups or species.
 Clearly, when estimating the timing of peak glucocorticoids, a simple baseline plus induced sampling scheme is sub optimal, but this scheme may be perform well in other situations where the maximum value is the target and there is relatively little variation in response time.
@@
 
-## DISCUSSION
-
-More sophisticated simulations that incorporate mechanistic processes or other molecules or components of stress response system.
-
-Empirical data is so limited and hard to collect, that it makes sense to proceed with simulations to guide and focus empirical work and to determine whether empirical study designs actually have the ability in principle to support hypotheses of interest given reasonable effect sizes.
-
-How to interpret some of the empirical studies that are published in light of simulations?
-
-More focus on measuring characteristics of the response other than just base/stress for all individuals. Group comparisons with treatments could look at shape changes.
-
-Based on the brief scenarios developed above:
-
-Goal is for this to be a pretty short section highlighting some main insights from above examples and other ways simulations could be used.
+## CONCLUSION
+
+While there has been increasing interest in understand within- and between-individual variation in the acute glucocorticoid response in recent years [CITE], the methods and data available to tackle these questions have changed relatively little.
+Many sophisticated statistical tools are now available and clear arguments have been made about the need to apply these approaches to endocrine traits [CITE].
+Arguably, the biggest roadblock at the moment is the limited availability of empirical data needed to test these ideas.
+Simulation offers one way forward, by allowing for more efficiently designed studies and by allowing researches to identify when the question of interest can *in principle* be answered with a given study design.
+Ideally, conceptual papers, empirical work, and simulation will proceed together to make progress in this field.
+The tools presented here only scratch the surface of the ways that data simulation can be applied to address pressing questions in evolutionary endocrinology.
+
+Nevertheless, even the simple demonstrations included in this paper suggest several ways that simulation could help move the field forward.
+One of the main benefits of simulating datasets is identifying unmeasured properties and assumptions of currently available data that can become targets for empirical work.
+For example, I demonstrated that the covariation between different components of the acute stress resopnse and the relative amount of variation in each of these can have profound effects on the ability to accurately measure any single component.
+Empirical work specifically designed to assess covariation and variance at different times could help to understand what conclusions we can reasonably draw from available data.
+One takeaway from these simulations is that variation in glucocorticoid-fitness relationships across seasons or life history stages can easily arise as a statistical artefact when between-individual variation in hormones also varies across the contexts.
+My study design example also suggests that there are potentially gains to be made by considering more diverse sampling designs tailored to the particular research question and study system.
+While standardized sample collection timing has been productive in this field, it also creates clear blind spots to certain types of variation between groups.
+
+In addition to providing insight in its own right, simulation has great potential to hone the design of future empirical studies by allowing for a principled analysis of various study design options and choices before costly data are collected and before animals are needlessly disturbed.
+For example, I showed that one of the most common sampling design schemes has essentially no ability to detect a difference in the speed of increase between two groups if they do not also differ in maximum values.
+It is perhaps not surprising to find that there is little published evidence for differences in the speed of the acute response when most data sampled to date cannot detect those differences.
+Across a wide range of disciplines there has been an increasing push for pre-registration, reproducible research, and transparent research practices [CITE].
+Simulation provides an opportunity for evolutionary endocrinologists to embrace these best practices by improving the quality of study design, allowing for more quantitative hypotheses and predictions, and providing a clear justification for experimental choices.
+
+This package and paper is meant only as a very cursory exploration of the ways that simulation can be applied to evolutionary endocrinology.
+I have no doubt that many more scenarios and complications could be added on to each of the simple examples presented here.
+Furthermore, there is ample room to create more sophisticated simulations that incorporate realistic mechanistic processes or interactions with other molecules and other components of the stress response system.
+I hope that this work will be a starting point to build and improve on as we work to understand the importance of variation in these flexible response systems.
 
@@
 
-I would like to thank John Wingfield for discussions about the ideas presented in this project and for providing an insightful overview of the history of study design in field based studies of acute stress responses.
-I also thank members of the Vitousek Lab for feedback and discussion on early versions of this project.
+I would like to thank Michaela Hau and her lab along with John Wingfield for discussions about the ideas presented in this project.
+John Wingfield also provided an insightful overview of the history of study design in field based studies of acute stress responses.
+Finally, I thank Maren Vitousek and her lab members for feedback and discussion on early versions of this project.
 

### 2022-05-05 3_r_scripts/Manuscript.Rmd
Prepping for revision.

@@
 ---
-title: "Functions for simulating data and designing studies of physiological flexibility in the acute glucocorticoid response to stressors"
+title: "Simulating physiological flexibility in the acute glucocorticoid response to stressors reveals limitations of current empirical approaches"
 author: 
@@
   - \usepackage[font={footnotesize}, labelfont={bf}]{caption}
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
 fontsize: 12pt
 output: 
+  word_document:
+  bookdown::pdf_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependencies: ["flafter"]
   bookdown::word_document2:
@@
 
+*Cornell Lab of Ornithology and Department of Ecology & Evolution*
+
+*cct63@cornell.edu*
+
+\newpage
+\raggedright
+
 ## ABSTRACT
@@
 The glucocorticoid mediated stress response is one of the major systems that allows vertebrates to rapidly adjust their physiology and behavior.
-Given it's role in responding to challenges, evolutionary physiologists have focused on the consequences of between-individual and, more recently, within-individual variation in the acute glucocorticoid response.
-Although sophisticated approaches are available to partition this variation statistically, empirical studies of physiological flexibility are severely limited by the logistical challenges of measuring the same animal multiple times during a single acute response or across multiple instances of acute responses.
-Empiricists have largely adopted the strategy of standardizing sampling as much as possible to allow for comparison between individuals, but this standardization also makes it very difficult to detect certain types of variation in the functional shape of acute response curves.
+Given its role in responding to challenges, evolutionary physiologists have focused on the consequences of between-individual and, more recently, within-individual variation in the acute glucocorticoid response.
+However, empirical studies of physiological flexibility are severely limited by the logistical challenges of measuring the same animal multiple times.
 Data simulation is a powerful approach when empirical data are limited, but has not been adopted to date in studies of physiological flexibility.
-In this paper, I describe the `simcoRt` package, which includes functions that can generate realistic acute glucocorticoid response data with user specified characteristics.
-Simulated animals can be sampled continuously through an acute response and across as many separate responses as desired, while varying key parameters (e.g., the degree of correlation between the speed and scope of a response).
-Using this simulation, I explore several possible scenarios to highlight areas where simulation might either provide new insight into physiological flexibility directly or aid in designing empirical studies that are better able to test the hypotheses of interest.
-All of the code required to reproduce the examples presented here is included in a public GitHub repository.
+In this paper, I develop a simulation that can generate realistic acute glucocorticoid response data with user specified characteristics.
+Simulated animals can be sampled continuously through an acute response and across as many separate responses as desired, while varying key parameters.
+Using the simulation, I develop several scenarios that address key questions in physiological flexibility.
+These scenarios demonstrate the conditions under which a single glucocorticoid trait can be accurately assessed with typical experimental designs, the consequences of covariation between different components of the acute stress response, and the way that context specific differences in variability of acute responses can influence the power to detect relationships between the strength of the acute stress response and fitness.
+I also describe how to use the simulation tools to aid in the design and evaluation of empirical studies.
+Recently there has been a great deal of interest in understanding the causes and consequences of physiological flexibility, but empirical data are often extremely limited and traditional sampling methods are poorly designed to directly address the questions of interest.
+This simulation represents a way forward by revealing critical aspects of physiological flexibility and by creating a tool for designing better empirical studies that integrate simulation and theory with data collection.
+
+*Keywords: acute stress response, physiological flexibility, glucocorticoids, evolutionary endocrinology*
 
@@
 Early studies often employed various control groups and sampled individual animals at a variety of time points over a long period in order to describe the full response curve for a particular group [e.g., a species or a breeding stage, @wingfield1992].
-These validations were considered essential to characterize key parameters of the acute response for each group being studied (i.e., baseline, rate of increase, maximum level, time of peak, and area under the curve; John Wingfield, personal communication).
+These validations were considered essential to characterize key parameters of the acute response for each group being studied (i.e., baseline, rate of increase, maximum level, time of peak, and area under the curve; John Wingfield, *personal communication*).
 The challenge of estimating these parameters becomes much more difficult when trying to describe the response for an individual animal rather than for a group, because glucocorticoids can often only be measured at two or three time points and only a small number of times per animal [e.g., @vitousek2018].
@@
 One of the characteristics of both the within-individual reaction norm and FVT literature is that empirical work has proceeded in very close coordination with simulation and statistical method development.
-In contrast, studies of endocrine flexibility often point to these methods, but don't address the ways that the particular logistical challenges of hormone measurement might suggest altered study designs.
-While many of the tools developed in these fields are transferable, studies of physiological flexibility would benefit from a focus on analysis development and testing that explicitly incorporates the particular details of these questions.
+In contrast, studies of endocrine flexibility often point to these methods, but don't address the ways that the particular logistical challenges of hormone measurement might necessitate different empirical approaches.
+I believe this is one reason that there are currently more conceptual papers arguing for a reaction norm approach to endocrine variation than there are empirical papers actually applying the approach.
+While many of the tools developed in these related fields are transferable, studies of physiological flexibility would benefit from a focus on analysis development and testing that explicitly incorporates the particular details and challenges of these questions.
 One way to accomplish these goals is to use simulations, but to my knowledge no studies of physiological flexibility have developed simulations of the acute stress response that address the issues discussed above.
@@
 
-## DESCRIPTION OF THE SIMULATION
-
-I developed a package called `simcoRt` in R version 4.0.2 [@rcore] with functions to generate acute physiological response curves.
+## MATERIALS AND METHODS
+
+### DESCRIPTION OF THE SIMULATION
+
+I developed a set of functions in R version 4.0.2 [@rcore] to generate acute physiological response curves.
 This simulation makes no assumptions about the  mechanistic process that results in the shape of a glucocorticoid response.
@@
 From this expressed response, individual data points are then collected at user specified times that would reflect an empirical study design (e.g., 2, 30, and 60 minutes).
-Additional noise is added to these data points to represent measurement error (e.g., assay error).
+Additional noise can be added to these data points to represent measurement error (e.g., assay error).
 
@@
 Simple plots illustrating repeated samples from the same individuals are also returned by default.
-
-## EXAMPLE APPLICATIONS OF SIMULATION
+I do not develop an example of repeatability in this manuscript, but the functions here could be used to determine the impact of different study design choices on repeatability estimates.
+
+### RESULTS
 
@@
 
-### *Simulating empirically parameterized data*
+#### *Simulating empirically parameterized data*
 
@@
 
-### *Accurately measuring a single glucocorticoid trait*
+#### *Accurately measuring a single glucocorticoid trait*
 
@@
 For example, the time chosen to take a stress-induced sample is often assumed to be either at the species peak or during a plateau period after the species peak.
-In some early studies, great care was taken to determine an average population level peak time [CITE], but many studies adopt the widely used 'standard' time of 30 minutes post capture without extensive validation.
+In some early studies, great care was taken to determine an average population level peak time [@wingfield1992], but many studies adopt the widely used 'standard' time of 30 minutes post capture without extensive validation [compiled in @vitousek2019].
 While there is a general assumption that sampling later than the peak is acceptable (and perhaps preferable) because animals will be sampled during a relatively stable high plateau, there is little empirical data to evaluate this assertion or to determine how much under or overshooting the species peak timing might influence inferences.
-Furthermore, even when the average peak timing is well established, differences in the amount of between-individual variation in the time to reach the peak or in peak values are common across species and even in different life history stages within species [CITE].
+Furthermore, even when the average peak timing is well established, differences in the amount of between-individual variation in the time to reach the peak or in peak values are common across species and even in different life history stages within species [@wingfield1992].
 The combinations of these patterns of variation could have major consequences on the accuracy of single point estimates taken at 30 minutes, but these questions cannot be addressed directly with empirical datasets where the true underlying values of each individual are unknown.
@@
 In one sense, this result is unsurprising because it is intuitive that large differences would be easier to detect, but there are important consequences of this fact for interpreting studies that seek to link between-individual variation in the magnitude of the stress response with other traits.
-For example, the magnitude of the acute stress response often varies substantially across life history stages [CITE].
+For example, the magnitude of the acute stress response often varies substantially across life history stages [@wingfield1992].
 Even if study designs are identical and maximum glucocorticoids are associated with performance, it will be easier to detect those patterns during life history stages with greater variation (see section on detecting fitness associations below).
 There is a weaker, but still substantial impact of variation in the time taken to reach maximum values on the accuracy of estimates in this simulation.
-Greater variation in time reduces the accuracy of estimates of maximal values.
+Greater variation in the speed of the response reduces the accuracy of estimates of maximal values.
 Finally, the timing of sampling relative to the average population peak timing also influences accuracy.
@@
 
-### *Exploring covariance between response components*
+#### *Exploring covariance between response components*
 
@@
 In these cases, measurements taken at particular time points contain information about multiple aspects of the response and without additional information it may be difficult to know what trait is being measured.
-The fact that each of these traits might be important and that they might covary has been discussed in a general sense [CITE], but simulations are uniquely powerful for exploring under exactly what conditions time point measure of glucocorticoids can or cannot be used as indicators of these traits.
+The fact that each of these traits might be important and that they might covary has been discussed in a general sense [e.g., @baugh2013], but simulations are uniquely powerful for exploring under exactly what conditions time point measure of glucocorticoids can or cannot be used as indicators of these traits.
 
@@
 Alternatively, the speed and maximum values might vary independently (figure \@ref(fig:cor-ex1)B; correlation = 0).
-Finally, individuals that are faster might max out at lower glucocorticoid values (figure \@ref(fig:cor-ex1)C; correlation = 0.6).
+Finally, individuals that are faster responders might max out at lower glucocorticoid values (figure \@ref(fig:cor-ex1)C; correlation = 0.6).
 While many researchers in this field might have intuitions about which of these scenarios is most likely to prevail, there is very little empirical data available to actually determine which is most common.
@@
 First, neither speed or maximum traits could be assessed accurately when between-individual variation in both traits was low (figure \@ref(fig:cor-ex2)A).
-This is potentially important for interpreting apparent differences in glucocorticoid fitness relationships because between-individual variation is known to differ across life history stages [CITE].
-Second, accurately assessing variation in speed was much harder-if not impossible-with single measures.
-
-Only when speed varied between individuals with very low variation in maximum was it possible to accurately measure speed, but this situation may be rare in natural populations.
-When speed was tightly correlated with maximum (figure \@ref(fig:cor-ex2)D) it was sometimes possible to attain reasonable estimates of speed (figure \@ref(fig:cor-ex2)C-D), but when speed was not correlated with maximum single measures were not good indicators of variation in speed (figure \@ref(fig:cor-ex2)A, C-D).
+This is potentially important for interpreting apparent differences in glucocorticoid fitness relationships because between-individual variation is known to differ across life history stages [@wingfield1992].
+Second, accurately assessing variation in speed was much harder---if not impossible---with single measures.
+
+It was only possible to accurately estimate speed when high between-individual variation in speed was coupled with low variation in maximal values, but this situation may be rare in natural populations.
+When speed was tightly correlated with maximum (figure \@ref(fig:cor-ex2)D) it was sometimes possible to attain reasonable estimates of speed (figure \@ref(fig:cor-ex2)C-D), but when speed was not correlated with maximum, single measures were not good indicators of variation in speed (figure \@ref(fig:cor-ex2)A, C-D).
 Finally, measuring variation in maximum values was much easier under many conditions (figure \@ref(fig:cor-ex2)C-D), but the accuracy of assessment of maximum values was also negative impacted by variation in speed and the degree of this impact differed depending on the correlation between the two traits (figure \@ref(fig:cor-ex2)).
@@
 
-### *Detecting links between fitness and responses*
+#### *Detecting links between fitness and responses*
 
 A common goal of recent studies is to establish whether variation in glucocorticoids is associated with fitness or some proxy for fitness [@schoenle2020].
-While there has been a great deal of discussion about the extent to which these relationships might differ with life history characteristics or between breeding stages [CITE], there has been relatively little consideration of the way that methodological limitations might limit the ability to detect these relationships even when they exist.
+While there has been a great deal of discussion about the extent to which these relationships might differ with life history characteristics or between breeding stages, there has been relatively little consideration of the way that methodological limitations might limit the ability to detect these relationships even when they exist.
 
@@
 Several patterns can be identified by examining the results of this simulation.
-First, the correlation between the true maximum glucocorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex) A-D).
-In all cases, however, the observed correlation is always lower than the true correlation and always lowest in the population with low between-individual variation.
+First, the correlation between the true maximum glucocorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex)A-D).
+In all cases, however, the observed correlation is lower than the true correlation and always lowest in the population with low between-individual variation.
 The ubiquity of this pattern is a product of the simulation structure, because adding measurement error or within-individual variation effectively adds noise to the true correlation.
@@
 Nevertheless, general patterns illustrated by the simulation are likely to pertain across a wide range of conditions.
-In this case, it is easiest to detect significant glucocorticoid-fitness relationships when both measurement error and within-individual variation are low (figure \@ref(fig:fitness-ex) A).
-It becomes harder to detect these true relationships when either measurement error (figure \@ref(fig:fitness-ex) B) or within-individual variation (figure \@ref(fig:fitness-ex) C) are high, but even in these more challenging situations the relationship can be detected the majority of the time if between-individual variation in maximum levels is high.
+In this case, it is easiest to detect significant glucocorticoid-fitness relationships when both measurement error and within-individual variation are low (figure \@ref(fig:fitness-ex)A).
+It becomes harder to detect these true relationships when either measurement error (figure \@ref(fig:fitness-ex)B) or within-individual variation (figure \@ref(fig:fitness-ex)C) are high, but even in these more challenging situations the relationship can be detected the majority of the time if between-individual variation in maximum levels is high.
 When both measurement error and within-individual variation are high, it is nearly impossible to detect glucocorticoid-fitness relationships with low-between individual variation, but in populations with high between-individual variation the relationship is still detected in about half of the simulations.
@@
 The fact that low between-individual variation in maximum glucocorticoids makes it harder to detect true glucocorticoid-fitness relationships across a wide range of conditions has important consequences for interpreting empirical results.
-Many studies have demonstrated different relationships (or lack thereof) between corticosterone and fitness at different life history stages [add our FEE & Bonier papers CITE], but it is also well known that the absolute amount of between individual variation in glucocorticoid traits varies considerably at different stages [CITE].
-Our simulation demonstrates that the power to detect true relationships will differ drastically across these conditions even with similar study designs and samples sizes, suggesting that great care is needed to conclusively demonstrate different glucocorticoid-fitness relationships across contexts.
-
-### *Designing optimal sampling strategies*
+Many studies have demonstrated different relationships (or lack thereof) between corticosterone and fitness at different life history stages [@vitousek2018; @bonier2009b], but it is also well known that the absolute amount of between individual variation in glucocorticoid traits varies considerably at different stages [@wingfield1992].
+Our simulation demonstrates that the power to detect true relationships will differ drastically across these conditions even with identical study designs and samples sizes, suggesting that great care is needed to conclusively differentiate true differences in glucocorticoid-fitness relationships across contexts from statistical artefacts.
+
+#### *Designing optimal sampling strategies*
 
@@
 I have implemented this difference by simulating two populations in which the treatment group has a steeper initial slope and also reaches the maximum value faster (figure \@ref(fig:prediction-ex)).
+Any number of possible hypotheses for a particular study system could be specified following a similar approach.
 
@@
 Given these constraints, I compare three different sampling designs: i) a study in which every animal is sampled at 1 minute, 30 minutes, and 60 minutes, ii) a study in which two sampling times between 1 and 60 minutes are randomly chosen for every animal, iii) a study in which two sampling times are randomly chosen for each animal, but weighted more heavily around the range of times when maximum levels are expected to be reached for the population.
+
 Note that the first sampling scheme closely mirrors the most common empirical design and in this case I have allowed an extra, third sample at 60 minutes, such that it includes 120, rather than 80, data points.
@@
 
-In this case, the standard sampling scheme performs poorly (figure \@ref(fig:treatment-ex) A), with no differences detectable between the two groups, despite the fact that the treatment group reaches it's maximum value on average 12 minutes (~40%) faster than the control group.
+In this case, the standard sampling scheme performs very poorly (figure \@ref(fig:treatment-ex)A), with no differences detectable between the two groups, despite the fact that the treatment group reaches it's maximum value on average 12 minutes (~40%) faster than the control group.
 In contrast, both the random sampling and weighted sampling schemes detect differences in the shape of the acute response (figure \@ref(fig:treatment-ex) B & C).
@@
 First, while strict standardization of the timing of samples has some clear advantages, it also comes with costs and likely makes it nearly impossible to detect certain types of variation between groups or individuals.
-In this case, standardized sampling performed much worse than the other two approaches despite the fact that the analysis included 50% more data and it should be clear that no amount of additional sampling would allow that approach to detect the differences between these groups.
+In this case, standardized sampling performed much worse than the other two approaches despite the fact that the analysis included 50% more data; it should be clear that no amount of additional sampling would allow that approach to detect this particular pattern of between group differences.
 Second, while it may be very difficult to accurately estimate the full shape of the acute stress response for *individuals*, the sampling schemes shown here demonstrate that it should be possible to describe these shapes accurately for groups (e.g., treatments, species, different contexts) even without extraordinarily large sample sizes.
-A similar argument about the power of randomly timed sampling has been put forward in the function valued trait literature [CITE], but this type of sampling scheme is rarely used in evolutionary endocrinology research.
+A similar argument about the power of randomly timed sampling has been put forward in the function valued trait literature [@gomulkiewicz2018], but this type of sampling scheme is rarely used in evolutionary endocrinology research.
+It is perhaps unsurprising that the few empirical papers that have emphasized the importance of different time courses (rather than only maximum) of the stress response have often focused on between group comparisons or investigated variation in the exact sampling time between individuals [e.g., @baugh2013; @small2017]
 
@@
 Clearly, when estimating the timing of peak glucocorticoids, a simple baseline plus induced sampling scheme is sub optimal, but this scheme may be perform well in other situations where the maximum value is the target and there is relatively little variation in response time.
-Creating simulations like this before studies are conducted has the potential increase the efficient use of researches time and funds, but also forces researches to think explicitly about quantitative predictions ahead of time.
+Creating simulations like this before studies are conducted has the potential to increase the efficient use of researches time and funds, but also forces researches to think explicitly about quantitative predictions ahead of time.
 These simulations could be included as part of a study pre-registration, grant application, or registered report to demonstrate exactly what data collection and analysis approaches are planned and to justify those decisions.
 
-## CONCLUSION
-
-While there has been increasing interest in understand within- and between-individual variation in the acute glucocorticoid response in recent years [CITE], the methods and data available to tackle these questions have changed relatively little.
-Many sophisticated statistical tools are now available and clear arguments have been made about the need to apply these approaches to endocrine traits [CITE].
-Arguably, the biggest roadblock at the moment is the limited availability of empirical data needed to test these ideas.
+## DISCUSSION
+
+While there has been increasing interest in understand within- and between-individual variation in the acute glucocorticoid response in recent years [@hau2016; @lema2013; @wada2014; @taff2016], the methods and data available to tackle these questions have changed relatively little.
+Many sophisticated statistical tools are now available and clear arguments have been made about the need to apply these approaches to endocrine traits, but relatively few empirical studies have effectively used these tools.
+Arguably, the biggest roadblock at the moment is the limited availability of empirical data needed to test hypotheses.
 Simulation offers one way forward, by allowing for more efficiently designed studies and by allowing researches to identify when the question of interest can *in principle* be answered with a given study design.
@@
 One of the main benefits of simulating datasets is identifying unmeasured properties and assumptions of currently available data that can become targets for empirical work.
-For example, I demonstrated that the covariation between different components of the acute stress resopnse and the relative amount of variation in each of these can have profound effects on the ability to accurately measure any single component.
+For example, I demonstrated that the covariation between different components of the acute stress response and the relative amount of variation in each of these can have profound effects on the ability to accurately measure any single component.
 Empirical work specifically designed to assess covariation and variance at different times could help to understand what conclusions we can reasonably draw from available data.
 One takeaway from these simulations is that variation in glucocorticoid-fitness relationships across seasons or life history stages can easily arise as a statistical artefact when between-individual variation in hormones also varies across the contexts.
-My study design example also suggests that there are potentially gains to be made by considering more diverse sampling designs tailored to the particular research question and study system.
-While standardized sample collection timing has been productive in this field, it also creates clear blind spots to certain types of variation between groups.
+The simulation exploring different sampling designs also suggests that there are potentially gains to be made by considering more diverse sampling designs tailored to the particular research question and study system.
+While standardized sample collection timing has allowed for large scale comparisons in this field [@vitousek2019], it also creates clear blind spots to certain types of variation between groups.
 
@@
 For example, I showed that one of the most common sampling design schemes has essentially no ability to detect a difference in the speed of increase between two groups if they do not also differ in maximum values.
-It is perhaps not surprising to find that there is little published evidence for differences in the speed of the acute response when most data sampled to date cannot detect those differences.
-Across a wide range of disciplines there has been an increasing push for pre-registration, reproducible research, and transparent research practices [CITE].
+It is perhaps not surprising to find that there is little published evidence for differences in the speed of the acute response when most study designs employed to date cannot *in principle* detect those differences.
+Across a wide range of disciplines there has been an increasing push for pre-registration, reproducible research, and transparent research practices [@odea2021].
 Simulation provides an opportunity for evolutionary endocrinologists to embrace these best practices by improving the quality of study design, allowing for more quantitative hypotheses and predictions, and providing a clear justification for experimental choices.
 
-This package and paper is meant only as a very cursory exploration of the ways that simulation can be applied to evolutionary endocrinology.
+This package and paper is meant only as an initial exploration of the ways that simulation can be applied to evolutionary endocrinology.
 I have no doubt that many more scenarios and complications could be added on to each of the simple examples presented here.
@@
 
-I would like to thank Michaela Hau and her lab along with John Wingfield for discussions about the ideas presented in this project.
-John Wingfield also provided an insightful overview of the history of study design in field based studies of acute stress responses.
+I would like to thank Michaela Hau and her lab for discussions about the ideas presented in this project.
+John Wingfield also provided an insightful overview of the history of study design in field based studies of the acute stress responses and pointed me to key references.
 Finally, I thank Maren Vitousek and her lab members for feedback and discussion on early versions of this project.

### 2022-06-30 3_r_scripts/Manuscript.Rmd
Working on revision.

@@
        xlab("Time (minutes)") + ylab("Corticosterone (ng/\U00B5l)") +
-       scale_color_manual(values = c("slateblue", "orange")) +
+       scale_color_manual(values = c("#9D93DE", "#FDAE42")) +
        #guides(color = "none") +

### 2022-06-30 3_r_scripts/Revision.Rmd
Working on revision.

--- 2022-05-05 3_r_scripts/Manuscript.Rmd
@@
 output: 
+  bookdown::word_document2:
+      number_sections: FALSE
+      toc: FALSE
+      extra_dependcies: ["flafter"]
   word_document:
@@
 These scenarios demonstrate the conditions under which a single glucocorticoid trait can be accurately assessed with typical experimental designs, the consequences of covariation between different components of the acute stress response, and the way that context specific differences in variability of acute responses can influence the power to detect relationships between the strength of the acute stress response and fitness.
-I also describe how to use the simulation tools to aid in the design and evaluation of empirical studies.
-Recently there has been a great deal of interest in understanding the causes and consequences of physiological flexibility, but empirical data are often extremely limited and traditional sampling methods are poorly designed to directly address the questions of interest.
-This simulation represents a way forward by revealing critical aspects of physiological flexibility and by creating a tool for designing better empirical studies that integrate simulation and theory with data collection.
+I also describe how to use the simulation tools to aid in the design and evaluation of empirical studies of physiological flexibility.
 
@@
 Successfully navigating these challenges often requires the ability to rapidly adjust behavior and physiology to match current conditions.
-For vertebrates, the glucocorticoid mediated stress response plays a major role in coordinating these changes when stressors are encountered [@wingfield1998; @sapolsky2000] and similar rapid response systems mediate changes in other taxa [@taborsky2020].
+For vertebrates, the glucocorticoid mediated stress response plays a major role in coordinating these changes when stressors are encountered [@wingfield1998; @sapolsky2000] and similar rapid response systems mediate changes in other taxa [@taborsky2021].
 Because of the central role that this response plays in coping with challenges, a great deal of research effort over the past 15 years has focused on understanding whether between-individual differences in the magnitude of this response predict coping ability and, ultimately, fitness [@breuner2008; @schoenle2020].
@@
 While these approaches are powerful, they have proven difficult to apply directly to endocrine flexibility data for two reasons.
-First, simulation studies suggest that many patterns may only be detectable with a level of repeated sampling that is possible for many behaviors (especially when collected autonomously), but that is currently not possible for most studies of endocrine flexibility.
-Second, and more fundamentally, these papers often focus on somewhat discrete measures of behavior (e.g., aggression score or activity level), whereas for acute glucocorticoid responses, the functional shape of the response itself may be the important trait and it may not be possible to summarize variation in the shape of the response with a single measure.
+First, simulation studies suggest that successfully modeling within-individual variation in flexible traits using an hierarchical modeling framework often requires a level of repeated sampling that is possible for many behaviors (especially when collected autonomously), but that is currently not possible for most studies of endocrine flexibility, because it would require sampling of many separate glucocorticoid responses per individual.
+Second, many behavioral papers focus on somewhat discrete measures (e.g., aggression score or activity level), whereas for acute glucocorticoid responses, the functional shape of the response itself may be the important trait.
+Fully describing the functional shape of a single acute glucocorticoid increase may require many samples in close succession, but for small vertebrates logistical and ethical constraints mean that it is rarely possible to take more than a few samples during the course of a single acute response.
 
 The function valued trait (FVT) framework is an alternative approach that explicitly considers the functional shape of a biological response [@gomulkiewicz2018; @stinchcombe2012; @kingsolver2015].
-While FVT approaches have been suggested for studies of endocrine flexibility [@taff2016], I am not aware of any papers that have applied this framework to empirical data on acute glucocorticoid responses, probably because sufficient data are not available.
+While FVT approaches have been suggested for studies of endocrine flexibility, I am not aware of any papers that have applied this framework to empirical data on acute glucocorticoid responses, probably because sufficient data are not available.
 Conceptually, however, this approach is a better match to the acute glucocorticoid response, because the shape of a response curve is explicitly considered as the phenotypic trait of interest.
@@
 This approach has been used to study a variety of phenotypes where values can be measured continuously or pooled across many individuals from the same group to accurately estimate the shape of a curve [see Table 1 in @stinchcombe2012].
-Applying the technique to endocrine flexibility at the within-individual level faces the same empirical challenges described for within-individual reaction norms above.
+Applying the technique to endocrine flexibility at the within-individual level faces the same empirical challenges described for within-individual reaction norms above, such as the need for repeated sampling of individuals and high temporal resolution of samples within individual physiological responses.
 Note that FVT and within-individual reaction norms approaches are not necessarily incompatible, but they have largely developed separately.
@@
 
+INSERT HISTORY AND CITATIONS FROM COCKREM
+
 For individual based studies, the most common approach to this problem is to standardize measurements as much as possible by measuring animals at the same time of the day during the same context, and by taking blood samples at standard times (often <3 and 30 minutes after capture) to characterize baseline and stress-induced glucocorticoids.
 This standardization allows for comparison between individuals, but in some cases it may also completely obscure the ability to detect variation in certain characteristics of the acute response curve.
-For example, if the speed (rate of initial increase) and scope (maximum value) of the acute response vary independently, samples taken at only two time points cannot accurately capture variation in either parameter.
+For example, if the speed (rate of initial increase or time required to reach maximum) and scope (maximum value) of the acute response vary independently, samples taken at only two time points cannot accurately capture variation in either parameter [@taff2022].
 Indeed, several discussions in recent years about methods such as the '3 minute rule' and the relative merits of 'area under the curve' versus time point measures of glucocorticoids are fundamentally related to a recognition of the importance of understanding variation in the functional shape of stress responses and whether different components of that shape covary within individuals [e.g., @cockrem2002; @small2017].
+While a great deal of empirical work has focused on characterizing the rate of initial increase [CITE], there has been relatively little work on understanding individual differences in the time required to reach maximal levels, and this attribute of speed is harder to estimate [@taff2022].
 
 One of the characteristics of both the within-individual reaction norm and FVT literature is that empirical work has proceeded in very close coordination with simulation and statistical method development.
-In contrast, studies of endocrine flexibility often point to these methods, but don't address the ways that the particular logistical challenges of hormone measurement might necessitate different empirical approaches.
-I believe this is one reason that there are currently more conceptual papers arguing for a reaction norm approach to endocrine variation than there are empirical papers actually applying the approach.
+In contrast, studies of endocrine flexibility often point to these methods, but don't address the ways that the particular logistical challenges of hormone measurement, such as the difficulty of repeatedly collecting samples within- or between-individuals, might necessitate different empirical approaches.
+I believe this is one reason that there are currently more conceptual papers arguing for a reaction norm approach to endocrine variation than there are empirical papers actually applying the approach [but see, @furtbauer2015; @taff2022; @lendvai2014; @houslay2022].
 While many of the tools developed in these related fields are transferable, studies of physiological flexibility would benefit from a focus on analysis development and testing that explicitly incorporates the particular details and challenges of these questions.
-One way to accomplish these goals is to use simulations, but to my knowledge no studies of physiological flexibility have developed simulations of the acute stress response that address the issues discussed above.
+One way to accomplish these goals is to integrate empirical work with simulations, but to my knowledge no studies of physiological flexibility have developed simulations of the acute stress response that address the issues discussed above.
+
+There is also a growing number of mathematical models of the  (Luttbeg, Taborsky)
 
@@
 
+The goal of this simulation is to provide a flexible tool that can produce realistic datasets of physiological flexibility for a variety of different systems and scenarios.
+As such, there are many possible applications and here I briefly highlight a few possibilities.
+First, I demonstrate that the simulation can produce datasets that are qualitatively similar to empirical data on acute glucocorticoid responses.
+Next, I explore three specific scenarios with the simulation that highlight challenges to empirical progress in this field.
+First, I ask how well maximum glucocorticoid measures can be estimated with a single time point measure in populations that differ systematically in features of the response.
+Second, I explore how different patterns of covariation between maximum glucocorticoids and one measure of speed (the time to reach maximum) influence the ability to reliably measure either component of the response.
+Third, I ask how different amounts of between- and within-individual variation in maximum glucocorticoid levels impact the ability to detect known relationships between glucocorticoids and fitness.
+These questions are all important for understanding physiologically flexibility and are difficult to fully address empirically, but this is by no means an exhaustive list of the questions or possible parameter permutations that could be explored with simulation.
+Finally, I demonstrate how the simulation tools can be used to evaluate the performance of different experimental designs given a set of logistical limitations (e.g., number of samples that can be collected) and a proposed hypothesis) and discuss how this procedure could be used as a planning tool to increase the power and reproducibility of future empirical work.
+
 ## MATERIALS AND METHODS
@@
 For illustration purposes, I refer to simulated glucocorticoid responses, but the simulation applies equally well to any physiological mediator of a rapid response.
-The package can be installed in R using the following command.
+The code for the complete simulation along with full documentation of each function and argument can be accessed in a GitHub repository with a current version permanently archived on Zenodo (LINK).
+The package can be installed directly within the R environment using the following command.
 
@@
 These parameters are sampled from a multivariate normal distribution with user specified mean, variance, and covariance for each parameter.
-I consider these values to be the 'true', unobserved, phenotype of the animal (setting aside the question of whether or not a 'true' physiological phenotype exists).
+By default, maximum glucocorticoid values are sampled from a normal distribution on the log scale and then exponentiated to determine absolute values.
+This results in a right skewed maximal glucocorticoid distribution on the absolute scale that is typical of many empirical datasets, but users can easily specify any parameters to be sampled from a normal or log normal distribution as required to match the characteristics of a particular study system.
+For the purposes of the simulation, I consider the sampled parameter values to be the 'true', unobserved, phenotype of the animal (setting aside the question of whether or not a 'true' physiological phenotype exists).
 
@@
 After sampling the parameters, values are interpolated for each one minute time point and a localized regression is fit to create a smoothed curve that represents the observed glucocorticoid response.
-From this expressed response, individual data points are then collected at user specified times that would reflect an empirical study design (e.g., 2, 30, and 60 minutes).
+From this expressed response, individual data points are then collected at user specified times that would reflect an empirical study design (e.g., 1, 30, and 60 minutes).
 Additional noise can be added to these data points to represent measurement error (e.g., assay error).
-
-The function also generates a simulated performance (e.g., fitness) measure, based on the underlying true values.
+This simulated dataset can then be treated as the input for any desired analyses and statistical approaches, while maintaining the ability to compare results to the 'true' values used in the simulation.
+
+The function also generates a simple simulated performance (e.g., fitness) measure, based on the underlying true parameter values sampled for each individual in the population (e.g., their baseline and maximum glucocorticoid value).
+The single fitness measure per animal is determined by allowing the user to specify the relative degree to which unmeasured traits plus each true parameter contribute to fitness outcomes.
 Data reflecting the true phenotypic values, the repeated expression of acute responses, and the observed time points can then be used in downstream analyses with any standard statistical approaches or software.
@@
       demo <- cort_sim2()
-      p <- plot_cort_sim(demo)
-      p
+
+          spoints <- unique(demo$simulated_dataset_long$time)
+          p1 <- ggplot(data = demo$simulated_dataset, mapping = aes(x = time, 
+              y = cort, by = animal_sample)) + geom_line(size = 0.5, 
+              alpha = 0.7, color = "coral3") + guides(color = FALSE) + theme_classic() + 
+              xlab("Time") + ylab("Corticosterone") + geom_vline(xintercept = spoints, 
+              linetype = "dashed") +
+              annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A", size = 5)
+          p2 <- ggplot(data = demo$timecourse_long, mapping = aes(x = time, 
+              y = cort, color = animal, by = animal_sample)) + geom_line(size = 0.5, 
+              color = "coral3", alpha = 0.7) + guides(color = FALSE) + 
+              theme_classic() + xlab("Time") + ylab("Corticosterone") + 
+              coord_cartesian(xlim = c(0, 60)) + geom_vline(xintercept = spoints, 
+              linetype = "dashed") +
+              annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B", size = 5)
+          p3 <- ggplot(data = demo$rank_timecourse, mapping = aes(x = time, 
+              y = rank, color = animal_sample2)) + geom_line(size = 1.6, 
+              alpha = 0.8) + guides(color = FALSE) + theme_classic() + 
+              scale_color_viridis(discrete = TRUE) + xlim(0, 60) + 
+              geom_vline(xintercept = spoints, linetype = "dashed") +
+              annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C", size = 5)
+
+      grid.arrange(p1, p2, p3, layout_matrix = rbind(c(1, 2, 2), 
+              c(3, 3, 3)))
 ```
@@
 
-### RESULTS
-
-The goal of this simulation is to provide a flexible tool that can produce realistic datasets of physiological flexibility for a variety of different systems and scenarios.
-As such, there are many possible applications and here I briefly highlight a few possibilities.
-These are by no means exhaustive, and I hope the simulation will be a useful tool to guide empirical work for specific hypotheses and study systems.
-Within each scenario, I have illustrated how the simulation functions might be used to address the particular question of interest, but I have not fully explored all the possible permutations of parameters systematically, because these will depend to a large extent on the empirical details of the system being studied.
-A complete set of reproducible code to create all of the examples presented in this paper is available on GitHub (https://github.com/cct663/speed_vs_scope).
-
-#### *Simulating empirically parameterized data*
+### METHODS FOR EACH SIMULATED SCENARIO
+
+Each of the scenarios described here uses the basic simulation functions described above with input parameters adjusted to address the question of interest.
+A complete set of reproducible code to create all of the examples presented here is available on GitHub and permanenty archived at Zenodo (https://github.com/cct663/speed_vs_scope).
+
+#### *Scenario 1: Simulating empirically parameterized data*
 
@@
 In addition to helping to design better studies, this approach has the advantage of increasing the transparency and reliability for studies of physiological flexibility, by making analysis choices and predictions clear before data are collected.
+
+To demonstrate this utility, I attempted to create simulated datasets that recreate the characteristics of the empirical data presented in Koolhaas et al.
+[-@koolhaas2010].
+As part of that study, a series of corticosterone measurements were collected during and after an acute stressor from 14 laboratory rats *Rattus norvegicus* using permanently implanted jugular vein canulae.
+I extracted data from Figure 6 in Koolhaas et al.
+[-@koolhaas2010] using WebPlotDigitizer (https://automeris.io/WebPlotDigitizer/) and then simulated data using the functions described above starting with the input values calculated directly from the empirical data.
+I then compared patterns of variation and population level response curves for the empirical and simulated data.
+
+#### *Scenario 2: Accurately measuring a single glucocorticoid trait*
+
+Single time point measures of glucocorticoids are often interpreted as representing meaningful variation between individuals.
+For example, variation in the level of glucocorticoids after 30 minutes of standardized restraint is typically interpreted as variation in the magnitude of the stress response [@taff2019].
+However, this interpretation rests on assumptions that are rarely explicitly tested with empirical data.
+For example, the time chosen to take a stress-induced sample is often assumed to be either at the species peak or during a plateau period after the species peak.
+In some early studies, great care was taken to determine an average population level peak time [e.g., @wingfield1992], but many studies adopt the widely used 'standard' time of 30 minutes post capture without extensive validation [compiled in @vitousek2019].
+While there is a general assumption that sampling later than the peak is acceptable (and perhaps preferable) because animals will be sampled during a relatively stable high plateau, there is little empirical data to evaluate this assertion or to determine how much under or overshooting the species peak timing might influence inferences.
+Furthermore, even when the average peak timing is well established, differences in the amount of between-individual variation in the time to reach the peak or in peak values are common across species and even in different life history stages within species [@wingfield1992].
+Some studies focus instead on the rate of initial increase in glucocorticoids with a sample taken at 10 or 15 minutes after disturbance [e.g., @cockrem2005; @cockrem2006; @love2003], often well before peak levels are reached.
+The combination of different patterns of within- and between-individual variation with the exact time(s) chosen for sampling could have consequences for the accuracy of point estimates taken at any single time point, but these questions cannot be addressed directly with empirical datasets where the true underlying values of each individual are unknown.
+
+I simulated a simple scenario to explore the consequences of variation in each of these parameters on the accuracy of estimating between individual differences in maximally expressed glucocorticoids during an acute response.
+For purposes of this illustration, I consider a single study design in which animals are sampled at 30 minutes.
+Using this design as a starting point, I systematically vary i) the timing of the population average peak (15, 30, or 45 minutes), ii) the amount of variation in maximum glucocorticoid levels reached (SD of 1 to 12 ng/$\mu$l), iii) and the amount of variation in the number of minutes taken to reach peak levels (SD of 1 to 20 minutes).
+All other variables in the simulation are constrained to be invariant between  individuals in the population (e.g., all individuals have identical baseline glucocorticoids in this case), though I consider cases in which multiple aspects of the rapid response are correlated with each other in the next scenario.
+I included moderate within-individual variability and a small amount of assay error across all iterations.
+For each combination of parameters, I simulated 200 animals and estimated the R^2^ value from a regression of the observed estimates of glucocorticoid levels at 30 minutes to the true known values.
+This simulation is likely a best case scenario for detection because it eliminates many sources of variation or noise that would be present in real data, but it illustrates the effect of variation in these three key parameters even when the exact same sampling design is employed.
+
+#### *Scenario 3: Exploring covariance between response components*
+
+In reality, fully characterizing the acute glucocorticoid response requires more than identifying just the maximum value reached.
+Individuals may differ in baseline levels, rate of initial increase, the speed of reaching the maximum level, time spent at maximum, and the speed of return to baseline.
+Moreover, each of these components of the endocrine response could be positively or negatively correlated with each other within and between individuals.
+In these cases, measurements taken at particular time points contain information about multiple aspects of the response and without additional information it may be difficult to know what trait is being measured.
+The fact that each of these traits might be important and that they might covary has been discussed in a general sense [e.g., @baugh2013], but simulations are uniquely powerful for exploring under exactly what conditions time point measure of glucocorticoids can or cannot be used as indicators of these traits.
+
+To illustrate this point, I explored the consequences of variation in the correlation between and relative amount of variation in just two aspects of the acute stress response: the maximum glucocorticoid level reached and the time required to reach the maximum level.
+For simplicity, I refer to the 'speed' of the response, but note that other aspects, such as the rate of initial increase, could also be considered as variation in the speed of response.
+When considering these two traits, a population of animals could plausibly display one of three patterns.
+Individuals that reach their maximum value faster might also reach higher values (figure \@ref(fig:cor-ex1)A; simulation correlation = -0.6).
+Alternatively, the speed and maximum values might vary independently (figure \@ref(fig:cor-ex1)B; correlation = 0).
+Finally, individuals that are faster responders might max out at lower glucocorticoid values (figure \@ref(fig:cor-ex1)C; correlation = 0.6).
+While many researchers in this field might have intuitions about which of these scenarios is most likely to prevail, there is very little empirical data available to actually determine which is most common.
+Moreover, regardless of the specifics for this particular correlation, the general pattern and considerations presented here will apply in similar ways to correlations between other aspects of the acute stress response.
+
+Using these three simulated populations as a starting point, I asked how well glucocorticoid values measured at one timepoint reflected true trait values.
+For each population I set an average population level speed of 30 minutes with other values in the simulation set at their default value.
+For every time point from 0 to 35 minutes I fit two simple linear regressions of the measured value on the true speed and maximum value and extracted the R^2^ value from the model.
+I repeated this simulation for all populations 50 times with 100 individuals sampled from the population each time.
+Finally, I repeated the entire set of simulations with each combination of low and high between-individual variation in the speed or maximum values (variation in speed: low = 2 minute SD, high = 12 minute SD; variation in maximum: low = 1ng/$\mu$l SD, high = 10ng/$\mu$l SD).
+
+#### *Scenario 4: Detecting links between fitness and responses*
+
+A common goal of recent studies is to establish whether variation in glucocorticoids is associated with fitness or some proxy for fitness [@schoenle2020].
+While there has been a great deal of discussion about the extent to which these relationships might differ with life history characteristics or between breeding stages, there has been relatively little consideration of the way that methodological limitations might limit the ability to detect these relationships even when they exist.
+
+Here, I imagine a simple scenario in which the 'true' maximum glucocorticoid level during an acute response explains 80% of the variation in fitness (clearly this is unrealistically high, but it is chosen for illustration only).
+I next construct a study in which researchers measure 50 individuals using a typical stress-induced (30 minute) sampling protocol.
+For simplicity, I set the other parameters in the simulation at their default values.
+Keeping the study design constant, I ask whether the glucocorticoid-fitness relationship can be recovered for two hypothetical populations that have low or high between-individual variation in maximum glucocorticoid levels.
+Using these populations, I asked asked how the ability to detect glucocorticoid-fitness relationships changed with different amounts of within-individual variation in acute response expression.
+For each combination of parameters, I simulated 50 populations and fit a simple linear regression model with observed glucocorticoid levels at 30 minutes as a predictor of fitness to ask whether the true glucocorticoid-fitness relationship was recovered.
+
+#### *Scenario 5: Designing optimal sampling strategies*
+
+One of the major benefits of simulating glucocorticoid response curves will be the ability to design optimal sampling strategies before data are collected.
+A simulation can be constrained to match any real world limitations (e.g., maximum number of samples possible per individual) and then explored to determine how to best allocate sampling resources.
+The specifics of this task will vary considerably with the study system and question being addressed, but here I illustrate one possible application.
+Consider an experiment in which the acute glucocorticoid response of a treatment group and control group are compared after some experimental manipulation.
+The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group, but will not differ in the maximum value itself.
+I have implemented this difference by simulating two populations in which the treatment group has a steeper initial slope and also reaches the maximum value faster.
+Any number of possible hypotheses for a particular study system could be specified following a similar approach.
+
+Using these simulated groups, I asked how well different study designs could detect the differences.
+Here we can impose any logistical constraints relevant to the study system.
+As an example, in this case we can only sample a maximum of 20 individuals per group, we can only sample each individual once post-treatment, and during that single sampling event we can take a blood sample at a maximum of three different time points, resulting in a total of 120 data points.
+Given these constraints, I compare several sampling designs.
+First, I explored 'standard' sampling strategies that are typical of empirical studies in this field that sample each individual animal at either i) 1 minute, 30 minutes, and 60 minutes, ii) 1 minute, 15 minutes, and 30 minutes, or iii) 1 minute, 15 minutes, and 60 minutes.
+Second, I explored two alternative sampling approaches that are not typically used in empirical work but might better capture the entire functional shape of response.
+These included, iv) a study in which three sampling times between 1 and 60 minutes are randomly chosen for every animal, and v) a study in which three sampling times are randomly chosen for each animal, but weighted more heavily around the range of times when maximum levels are expected to be reached for the population.
+
+For illustration purposes I sampled directly from the 'true' response curves in this example so that there is no additional measurement error added.
+To evaluate these schemes I compare estimates of the acute response curve for each group to the 'true' known curves.
+Note that a more complete analysis of a sampling schemes performance should include many more iterations and full statistical comparisons, but the details here will be highly dependent on the study system and goals, so I provide this simple example to illustrate the approach rather than to make any more widely applicable conclusions.
+
+### RESULTS
+
+#### *Scenario 1: Simulating empirically parameterized data*
+
+The simulation functions were able to produce a new synthetic dataset that has similar variation and patterns to the empirical data in Koolhaas et al.
+[-@koolhaas2010] (Figure \@ref(fig:kool-demo)A).
+The average population wide response curve shape also closely matched the empirical data (Figure \@ref(fig:kool-demo)B).
+In this case, the plotted simulation data include the same number of animals sampled at the same time points as the empirical data, but these sampling points and total sample size can easily be changed as desired.
+The parameterized simulation can now be used to test the sensitivity of any number of experimental designs before additional data is collected, such as different sample sizes or sampling time points.
 
@@
 
-To demonstrate this utility, I have redrawn data from Koolhaas et al.
-[-@koolhaas2010].
-As part of that study, a series of corticosterone measurements were collected during and after an acute stressor from 14 laboratory rats *Rattus norvegicus* using permanently implanted jugular vein canulae.
-I next simulated data using the functions described above starting with the input values calculated directly from the empirical data.
-The simulation creates a new dataset that has similar variation and patterns to the empirical data (Figure \@ref(fig:kool-demo)A) along with a population wide corticosterone response curve shape that closely matches the empirical data (Figure \@ref(fig:kool-demo)B).
-In this case, the plotted simulation data include the same number of animals sampled at the same time points as the empirical data, but these sampling points and total sample size can easily be changed as desired.
-The parameterized simulation can now be used to test the sensitivity of any number of experimental designs before additional data is collected.
-
-#### *Accurately measuring a single glucocorticoid trait*
-
-Single time point measures of glucocorticoids are often interpreted as representing meaningful variation between individuals.
-For example, variation in the level of glucocorticoids after 30 minutes of standardized restraint is typically interpreted as variation in the magnitude of the stress response [@taff2019].
-However, this interpretation rests on assumptions that are rarely explicitly tested with empirical data.
-For example, the time chosen to take a stress-induced sample is often assumed to be either at the species peak or during a plateau period after the species peak.
-In some early studies, great care was taken to determine an average population level peak time [@wingfield1992], but many studies adopt the widely used 'standard' time of 30 minutes post capture without extensive validation [compiled in @vitousek2019].
-While there is a general assumption that sampling later than the peak is acceptable (and perhaps preferable) because animals will be sampled during a relatively stable high plateau, there is little empirical data to evaluate this assertion or to determine how much under or overshooting the species peak timing might influence inferences.
-Furthermore, even when the average peak timing is well established, differences in the amount of between-individual variation in the time to reach the peak or in peak values are common across species and even in different life history stages within species [@wingfield1992].
-The combinations of these patterns of variation could have major consequences on the accuracy of single point estimates taken at 30 minutes, but these questions cannot be addressed directly with empirical datasets where the true underlying values of each individual are unknown.
-
-Here, I simulate a simple scenario exploring the consequences of variation in each of these parameters on the accuracy of estimating between individual differences in maximally expressed glucocorticoids during an acute response.
-For purposes of this illustration, I consider a single study design in which animals are sampled at 30 minutes.
-Using this design as a starting point, I systematically vary i) the timing of the population average peak (15, 30, or 45 minutes), ii) the amount of variation in maximum glucocorticoid levels reached, iii) and the amount of variation in the number of minutes taken to reach peak levels.
-All other variables in the simulation are constrained to be invariant between individuals in the population (e.g., all individuals have identical baseline glucocorticoids in this case), though I consider cases in which multiple aspects of the rapid response are correlated with each other in the next section.
-I included moderate within-individual variability and a small amount of assay error across all iterations.
-For each combination of parameters, I simulated 200 animals and estimated the R^2^ value from a regression of the observed estimates of glucocorticoid levels at 30 minutes to the true known values.
-This simulation is likely a best case scenario because it eliminates many sources of variation or noise that would be present in real data, but it illustrates the effect of variation in these three key parameters even when the exact same sampling design is employed.
+#### *Scenario 2: Accurately measuring a single glucocorticoid trait*
+
+The amount of between-individual variation in maximum glucocorticoid values has a profound effect on the ability to detect true maximal levels with samples taken at 30 minutes \@ref(fig:max-sim).
+A single sample taken at 30 minutes was highly correlated with true maximal glucocorticoid levels when between-individual variation wass high, but with low between-individual variation in maximum a single sample was uninformative.
+I simulated a wide range of variation that may include unrealistically high and low values, but the result highlights the importance of considering the amount of variation expected in a study population.
+There is a weaker, but still substantial impact of variation in the time taken to reach maximum values on the accuracy of estimates in this simulation.
+Greater variation in the speed (time to reach maximum) of the response reduces the accuracy of estimates of maximal values.
+Finally, the timing of sampling relative to the average population peak timing also influences accuracy.
+Measuring after the average peak time results in the most accurate estimates across a range of parameter values, while measuring before the average peak time produces the least accurate measures, particularly when there is also high variation in the time to reach maximum values between individuals.
 
@@
 
-Results of this simulation are summarized in figure \@ref(fig:max-sim).
-The amount of between-individual variation in the maximum glucocorticoid value has a profound effect on the ability to detect true maximal levels with samples taken at 30 minutes.
-In one sense, this result is unsurprising because it is intuitive that large differences would be easier to detect, but there are important consequences of this fact for interpreting studies that seek to link between-individual variation in the magnitude of the stress response with other traits.
-For example, the magnitude of the acute stress response often varies substantially across life history stages [@wingfield1992].
-Even if study designs are identical and maximum glucocorticoids are associated with performance, it will be easier to detect those patterns during life history stages with greater variation (see section on detecting fitness associations below).
-There is a weaker, but still substantial impact of variation in the time taken to reach maximum values on the accuracy of estimates in this simulation.
-Greater variation in the speed of the response reduces the accuracy of estimates of maximal values.
-Finally, the timing of sampling relative to the average population peak timing also influences accuracy.
-Measuring after the average peak time results in the most accurate estimates across a range of parameter values, while measuring before the average peak time produces the least accurate measures, particularly when there is also high variation in the time to reach maximum values between individuals.
-This simple example demonstrates clearly that the same experimental design will perform better or worse depending on the combination of glucocorticoid regulation parameters in the population being studied.
-
-#### *Exploring covariance between response components*
-
-In reality, fully characterizing the acute glucocorticoid response requires more than identifying just the maximum value reached.
-Individuals may differ in baseline levels, rate of initial increase, the speed of reaching the maximum level, time spent at maximum, and the speed of return to baseline.
-Moreover, each of these components of the endocrine response could be positively or negatively correlated with each other within and between individuals.
-In these cases, measurements taken at particular time points contain information about multiple aspects of the response and without additional information it may be difficult to know what trait is being measured.
-The fact that each of these traits might be important and that they might covary has been discussed in a general sense [e.g., @baugh2013], but simulations are uniquely powerful for exploring under exactly what conditions time point measure of glucocorticoids can or cannot be used as indicators of these traits.
+#### *Scenario 3: Exploring covariance between response components*
+
+The time that samples were taken at, relative amount of variation in speed and maximum, and degree of correlation between the speed and maximum all had substantial impacts on the ability to infer true trait values from single time point glucocorticoid measures (figure \@ref(fig:cor-ex2)).
+Neither speed nor maximum traits could be assessed accurately when between-individual variation in both traits was low (figure \@ref(fig:cor-ex2)A).
+Overall, accurately assessing variation in speed was much harder---if not impossible---with single measures.
+
+It was only possible to accurately estimate speed when high between-individual variation in speed was coupled with low variation in maximal values, but this situation may be rare in natural populations.
+When speed was tightly correlated with maximum (figure \@ref(fig:cor-ex2)D) it was sometimes possible to attain reasonable estimates of speed (figure \@ref(fig:cor-ex2)C-D), but when speed was not correlated with maximum, single measures were not good indicators of variation in speed (figure \@ref(fig:cor-ex2)A, C-D).
+Finally, measuring variation in maximum values was much easier under many conditions (figure \@ref(fig:cor-ex2)C-D), but the accuracy of assessment of maximum values was also negatively impacted by variation in speed and the degree of this impact differed depending on the correlation between the two traits (figure \@ref(fig:cor-ex2)).
 
@@
 ```
-
-To illustrate this point, I explored the consequences of variation in the correlation between and relative amount of variation in just two aspects of the acute stress response: the maximum glucocorticoid level reached and the time required to reach the maximum level.
-For simplicity, I refer to the 'speed' of the response, but note that other aspects, such as the rate of initial increase, could also be considered as variation in the speed of response.
-When considering these two traits, a population of animals could plausibly display one of three patterns.
-Individuals that reach their maximum value faster might also reach higher values (figure \@ref(fig:cor-ex1)A; simulation correlation = -0.6).
-Alternatively, the speed and maximum values might vary independently (figure \@ref(fig:cor-ex1)B; correlation = 0).
-Finally, individuals that are faster responders might max out at lower glucocorticoid values (figure \@ref(fig:cor-ex1)C; correlation = 0.6).
-While many researchers in this field might have intuitions about which of these scenarios is most likely to prevail, there is very little empirical data available to actually determine which is most common.
-Moreover, regardless of the specifics for this particular correlation, the general pattern and considerations presented here will apply in similar ways to correlations between other aspects of the acute stress response.
 
@@
 
-Using these three simulated populations as a starting point, I asked how well glucocorticoid values measured at one timepoint reflected true trait values.
-For each population I set an average population level speed of 30 minutes with other values in the simulation set at their default value.
-For every time point from 0 to 35 minutes I fit two simple linear regressions of the measured value on the true speed and maximum value and extracted the R^2^ value from the model.
-I repeated this simulation for all populations 50 times with 100 individuals sampled from the population each time.
-Finally, I repeated the entire set of simulations with each combination of low and high between-individual variation in the speed or maximum values (variation in speed: low = 2 minute SD, high = 12 minute SD; variation in maximum: low = 1ng/$\mu$l SD, high = 10ng/$\mu$l SD).
-
-The time that samples were taken at, relative amount of variation in speed and maximum, and degree of correlation between the speed and maximum all had substantial impacts on the ability to infer true trait values from single time point glucocorticoid measures (figure \@ref(fig:cor-ex2)).
-While these scenarios do not explore all possible parameter space, there are several clear conclusions that can be made.
-First, neither speed or maximum traits could be assessed accurately when between-individual variation in both traits was low (figure \@ref(fig:cor-ex2)A).
-This is potentially important for interpreting apparent differences in glucocorticoid fitness relationships because between-individual variation is known to differ across life history stages [@wingfield1992].
-Second, accurately assessing variation in speed was much harder---if not impossible---with single measures.
-
-It was only possible to accurately estimate speed when high between-individual variation in speed was coupled with low variation in maximal values, but this situation may be rare in natural populations.
-When speed was tightly correlated with maximum (figure \@ref(fig:cor-ex2)D) it was sometimes possible to attain reasonable estimates of speed (figure \@ref(fig:cor-ex2)C-D), but when speed was not correlated with maximum, single measures were not good indicators of variation in speed (figure \@ref(fig:cor-ex2)A, C-D).
-Finally, measuring variation in maximum values was much easier under many conditions (figure \@ref(fig:cor-ex2)C-D), but the accuracy of assessment of maximum values was also negative impacted by variation in speed and the degree of this impact differed depending on the correlation between the two traits (figure \@ref(fig:cor-ex2)).
-Beyond the specifics of this particular example, what these results demonstrate clearly is that understanding what aspect of the glucocorticoid response is being measured by any particular study design depends on extensive knowledge of the overall shape and amount of variation in different aspects of the acute stress response.
-
-#### *Detecting links between fitness and responses*
-
-A common goal of recent studies is to establish whether variation in glucocorticoids is associated with fitness or some proxy for fitness [@schoenle2020].
-While there has been a great deal of discussion about the extent to which these relationships might differ with life history characteristics or between breeding stages, there has been relatively little consideration of the way that methodological limitations might limit the ability to detect these relationships even when they exist.
-
-```{r fitness-ex, echo = FALSE, fig.height = 8, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Relationship between observed maximum glucocorticoid values and fitness for simulated populations that have low between-individual variation (blue) or high between-individual variation (red).
+#### *Scenario 4: Detecting links between fitness and responses*
+
+Several patterns can be identified by examining the results of this simulation.
+First, as specified by the simulation parameters, the correlation between the true maximum glucocorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex)A-B).
+In all cases, however, the observed correlation is lower than the true correlation and always lowest in the population with low between-individual variation.
+The ubiquity of this pattern is a product of the simulation structure, because adding within-individual variation effectively adds noise to the true correlation.
+It is important to note that in the real world, it is unlikely that this pattern would be so universal, because unmeasured variables could influence both fitness and glucocorticoids.
+For example, if habitat quality directly alters fitness and glucocorticoids, the observed correlation could be stronger than the 'true' correlation.
+Thus, interpretation of these results should be made cautiously in light of the simplicity of the simulation compared to real world conditions.
+Nevertheless, general patterns illustrated by the simulation are likely to pertain across a wide range of conditions.
+
+It was easiest to detect statistical evidence for known glucocorticoid-fitness relationships when within-individual variation was low (figure \@ref(fig:fitness-ex)A).
+Under these conditions, the true pattern was recovered in nearly all simulated populations with high between-individual variation and in approximately half of the populations with low between-individual variation.
+It becomes harder to detect these true relationships within-individual variation is high (figure \@ref(fig:fitness-ex)B) are high, but even in these more challenging situations the relationship can be detected the majority of the time if between-individual variation in maximum levels is high.
+When within-individual variation is high and between-individual variation is low, it is nearly impossible to detect glucocorticoid-fitness relationships.
+
+```{r fitness-ex, echo = FALSE, fig.height = 4, fig.width = 8, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Relationship between observed maximum glucocorticoid values and fitness for simulated populations that have low between-individual variation in maximum glucocorticoids (blue) or high between-individual variation in maximum glucocorticoids (red).
 Each point is the result of a separate simulation of 50 individuals using the settings described in the text.
 Filled circles are simulations in which observed glucocorticoid values at 30 minutes were significantly correlated with fitness and crosses are simulations in which the relationship was not significant.
-Panels illustrate conditions with low measurement error (A, C) versus high measurement error (B, D) and low within-individual variation (A, B) versus high within-individual variation (C, D).
-For each simulation, the correlation between true maximum glucocorticoids fitness is plotted on the y-axis and the correaltion with observed values is plotted on the x-axis."}
-#draw figure ----
+Panels illustrate conditions with low within-individual variation (A) versus high within-individual variation (B).
+For each simulation, the correlation between true maximum glucocorticoids fitness is plotted on the y-axis and the correlation with observed values is plotted on the x-axis."}
+# #draw figure ----
+# set.seed(99)
 # n_sam <- 50
@@
 # 
+# # Low within variation
+# 
 # for(i in 1:n_sam){
-#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.1, max_error = 0.2,
+#       f1 <- simcoRt::cort_sim2(data = simcoRt::cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 25, speed_sd = 1),
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.2, max_error = 0.15,
 #                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
@@
 # 
-#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#     f2 <- simcoRt::cort_sim2(data = simcoRt::cort_sim1(n = 50, speed_mu = 25, max_mu = log(45), max_sd = 0.03, speed_sd = 1),
 #                                 bleed_times = c(2, 30), sample_times = 1,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.1, max_error = 0.2)
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.2, max_error = 0.15)
 #     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
@@
 # 
-# fd1$group <- "High"
+# fd1$group <- "Norm"
 # fd2$group <- "Low"
@@
 #   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
-#   scale_color_manual(values = c("coral3", "slateblue")) +
-#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   scale_color_manual(values = c("slateblue", "coral3")) +
+#   scale_fill_manual(values = c("slateblue", "coral3")) +
 #   theme_bw() +
 #   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   ggtitle("Low measurement error \nLow within-individual variation") +
-#   xlab("") + ylab("True correlation") +
+#   ggtitle("Low within-individual variation") +
+#   xlab("Observed correlation") + ylab("True correlation") +
 #   theme(legend.position = c(0.75, 0.15)) +
@@
 #   #theme(legend.background = element_rect(fill = alpha("white", 0))) +
-#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A", size = 8) +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A", size = 6) +
 #   theme(axis.title = element_text(size = 13))
@@
 # 
-# ## Low assay, high within
-# 
+# ## High within
+# 
+# set.seed(99)
 # for(i in 1:n_sam){
-#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.4, max_error = 0.1,
+#     f1 <- simcoRt::cort_sim2(data = simcoRt::cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 25, speed_sd = 1),
+#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.2, max_error = 0.6,
 #                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
@@
 # 
-#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
+#     f2 <- simcoRt::cort_sim2(data = simcoRt::cort_sim1(n = 50, speed_mu = 25, max_mu = log(45), max_sd = 0.03, speed_sd = 1),
 #                                 bleed_times = c(2, 30), sample_times = 1,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.4, max_error = 0.1)
+#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.2, max_error = 0.6)
 #     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
@@
 # 
-# fd1$group <- "High"
-# fd2$group <- "Low"
+# fd1$group <- "Norm"
+# fd2$group <- "High"
 # 
@@
 #   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
-#   scale_color_manual(values = c("coral3", "slateblue")) +
-#   scale_fill_manual(values = c("coral3", "slateblue")) +
+#   scale_color_manual(values = c("slateblue", "coral3")) +
+#   scale_fill_manual(values = c("slateblue", "coral3")) +
 #   theme_bw() +
 #   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   ggtitle("High measurement error \nLow within-individual variation") +
-#   xlab("") + ylab("") +
-#   #theme(legend.position = c(0.82, 0.15)) +
-#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
-#   guides(fill = "none", color = "none") +
-#   scale_shape_manual(values = c(21, 3), guide = "none") +
-#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B", size = 8) +
-#   theme(axis.title = element_text(size = 13)) 
-#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "High measurement error \nLow within-individual variation", hjust = 0)
-# 
-# pB <- ggExtra::ggMarginal(pb, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both") 
-# 
-# ## Low assay, high within
-# 
-# for(i in 1:n_sam){
-#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.1, max_error = 0.6,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
-#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
-#     f1b <- subset(f1b, f1b$time == 30)
-#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
-#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
-#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
-#     
-#     
-#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.1, max_error = 0.6)
-#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
-#     f2b <- subset(f2b, f2b$time == 30)
-#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
-#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
-#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
-#     print(paste(i, "of", n_sam, sep = " "))
-# }
-# 
-# fd1 <- fit_dat[, 1:3]
-# fd2 <- fit_dat[, 4:6]
-# colnames(fd1) <- c("true", "obs", "p")
-# colnames(fd2) <- c("true", "obs", "p")
-# 
-# fd1$dif <- fd1$obs - fd1$true
-# fd2$dif <- fd2$obs - fd2$true
-# 
-# fd1$group <- "High"
-# fd2$group <- "Low"
-# 
-# f3 <- rbind(fd1, fd2)
-# 
-# f3$group <- as.factor(f3$group)
-# f3$shp <- NA
-# for(i in 1:nrow(f3)){
-#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
-# }
-# f3c <- f3
-# pc <- ggplot(data = f3c, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
-#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
-#   #geom_smooth(method = "lm", alpha = 0.3) +
-#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
-#   scale_color_manual(values = c("coral3", "slateblue")) +
-#   scale_fill_manual(values = c("coral3", "slateblue")) +
-#   theme_bw() +
-#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   ggtitle("Low measurement error \nHigh within-individual variation") +
-#   xlab("Observed correlation") + ylab("True correlation") +
-#   #theme(legend.position = c(0.82, 0.15)) +
-#   #guides(fill = guide_legend("Between-individual \nvariation"), color = guide_legend("Between-individual \nvariation")) +
-#   guides(fill = "none", color = "none") +
-#   scale_shape_manual(values = c(21, 3), guide = "none") +
-#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C", size = 8) +
-#   theme(axis.title = element_text(size = 13)) 
-#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "Low measurement error \nHigh within-individual variation", hjust = 0)
-# 
-# pC <- ggExtra::ggMarginal(pc, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
-# 
-# ## high assay, high within
-# 
-# for(i in 1:n_sam){
-#       f1 <- cort_sim2(data = cort_sim1(n = 50, max_mu = log(45), max_sd = 0.21, speed_mu = 30, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1, assay_error = 0.3, max_error = 0.6,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5))
-#     f1b <- plyr::join(f1$simulated_dataset_long[, c("animal", "time", "cort")], f1$true_values, "animal")
-#     f1b <- subset(f1b, f1b$time == 30)
-#     fit_dat$f1b_true[i] <- cor(f1b$max, f1b$performance)
-#     fit_dat$f1b_obs[i] <- cor(f1b$cort, f1b$performance)
-#     fit_dat$f1b_p[i] <- coef(summary(lm(performance ~ cort, data = f1b)))[2, 4]
-#     
-#     
-#     f2 <- cort_sim2(data = cort_sim1(n = 50, speed_mu = 30, max_mu = log(45), max_sd = 0.03, speed_sd = 1), 
-#                                 bleed_times = c(2, 30), sample_times = 1,
-#                     performance_contributions = c(0, 0, 5, 0, 0, 0, 0, 5), assay_error = 0.3, max_error = 0.6)
-#     f2b <- plyr::join(f2$simulated_dataset_long[, c("animal", "time", "cort")], f2$true_values, "animal")
-#     f2b <- subset(f2b, f2b$time == 30)
-#     fit_dat$f2b_true[i] <- cor(f2b$max, f2b$performance)
-#     fit_dat$f2b_obs[i] <- cor(f2b$cort, f2b$performance)
-#     fit_dat$f2b_p[i] <- coef(summary(lm(performance ~ cort, data = f2b)))[2, 4]
-#     print(paste(i, "of", n_sam, sep = " "))
-# }
-# 
-# fd1 <- fit_dat[, 1:3]
-# fd2 <- fit_dat[, 4:6]
-# colnames(fd1) <- c("true", "obs", "p")
-# colnames(fd2) <- c("true", "obs", "p")
-# 
-# fd1$dif <- fd1$obs - fd1$true
-# fd2$dif <- fd2$obs - fd2$true
-# 
-# fd1$group <- "High"
-# fd2$group <- "Low"
-# 
-# f3 <- rbind(fd1, fd2)
-# 
-# f3$group <- as.factor(f3$group)
-# f3$shp <- NA
-# for(i in 1:nrow(f3)){
-#   ifelse(f3$p[i] < 0.05, f3$shp[i] <- "A", f3$shp[i] <- "B")
-# }
-# f3d <- f3
-# pd <- ggplot(data = f3d, mapping = aes(x = obs, y = true, colour = group, fill = group)) +
-#   geom_point(alpha = 0.7, mapping = aes(shape = shp)) + 
-#   #geom_smooth(method = "lm", alpha = 0.3) +
-#   #geom_boxplot(mapping = aes(x = true, y = obs), orientation = "y", alpha = 0.3, outlier.shape = NA, width = 1) +
-#   scale_color_manual(values = c("coral3", "slateblue")) +
-#   scale_fill_manual(values = c("coral3", "slateblue")) +
-#   theme_bw() +
-#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   ggtitle("High measurement error \nHigh within-individual variation") +
+#   ggtitle("High within-individual variation") +
 #   xlab("Observed correlation") + ylab("") +
@@
 #   scale_shape_manual(values = c(21, 3), guide = "none") +
-#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "D", size = 8) +
+#   annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B", size = 6) +
 #   theme(axis.title = element_text(size = 13))
-#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "High measurement error \nHigh within-individual variation", hjust = 0)
-# 
-# pD <- ggExtra::ggMarginal(pd, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
-# 
-# pAD <- ggpubr::ggarrange(pA, pB, pC, pD, nrow = 2, ncol = 2)
-# 
-# saveRDS(pAD, here::here("5_other_outputs/fitness-ex.rds"))
+#   #annotate(geom = "text", x = 0.19, y = 0.51, label = "High measurement error \nLow within-individual variation", hjust = 0)
+# 
+# pB <- ggExtra::ggMarginal(pb, groupFill = TRUE, groupColour = TRUE, type = "boxplot", margins = "both")
+# 
+# pAB <- ggpubr::ggarrange(pA, pB)
+# 
+# 
+# saveRDS(pAB, here::here("5_other_outputs/fitness-ex.rds"))
 
 # reload and print ----
-pAD <- readRDS(here::here("5_other_outputs/fitness-ex.rds"))
-pAD
+pAB <- readRDS(here::here("5_other_outputs/fitness-ex.rds"))
+pAB
 
@@
 
-Here, I imagine a simple scenario in which the 'true' maximum glucocorticoid level during an acute response explains 80% of the variation in fitness (clearly this is unrealistically high, but it is chosen for illustration only).
-I next construct a study in which researchers measure 50 individuals using a typical stress-induced (30 minute) sampling protocol.
-For simplicity, I set the other parameters in the simulation at their default values.
-Keeping the study design constant, I ask whether the glucocorticoid-fitness relationship can be recovered for two hypothetical populations that have low or high between-individual variation in maximum glucocorticoid levels.
-For each of these two populations, I ask how the ability to detect glucocorticoid-fitness relationships changes with different amounts of within-individual variation in acute response expression and with differing amounts of measurement error.
-For each combination of parameters, I simulated 50 populations and fit a simple linear regression model with observed glucocorticoid levels at 30 minutes as a predictor of fitness to ask whether the true glucocorticoid-fitness relationship was recovered.
-
-Several patterns can be identified by examining the results of this simulation.
-First, the correlation between the true maximum glucocorticoid value and fitness does not differ for populations simulated with high or low between-individual variation (figure \@ref(fig:fitness-ex)A-D).
-In all cases, however, the observed correlation is lower than the true correlation and always lowest in the population with low between-individual variation.
-The ubiquity of this pattern is a product of the simulation structure, because adding measurement error or within-individual variation effectively adds noise to the true correlation.
-It is important to note that in the real world, it is unlikely that this pattern would be so universal, because unmeasured variables could influence both fitness and glucocorticoids.
-For example, if habitat quality directly alters fitness and glucocorticoids, the observed correlation could be stronger than the 'true' correlation.
-Thus, interpretation of these results should be made cautiosly in light of the simplicity of the simulation compared to real world conditions.
-
-Nevertheless, general patterns illustrated by the simulation are likely to pertain across a wide range of conditions.
-In this case, it is easiest to detect significant glucocorticoid-fitness relationships when both measurement error and within-individual variation are low (figure \@ref(fig:fitness-ex)A).
-It becomes harder to detect these true relationships when either measurement error (figure \@ref(fig:fitness-ex)B) or within-individual variation (figure \@ref(fig:fitness-ex)C) are high, but even in these more challenging situations the relationship can be detected the majority of the time if between-individual variation in maximum levels is high.
-When both measurement error and within-individual variation are high, it is nearly impossible to detect glucocorticoid-fitness relationships with low-between individual variation, but in populations with high between-individual variation the relationship is still detected in about half of the simulations.
-
-The fact that low between-individual variation in maximum glucocorticoids makes it harder to detect true glucocorticoid-fitness relationships across a wide range of conditions has important consequences for interpreting empirical results.
-Many studies have demonstrated different relationships (or lack thereof) between corticosterone and fitness at different life history stages [@vitousek2018; @bonier2009b], but it is also well known that the absolute amount of between individual variation in glucocorticoid traits varies considerably at different stages [@wingfield1992].
-Our simulation demonstrates that the power to detect true relationships will differ drastically across these conditions even with identical study designs and samples sizes, suggesting that great care is needed to conclusively differentiate true differences in glucocorticoid-fitness relationships across contexts from statistical artefacts.
-
-#### *Designing optimal sampling strategies*
-
-One of the major benefits of simulating glucocorticoid response curves will be the ability to design optimal sampling strategies before data are collected.
-A simulation can be constrained to match any real world limitations (e.g., maximum number of samples possible per individual) and then explored to determine how to best allocate sampling resources.
-The specifics of this task will vary considerably with the study system and question being addressed, but here I illustrate one possible application.
-Consider an experiment in which the acute glucocorticoid response of a treatment group and control group are compared after some experimental manipulation.
-The details of the manipulation are unimportant here, but suppose that the prediction is that this manipulation should result in a difference in the speed of the corticosterone response between our two groups, such that the treatment group will reach it's maximum glucocorticoid value faster than the control group, but will not differ in the maximum value itself.
-I have implemented this difference by simulating two populations in which the treatment group has a steeper initial slope and also reaches the maximum value faster (figure \@ref(fig:prediction-ex)).
-Any number of possible hypotheses for a particular study system could be specified following a similar approach.
+#### *Scenario 5: Designing optimal sampling strategies*
+
+In this simulated scenario, standard sampling schemes cannot fully describe the difference between treatment groups (Figure \@ref(fig:prediction-ex); \@ref(fig:treatment-ex)A-C).
+Sampling at 1, 30, and 60 minutes completely fails to distinguish the difference between groups, despite the fact that the treatment group reaches it's maximum value on average 12 minutes (~40%) faster than the control group.
+Many empirical studies include an earlier sample at 15 minutes to capture the rate of initial increase [e.g., @cockrem2005, @cockrem2006, @cockrem2007, @huber2021, @silverin1998].
+Shifting one of the sampling time points in this simulation to 15 minutes does recover a clear treatment difference in initial increase, but cannot capture the large difference in the time required to reach maximum values between the groups \@ref(fig:treatment-ex)B-C).
+
+In contrast, both the random sampling and weighted sampling schemes detect differences in the overall shape of the acute response (figure \@ref(fig:treatment-ex) D-E).
+In this particular scenario, there is no clear difference between these two approaches and both perform well in describing both the difference in initial increase and the difference in the time required to reach maximal levels in each group.
 
@@
 
-Next, we can ask how well different study designs can detect this difference.
-Here we can impose any logistical constraints relevant to the study system.
-As an example, in this case we can only sample a maximum of 20 individuals per group, we can only sample each individual once post-treatment, and during that single sampling event we can take a blood sample at a maximum of two different time points, resulting in a total of 80 data points.
-Given these constraints, I compare three different sampling designs: i) a study in which every animal is sampled at 1 minute, 30 minutes, and 60 minutes, ii) a study in which two sampling times between 1 and 60 minutes are randomly chosen for every animal, iii) a study in which two sampling times are randomly chosen for each animal, but weighted more heavily around the range of times when maximum levels are expected to be reached for the population.
-
-Note that the first sampling scheme closely mirrors the most common empirical design and in this case I have allowed an extra, third sample at 60 minutes, such that it includes 120, rather than 80, data points.
-For illustration purposes I sampled directly from the 'true' response curves in this example so that there is no additional measurement error added.
-To evaluate these schemes I compare estimates of the acute response curve for each group to the 'true' known curves shown in figure \@ref(fig:prediction-ex).
-Note that a more complete analysis of a sampling schemes performance should include many more iterations and full statistical comparisons, but the details here will be highly dependent on the study system and goals, so I provide this simple example to illustrate the approach rather than to make any more widely applicable conclusions.
-
-```{r treatment-ex, echo = FALSE, fig.height = 2.8, fig.width = 7.3, fig.pos = "!ht", fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Three possible sampling schemes to compare two groups.
-For standard sampling (A), every individual is sampled at exactly 2, 30, and 60 minutes.
-For random sampling (B) each individual is sampled at two random points between 1 and 60 minutes.
-For weighted sampling (C) two sampling times are chosen for each individual from a normal distribution with mean of 32 and sd of 9 minutes.
-In all three panels, solid lines are the true group averages, dashed lines are the estimates based on samples, and points are individual samples collected."}
+```{r treatment-ex, echo = FALSE, fig.height = 5.6, fig.width = 7.3, fig.align = "center", message = FALSE, warning = FALSE, fig.cap = "Five possible sampling schemes to compare two groups.
+For standard sampling (A-C), every individual is sampled at exactly 3 time points that include baseline along with samples at 15, 30, or 60 minutes.
+For random sampling (D) each individual is sampled at three random points between 1 and 60 minutes.
+For weighted sampling (E) three sampling times are chosen for each individual from a normal distribution with mean of 32 and sd of 9 minutes.
+In all panels, solid lines are the true group averages, dashed black lines are the estimates based on samples, shaded intervals are the 95% estimates from a simple generalized additive model, and points are individual samples collected."}
 
@@
        theme(axis.title = element_text(size = 14)) +
-       geom_smooth(se = FALSE, size = 1.1, mapping = aes(by = type, color = type)) +
+       geom_smooth(se = FALSE, size = 1, mapping = aes(by = type, color = type)) +
        theme(legend.title = element_blank())
+
+# reviewer requested scheme
+      ddx <- subset(comp_sim2, comp_sim2$time == 1 | comp_sim2$time == 15 | comp_sim2$time == 30)
+      ddx$type <- paste(ddx$type, "sub", sep = "_")
+
+     pxa <- p1 +
+       #geom_line(data = ddx, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
+              geom_point(data = ddx, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6, size = 0.8) +
+       geom_smooth(data = ddx, size = 0.8, mapping = aes(x = time, y = cort, by = type, color = "black", fill = type), linetype = "twodash") +
+       scale_fill_manual(values = c("#968BDC", "#FFC04C")) +
+       scale_color_manual(values = c("black", "slateblue", "#968BDC", "orange", "#FFC04C")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B") +
+       guides(fill = "none") +
+       ylab("") +
+       ggtitle("Standard: 1, 15, 30")
+
+# reviewer requested scheme 2
+      ddx <- subset(comp_sim2, comp_sim2$time == 1 | comp_sim2$time == 15 | comp_sim2$time == 60)
+      ddx$type <- paste(ddx$type, "sub", sep = "_")
+
+     pxa2 <- p1 +
+       #geom_line(data = ddx, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
+              geom_point(data = ddx, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6, size = 0.8) +
+       geom_smooth(data = ddx, size = 0.8, mapping = aes(x = time, y = cort, by = type, color = "black", fill = type), linetype = "twodash") +
+       scale_fill_manual(values = c("#968BDC", "#FFC04C")) +
+       scale_color_manual(values = c("black", "slateblue", "#968BDC", "orange", "#FFC04C")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C") +
+       guides(fill = "none") +
+       ylab("") +
+       ggtitle("Standard: 1, 15, 60")
 
@@
      dd1 <- subset(comp_sim2, comp_sim2$time == 1 | comp_sim2$time == 30 | comp_sim2$time == 60)
+     dd1$type <- paste(dd1$type, "sub", sep = "_")
 
@@
        #geom_line(data = dd1, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
-       geom_smooth(data = dd1, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed") +
-       scale_fill_manual(values = c("slateblue", "orange")) +
+              geom_point(data = dd1, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6, size = 0.8) +
+       geom_smooth(data = dd1, size = 0.8, mapping = aes(x = time, y = cort, by = type, color = "black", fill = type), linetype = "twodash") +
+       scale_fill_manual(values = c("#968BDC", "#FFC04C")) +
+       scale_color_manual(values = c("black", "slateblue", "#968BDC", "orange", "#FFC04C")) +
        annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A") +
        guides(fill = "none") +
-       geom_point(data = dd1, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
-       ggtitle("Standard sampling")
+       ggtitle("Standard: 1, 30, 60")
 
@@
        sub <- subset(cs2, cs2$an2 == ans$animal_sample[i])
-       subs <- sub[sample(seq(1, 60, 1), 2, replace = FALSE), ]
+       subs <- sub[sample(seq(1, 60, 1), 3, replace = FALSE), ]
        dd2 <- rbind(dd2, subs)
@@
      dd2 <- subset(dd2, is.na(dd2$an2) == FALSE)
+     dd2$type <- paste(dd2$type, "sub", sep = "_")
 
@@
        #geom_line(data = dd2, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
-       geom_smooth(data = dd2, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed") +
-       scale_fill_manual(values = c("slateblue", "orange")) +
-       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B") +
+       geom_point(data = dd2, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6, size = 0.8) +
+       geom_smooth(data = dd2, size = 0.8, mapping = aes(x = time, y = cort, by = type, color = "black", fill = type), linetype = "twodash") +
+       scale_fill_manual(values = c("#968BDC", "#FFC04C")) +
+       scale_color_manual(values = c("black", "slateblue", "#968BDC", "orange", "#FFC04C")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "D") +
        guides(fill = "none") +
-       ylab("") +
-       geom_point(data = dd2, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
        ggtitle("Random sampling")
@@
      for(i in 1:nrow(ans)){
-       choose <- round(rnorm(2, 32, 9))
+       choose <- round(rnorm(3, 32, 9))
        sub <- subset(cs2, cs2$an2 == ans$animal_sample[i])
@@
      dd3 <- subset(dd3, is.na(dd3$an2) == FALSE)
+     dd3$type <- paste(dd3$type, "sub", sep = "_")
 
@@
        #geom_line(data = dd3, linetype = "dashed", size = 1.1, mapping = aes(x = time, y = cort, by = type, color = type)) +
-       geom_smooth(data = dd3, mapping = aes(x = time, y = cort, by = type, color = type, fill = type), linetype = "dashed", span = 0.9) +
-       scale_fill_manual(values = c("slateblue", "orange")) +
-       scale_color_manual(values = c("slateblue", "orange")) +
-       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C") +
+       geom_point(data = dd3, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6, size = 0.8) +
+       geom_smooth(data = dd3, size = 0.8, mapping = aes(x = time, y = cort, by = type, color = "black", fill = type), linetype = "twodash", span = 0.9) +
+       scale_fill_manual(values = c("#968BDC", "#FFC04C")) +
+       scale_color_manual(values = c("black", "slateblue", "#968BDC", "orange", "#FFC04C")) +
+       annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "E") +
        #theme(legend.position = c(.6, .2)) +
@@
        ylab("") +
-       geom_point(data = dd3, mapping = aes(x = time, y = cort, color = type, fill = type), alpha = 0.6) +
        ggtitle("Weighted sampling")
 
-     ggpubr::ggarrange(p1a, p1b, p1c, nrow = 1)
+     #ggpubr::ggarrange(p1a, pxa, pxa2, p1b, p1c, nrow = 2, ncol = 3)
+
+    grid.arrange(grobs = list(p1a, pxa, pxa2, p1b, p1c),
+                  widths = c(1, 1, 1, 1, 1, 1),
+                  layout_matrix = rbind(c(1, 1, 2, 2, 3, 3),
+                                        c(NA, 4, 4, 5, 5, NA)))
 
@@
 
-In this case, the standard sampling scheme performs very poorly (figure \@ref(fig:treatment-ex)A), with no differences detectable between the two groups, despite the fact that the treatment group reaches it's maximum value on average 12 minutes (~40%) faster than the control group.
-In contrast, both the random sampling and weighted sampling schemes detect differences in the shape of the acute response (figure \@ref(fig:treatment-ex) B & C).
-In this particular scenario, there is no clear difference between these two approaches.
+## DISCUSSION
+
+I demonstrate that simple simulation tools can produce datasets that closely match empirically observed acute glucocorticoid responses.
+Once developed, simulations can then be used to explore a wide range of hypothetical scenarios and to guide empirical studies.
+I explore a few possible scenarios with a limited range of parameters here.
+Nevertheless, even the simple demonstrations included in this paper suggest several ways that simulation could help move studies of physiological flexibility forward.
+One of the main benefits of simulating datasets is identifying unmeasured properties and assumptions of currently available data that can become targets for future empirical work.
+
+For example, I demonstrate that the ability to accurately measure a single glucocorticoid trait of interest (maximum glucocorticoid level) is strongly impacted by the amount of between-individual variation in the population and more subtly influenced by between-individual variation in speed (the time required to reach maximum levels) and by the timing of sample collection relative to the average population peak timing.
+In one sense, this result is unsurprising because it is intuitive that large differences in maximum glucocorticoids should be easier to detect, but there are important consequences of this fact for interpreting studies that seek to link between-individual variation in the magnitude of the stress response with other traits.
+The magnitude of the acute stress response often varies substantially across life history stages [@wingfield1992].
+Even if study designs are identical it will be easier to accurately measure individual glucocorticoid traits under some conditions with the same level of sampling.
+This simple scenario shows that the same sampling regime will perform better or worse depending on the combination of glucocorticoid regulation parameters in the population being studied for statistical, rather than biological, reasons.
+
+The pattern of within-individual covariation between separate aspects of the acute glucocorticoid response (e.g., speed and maximum values) can also have important effects on the interpretation of empirical results.
+I found that covariation between speed and magnitude of the acute stress response and the relative amount of variation in each of these traits have profound effects on the ability to accurately measure either single component.
+This result occurs because measures taken at any given time point reflect a mixture of speed and magnitude.
+Empirical work specifically designed to assess covariation and variance at different times could help to understand what conclusions we can reasonably draw from available data.
+This simulation also clearly demonstrates that measuring timing components of the acute glucocorticoid response (e.g., the time required to reach maximum) is much more challenging over a wide range of conditions than measuring maximal levels.
+Although a great deal of empirical work has focused on estimating differences in the initial rate of increase in glucocorticoids [e.g., @cockrem2007; @cockrem2005; @wingfield1992], much less is known about between-individual variation in the time required to reach maximal levels.
+
+Beyond the specifics of this particular example, what these results demonstrate is that understanding what aspect of the glucocorticoid response is being measured by any particular study design depends on extensive knowledge of the overall shape and amount of variation in different aspects of the acute stress response.
+There is still a great need for observational sampling to characterize within- and between-individual characteristics of glucocorticoid responses, particularly when new species or contexts are being studied.
+An increased focus on descriptive data collection can subsequently contribute to more targeted and powerful hypothesis tests and better parameterized simulations.
+
+I also showed that the amount of within- and between-individual variation in glucocorticoid traits directly impacts the ability to detect statistical support for known relationships between the trait of interest and fitness.
+The fact that low between-individual variation and high within-individual variation in maximum glucocorticoids make it harder to detect true glucocorticoid-fitness relationships across a wide range of conditions has important consequences for interpreting empirical results.
+Many studies have demonstrated different relationships (or lack thereof) between glucocorticoids and fitness at different life history stages [@vitousek2018; @bonier2009b], but it is also well known that the absolute amount of between individual variation in glucocorticoid traits varies considerably at different stages [@wingfield1992].
+My simulation demonstrates that the power to detect true relationships will differ drastically across these conditions even with identical study designs and samples sizes; thus, apparent differences in glucocorticoid-fitness relationships across seasons or stages can easily arise as statistical artefacts when between-individual variation in hormones also differs across contexts.
+Great care may be needed to conclusively differentiate true differences in these relationships from statistical artefacts.
+Of course, it may be common for physiological traits have stronger direct impacts on fitness in some contexts than others [@bonier2009b], but simulation guided studies can help to ensure that the statistical approach to detect differences is equally powerful in different contexts.
+
+Finally, I show that many standard sampling schemes perform poorly in some situations and for certain questions.
 A few clear takeaways can be derived from these results.
 First, while strict standardization of the timing of samples has some clear advantages, it also comes with costs and likely makes it nearly impossible to detect certain types of variation between groups or individuals.
-In this case, standardized sampling performed much worse than the other two approaches despite the fact that the analysis included 50% more data; it should be clear that no amount of additional sampling would allow that approach to detect this particular pattern of between group differences.
-Second, while it may be very difficult to accurately estimate the full shape of the acute stress response for *individuals*, the sampling schemes shown here demonstrate that it should be possible to describe these shapes accurately for groups (e.g., treatments, species, different contexts) even without extraordinarily large sample sizes.
-A similar argument about the power of randomly timed sampling has been put forward in the function valued trait literature [@gomulkiewicz2018], but this type of sampling scheme is rarely used in evolutionary endocrinology research.
-It is perhaps unsurprising that the few empirical papers that have emphasized the importance of different time courses (rather than only maximum) of the stress response have often focused on between group comparisons or investigated variation in the exact sampling time between individuals [e.g., @baugh2013; @small2017]
-
-This simulation is particular to a single very specific scenario, but a similar scenario could be designed for any number of studies and any number of predictions about how the speed, scope, or other attributes of the glucocorticoid response are expected to change with a treatment or between different groups or species.
-Clearly, when estimating the timing of peak glucocorticoids, a simple baseline plus induced sampling scheme is sub optimal, but this scheme may be perform well in other situations where the maximum value is the target and there is relatively little variation in response time.
-Creating simulations like this before studies are conducted has the potential to increase the efficient use of researches time and funds, but also forces researches to think explicitly about quantitative predictions ahead of time.
+It should be clear that no amount of additional sampling would allow that approach to detect the difference in time to reach maximum glucocorticoid levels developed in the scenario here.
+Second, while it may be very difficult to accurately estimate the full shape of the acute stress response for *individuals*, alternative sampling schemes with random or weighted sample timing make it possible to describe these shapes accurately for groups (e.g., treatments, species, different contexts) even without extraordinarily large sample sizes.
+A similar argument about the power of randomly timed sampling has been put forward in the function valued trait literature [@gomulkiewicz2018], but this type of sampling scheme is rarely used in evolutionary endocrinology research [but see, @vitousek2022].
+It is perhaps unsurprising that empirical papers that have emphasized the importance of different time courses (rather than only maximum) of the stress response often focus on between group comparisons or investigate variation in the exact sampling time between individuals [e.g., @baugh2013; @small2017].
+These alternative sampling schemes also come with drawbacks (e.g., the inability to directly compare individuals sampled at different times) and I don't suggest that they are universally better options than standardization; rather, different schemes will perform better or worse given the specific scenario and the exact hypothesis being tested.
+
+The simulation of different sampling schemes in particular addresses a single specific scenario, but a similar scenario could be designed for any number of studies and any number of predictions about how the speed, scope, or other attributes of the glucocorticoid response are expected to change with a treatment or between different groups or species.
+Clearly, when estimating the timing of peak glucocorticoids, a simple baseline plus induced sampling scheme is sub optimal, but this scheme may perform well in other situations where the maximum value is the target and there is relatively little variation in response time.
+Creating simulations like this before studies are conducted has the potential to increase the efficient use of researchers time and funds, but also forces researchers to think explicitly about quantitative predictions ahead of time.
 These simulations could be included as part of a study pre-registration, grant application, or registered report to demonstrate exactly what data collection and analysis approaches are planned and to justify those decisions.
-
-## DISCUSSION
-
-While there has been increasing interest in understand within- and between-individual variation in the acute glucocorticoid response in recent years [@hau2016; @lema2013; @wada2014; @taff2016], the methods and data available to tackle these questions have changed relatively little.
-Many sophisticated statistical tools are now available and clear arguments have been made about the need to apply these approaches to endocrine traits, but relatively few empirical studies have effectively used these tools.
-Arguably, the biggest roadblock at the moment is the limited availability of empirical data needed to test hypotheses.
-Simulation offers one way forward, by allowing for more efficiently designed studies and by allowing researches to identify when the question of interest can *in principle* be answered with a given study design.
-Ideally, conceptual papers, empirical work, and simulation will proceed together to make progress in this field.
-The tools presented here only scratch the surface of the ways that data simulation can be applied to address pressing questions in evolutionary endocrinology.
-
-Nevertheless, even the simple demonstrations included in this paper suggest several ways that simulation could help move the field forward.
-One of the main benefits of simulating datasets is identifying unmeasured properties and assumptions of currently available data that can become targets for empirical work.
-For example, I demonstrated that the covariation between different components of the acute stress response and the relative amount of variation in each of these can have profound effects on the ability to accurately measure any single component.
-Empirical work specifically designed to assess covariation and variance at different times could help to understand what conclusions we can reasonably draw from available data.
-One takeaway from these simulations is that variation in glucocorticoid-fitness relationships across seasons or life history stages can easily arise as a statistical artefact when between-individual variation in hormones also varies across the contexts.
-The simulation exploring different sampling designs also suggests that there are potentially gains to be made by considering more diverse sampling designs tailored to the particular research question and study system.
-While standardized sample collection timing has allowed for large scale comparisons in this field [@vitousek2019], it also creates clear blind spots to certain types of variation between groups.
-
-In addition to providing insight in its own right, simulation has great potential to hone the design of future empirical studies by allowing for a principled analysis of various study design options and choices before costly data are collected and before animals are needlessly disturbed.
-For example, I showed that one of the most common sampling design schemes has essentially no ability to detect a difference in the speed of increase between two groups if they do not also differ in maximum values.
-It is perhaps not surprising to find that there is little published evidence for differences in the speed of the acute response when most study designs employed to date cannot *in principle* detect those differences.
 Across a wide range of disciplines there has been an increasing push for pre-registration, reproducible research, and transparent research practices [@odea2021].
@@
 
-This package and paper is meant only as an initial exploration of the ways that simulation can be applied to evolutionary endocrinology.
-I have no doubt that many more scenarios and complications could be added on to each of the simple examples presented here.
-Furthermore, there is ample room to create more sophisticated simulations that incorporate realistic mechanistic processes or interactions with other molecules and other components of the stress response system.
-I hope that this work will be a starting point to build and improve on as we work to understand the importance of variation in these flexible response systems.
+While there has been increasing interest in understanding within- and between-individual variation in the acute glucocorticoid response in recent years [@hau2016; @lema2013; @wada2014; @taff2016], the methods and data available to tackle these questions have changed relatively little.
+Many sophisticated statistical tools are now available and clear arguments have been made about the need to apply these approaches to endocrine traits, but relatively few empirical studies have effectively used these tools [but see, @furtbauer2015; @houslay2022].
+Arguably, the biggest roadblock at the moment is the limited availability of empirical data needed to test hypotheses.
+Although there is a rich history of empirical work focused on understanding the acute glucocorticoid response [e.g., @wingfield1992; @cockrem2002; @romero2002; @vitousek2019], many current questions require a level of repeated sampling within- and between-individuals that is difficult to achieve.
+Simulation offers one way forward, by allowing for more efficiently designed studies, by allowing researchers to identify when the question of interest can *in principle* be answered with a given study design, and by providing a tool to translate recent theoretical advances [e.g., @luttbeg2021; @grindstaff2022; @taborsky2021] into tractable, hypothesis driven empirical studies.
+Ideally, conceptual papers, empirical work, theory, and simulation will proceed together to make progress in this field.
+The tools presented here only scratch the surface of the ways that data simulation can be applied to address pressing questions in evolutionary endocrinology.
 
@@
 I would like to thank Michaela Hau and her lab for discussions about the ideas presented in this project.
-John Wingfield also provided an insightful overview of the history of study design in field based studies of the acute stress responses and pointed me to key references.
+John Wingfield also provided an insightful overview of the history of study design in field-based studies of the acute stress responses and pointed me to key references.
 Finally, I thank Maren Vitousek and her lab members for feedback and discussion on early versions of this project.

### 2022-06-30 9_old_not_needed/1_RWBL_script.R
Fixing up repo.

@@ -1,362 +1,0 @@
-### Script for speed vs. scope manuscript
-# Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
-
-
-## Load packages ----
-    pacman::p_load(ggplot2, here, tidyverse, viridis, gridExtra)
-
-## Load RWBL field data ----
-        rw_dat <- read.delim(here::here("1_raw_data/rwbl_data.txt"))      
-        rw_dat <- subset(rw_dat, rw_dat$Bird_no != 10)    # This bird is missing most data
-        rw_dat <- subset(rw_dat, rw_dat$band != "1372-33452")
-    
-    ## Replacing labeling on a couple rows to make them conform to level names
-        rw_dat$LH.stage<-gsub("mid-breeding","early-breeding",rw_dat$LH.stage)	
-        rw_dat$LH.stage[38]<-"early-breeding"
-    
-    ## Make a list of colors in order of rows to plot LH stages in different colors
-        collist<-rep(NA,nrow(rw_dat))
-        for(i in 1:nrow(rw_dat)){
-          if(rw_dat$LH.stage[i]=="early-breeding"){collist[i]<-"lightblue"}
-          if(rw_dat$LH.stage[i]=="late-breeding"){collist[i]<-"gray60"}
-          if(rw_dat$LH.stage[i]=="molt"){collist[i]<-"red"}
-        }
-
-## Analysis of RWBL field data ----
-        ## Loop through data frame to calculate slope and max cort achieved for each bird's samples
-            for(i in 1:nrow(rw_dat)){
-              # Set up dataframe for each bird with just the first three samples (less than 10 minutes)
-                calc <- as.data.frame(matrix(nrow = 3, ncol = 3))
-                colnames(calc) <- c("time", "cort", "lcort")
-                calc[, 1] <- t(rw_dat[i, c("tube1_tsec", "tube2_tsec", "tube3_tsec")])
-                calc[, 2] <- t(rw_dat[i, c("tube1_cort", "tube2_cort", "tube3_cort")])
-                calc[, 3] <- log(calc[, 2] + 1)
-                rownames(calc) <- c(1, 2, 3)
-                calc <- subset(calc, calc$time < 600 & is.na(calc$cort) == FALSE)
-              # Calculate the slope over the first 8 minutes from a simple linear regression. Both normal and log scale.   
-                calc$time2 <- calc$time / 60
[…]

### 2022-06-30 9_old_not_needed/2_cortResponse_script.R

@@ -1,154 +1,0 @@
-### Script for speed vs. scope manuscript
-# Written by Conor Taff; cct63@cornell.edu; cct663@gmail.com
-
-## Load packages ----
-    pacman::p_load(ggplot2, here, MASS, MBESS, tidyverse, viridis, gridExtra, rptR, simcoRt, sp)
-
-## Initial demonstration ----
-      set.seed(100)   # makes it reproducible by using same random seed
-      demo <- cort_sim2()
-      p <- plot_cort_sim(demo)
-      ggsave(here::here("3_r_scripts/demo.png"), p, device = "png", width = 10.2, height = 7.5)
-      
-      ggsave(here::here("3_r_scripts/demo_fullx.png"), 
-        plot_cort_sim_1(demo, x_max = 160, y_max = 100),
-        device = "png", width = 9, height = 5, units = "in")
-      
-## Variation in speed vs. scope ----
-    set.seed(955)
-    d1 <- cort_sim2(cort_sim1(speed_mu = 25, speed_sd = 0, max_sd = 0.2))  
-    d2 <- cort_sim2(cort_sim1(max_sd = 0.2))
-    d3 <- cort_sim2(cort_sim1(max_sd = 0, speed_sd = 10))
-    p1 <- plot_cort_sim_1(d1)
-    p1 <- p1 + annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "A")
-    p2 <- plot_cort_sim_1(d2)
-    p2 <- p2 + annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "B")
-    p3 <- plot_cort_sim_1(d3)
-    p3 <- p3 + annotate(geom = "text", x = -Inf, y = Inf, hjust = -.5, vjust = 1.5, label = "C")
-    ggsave(here::here("3_r_scripts/relative_variation.png"),
-           ggpubr::ggarrange(p1, p2, p3, nrow = 1),
-           device = "png", width = 9.2, height = 3, units = "in")
-                    
-## Simulate data ----
-    #plot_cort_sim(cort_sim2(cort_sim1(n = 60, speed_mu = 18, speed_sd = 3, base_mu = 9, base_sd = 3, cor_base_max = 0.3, cor_base_speed = -0.5, cor_speed_max = -0.7, cor_max_return = 0.4), 
-    #                        bleed_times = c(1, 15, 30), sample_times = 1))
-          
[…]

### 2022-06-30 3_r_scripts/Revision.Rmd
Editing text.

@@
 For illustration purposes, I refer to simulated glucocorticoid responses, but the simulation applies equally well to any physiological mediator of a rapid response.
-The code for the complete simulation along with full documentation of each function and argument can be accessed in a GitHub repository with a current version permanently archived on Zenodo (LINK).
+The code for the complete simulation along with full documentation of each function and argument can be accessed in a GitHub repository with a current version permanently archived on Zenodo (https://doi.org/10.5281/zenodo.6784207).
 The package can be installed directly within the R environment using the following command.
@@
 Each of the scenarios described here uses the basic simulation functions described above with input parameters adjusted to address the question of interest.
-A complete set of reproducible code to create all of the examples presented here is available on GitHub and permanenty archived at Zenodo (https://github.com/cct663/speed_vs_scope).
+A complete set of reproducible code to create all of the examples presented here is available on GitHub and permanenty archived at Zenodo (https://doi.org/10.5281/zenodo.6784203).
 

### 2022-07-01 3_r_scripts/Revision.Rmd
minor changes

@@
 These validations were considered essential to characterize key parameters of the acute response for each group being studied (i.e., baseline, rate of increase, maximum level, time of peak, and area under the curve; John Wingfield, *personal communication*).
-The challenge of estimating these parameters becomes much more difficult when trying to describe the response for an individual animal rather than for a group, because glucocorticoids can often only be measured at two or three time points and only a small number of times per animal [e.g., @vitousek2018].
+Indeed, there is a long and rich history of empirical work characterizing differences in each aspect of the acute glucocorticoid response [e.g., @cockrem2002; @wingfield1992; @breuner1999; @love2003] and empirical data has contributed to a variety of conceptual models of acute glucocorticoid regulation [e.g., @wingfield1998; @romero2009].
+More recently, mathematical models have been used to generate predictions about how flexibility and variation in different parts of the response system might respond to particular conditions [e.g., @luttbeg2021; @grindstaff2022; @taborsky2021].
+However, the challenge of estimating these parameters and of decomposing within- and between-individual variance becomes much more difficult when trying to fully describe the response for an individual animal rather than for a group, because glucocorticoids can often only be measured at two or three time points and only a small number of times per animal [e.g., @vitousek2018].
 Because these studies require an estimate for each individual, the solutions used by older studies that added additional animals to allow for sampling at more time points are not available.
-
-INSERT HISTORY AND CITATIONS FROM COCKREM
 
@@
 Indeed, several discussions in recent years about methods such as the '3 minute rule' and the relative merits of 'area under the curve' versus time point measures of glucocorticoids are fundamentally related to a recognition of the importance of understanding variation in the functional shape of stress responses and whether different components of that shape covary within individuals [e.g., @cockrem2002; @small2017].
-While a great deal of empirical work has focused on characterizing the rate of initial increase [CITE], there has been relatively little work on understanding individual differences in the time required to reach maximal levels, and this attribute of speed is harder to estimate [@taff2022].
+While a great deal of empirical work has focused on characterizing the rate of initial increase [@cockrem2005; @cockrem2006], there has been relatively little work on understanding individual differences in the time required to reach maximal levels, and this attribute of speed is harder to estimate [@taff2022].
 
@@
 One way to accomplish these goals is to integrate empirical work with simulations, but to my knowledge no studies of physiological flexibility have developed simulations of the acute stress response that address the issues discussed above.
-
-There is also a growing number of mathematical models of the  (Luttbeg, Taborsky)
 
@@
 As such, there are many possible applications and here I briefly highlight a few possibilities.
-First, I demonstrate that the simulation can produce datasets that are qualitatively similar to empirical data on acute glucocorticoid responses.
+Initially, I demonstrate that the simulation can produce datasets that are qualitatively similar to empirical data on acute glucocorticoid responses.
 Next, I explore three specific scenarios with the simulation that highlight challenges to empirical progress in this field.
@@
 These questions are all important for understanding physiologically flexibility and are difficult to fully address empirically, but this is by no means an exhaustive list of the questions or possible parameter permutations that could be explored with simulation.
-Finally, I demonstrate how the simulation tools can be used to evaluate the performance of different experimental designs given a set of logistical limitations (e.g., number of samples that can be collected) and a proposed hypothesis) and discuss how this procedure could be used as a planning tool to increase the power and reproducibility of future empirical work.
+Finally, I demonstrate how the simulation tools can be used to evaluate the performance of different experimental designs given a set of logistical limitations (e.g., number of samples that can be collected) and discuss how this procedure could be used as a planning tool to increase the power and reproducibility of future empirical work.
 
